### 2023-10-15 US/before_body.tex
Initial commit

\noindent \textbf{Keywords:} Uncertainty, Monetary transmission, Non-linearity, Sign restriction, Vector-Autoregression \linebreak
\noindent \textbf{JEL codes:} E32, E52, C32 
\newpage
\bgroup
\large{\textbf{Acknowledgements}}\\
\egroup

I would like to express my gratitude to a number of people who helped me see this project through. First and foremost to my PhD advisor, István Kónya for his supervision and guidance over the years of my PhD; to Tohid Atashbar, my former IMF internship supervisor for his co-authorship on the IMF Working Paper, as well as for his suggestions on this paper; to Gergely Gánics for his suggestions on fine-tuning the methodology used in this paper; to Anh Nguyen and Dániel Baksa for their suggestions on alternative specifications that would support my results; as well as to former colleagues from my IMF internship and current colleagues from academia, who provided insightful comments during seminar presentations.

\newpage 
\bgroup
\hypersetup{linkcolor = black}
\renewcommand{\baselinestretch}{1}\normalsize
\tableofcontents
\newpage
\listoffigures
\egroup
\newpage

### 2023-10-15 US/depo/old_version/tex/abstract2.tex
Initial commit

\documentclass[12pt,a4paper]{article}
%\DeclareMathSizes{12}{14}{10}{6}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage[outer=25mm,inner=35mm,top=25mm,bottom=25mm]{geometry}

\usepackage{indentfirst}
\usepackage{colortbl}
\usepackage{amsmath}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage[export]{adjustbox}
\usepackage{listings}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{apacite}

\frenchspacing
\linespread{1.5}

\begin{document}
	\noindent {\Large Corvinus University of Budapest\\
		Institute of Economics}\\

	\vspace{35mm}

	\begin{center}
		{\huge The asymmetric effect of uncertainty on monetary transmission}\\
	\end{center}

	\vspace{90mm}

	\begin{flushright}
		{\Large Péter Horváth}\\
		%{\Large Makrogazdasági- és piacelemző szakirány}\\
		{\Large 2022}
	\end{flushright}

	\vspace{10mm}
	\begin{center}
		{\Large Supervisor : István Kónya}
	\end{center}

	\thispagestyle{empty}

	\pagebreak
	\setcounter{tocdepth}{2}
	%	\tableofcontents
	%\listoftables
	%\listoffigures

	\pagebreak
	\begin{center} \section*{Abstract} \end{center}

Recently in macroeconomic research, the impact of uncertainty gained much traction as an important factor in studying the business cycle of economies. The pre-existing literature is already rich and diverse as papers cover among others how a country’s own uncertainty impacts its business cycle, how world (or US) level uncertainty spills over to smaller economies, and the topic of asymmetries is has been shown that uncertainty shocks in recession times hit harder and higher uncertainty leading to supply and demand shocks having a more severe impact on the economy. In this paper, I aim to contribute to this already rich base of literature by exploring the asymmetric relationship between monetary policy and uncertainty. More specifically, I will be studying how asymmetrical of an effect uncertainty has on the efficacy of the monetary transmission mechanism of the interest rate channel., i.e.: I am interested in how much an interest rate shock is able to dampen the inflation rate, and what is the cost of stopping inflation in terms of economic growth during times of high uncertainty. The notion is that the transmission mechanism should be less effective in achieving its goal when uncertainty is high, thus highlighting the importance of managing expectations in order to control economic uncertainty. 
As the question of interest discussed above is of an empirical nature, I will be using the toolbox of vector autoregressive modelling. To get the highest frequency and quality data possible, I will rely on monthly US data from the FRED database using the FFR, the CPI and the Industrial Production Index (as a proxy of GDP growth).As the price puzzle is a challenging problem in the literature of monetary macroeconomics, I will rely on using the sign restriction identification scheme in a Bayesian VAR setting. To study the nonlinear relationship between uncertainty and monetary policy, I will first establish a baseline estimation, by running the model on the whole time series available, so that I have base for comparison on the transmission mechanism. Next, I will identify periods in the time series where uncertainty was at its peak and most volatile. Over these selected periods I will estimate models (using the same estimation methods) in 5 year (60 month) rolling windows. To compare results, I will plot the impulse responses of an interest rate shock from the baseline and high uncertainty models and compare them, in the latter case – as it is comprised of several model estimates – I will use median impulse responses. The results will show some asymmetries, as an interest rate shock will produce a considerably larger drop in economic growth, however there is no substantial change in the impact on inflation.\\

\bigskip

\noindent \textbf{Keywords:} Uncertainty, Monetary transmission, Asymmetry, Sign restriction, Bayesian VAR

	\pagebreak

\end{document}

### 2023-10-15 US/depo/old_version/tex/uncertainty_project.tex
Initial commit

\documentclass[12pt,a4paper]{article}
%\DeclareMathSizes{12}{14}{10}{6}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage[outer=25mm,inner=35mm,top=25mm,bottom=25mm]{geometry}

\usepackage{indentfirst}
\usepackage{colortbl}
\usepackage{amsmath}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage[export]{adjustbox}
\usepackage{listings}
\usepackage{wrapfig}
\usepackage{float}
\usepackage[
backend=biber,
style=mla,
sorting=ynt
]{biblatex}
\usepackage{optidef}
\addbibresource{ref.bib}

\frenchspacing
\linespread{1.5}

\begin{document}
	\noindent {\Large Corvinus University of Budapest\\
		Institute of Economics}\\

	\vspace{35mm}

	\begin{center}
		{\huge The asymmetric effect of uncertainty on monetary transmission}\\
	\end{center}

	\vspace{90mm}

	\begin{flushright}
		{\Large Péter Horváth}\\
		%{\Large Makrogazdasági- és piacelemző szakirány}\\
		{\Large 2022}
	\end{flushright}

	\vspace{10mm}
	\begin{center}
		{\Large Supervisor : István Kónya}
	\end{center}

	\thispagestyle{empty}

	\pagebreak
	\setcounter{tocdepth}{2}
%	\tableofcontents
	%\listoftables
	%\listoffigures

	\pagebreak
\begin{center} \section*{Abstract} \end{center}

This paper discusses the asymmetric effect of uncertainty on the monetary transmission mechanism. The notion is that during periods characterized by high uncertainty spikes, the transmission mechanism of the interest rate channel is less effective, thus implying the importance of using forward guidance alongside interest rate decisions. To study this nonlinear relationship, I use a rolling window approach to estimate sign restricted VAR models over these periods in question. The results of the estimation show a minor difference in the impact of an interest rate shock during high-uncertainty periods, compared to the estimate over the whole time series.\\

\bigskip

\noindent \textbf{Keywords:} Uncertainty, Monetary transmission, Asymmetry, Sign restriction, Vector-Autoregression

\pagebreak

\section{Introduction}
Over the recent years, uncertainty and asymmetries are two subjects that gained considerable traction in macroeconomics research. Both fields of research already have a rich literature, with there being a consensus on uncertainty shocks having an impact resembling that of a negative demand shock on economic activity (\textcolor{blue}{\cite{CARRIERESWALLOW2013316}}, \textcolor{blue}{\cite{COLOMBO201339}}, \textcolor{blue}{\cite{CALDARA2016185}}, \textcolor{blue}{\cite{CHENG2018305}}, \textcolor{blue}{\cite{BONCIANI2020102236}}, \textcolor{blue}{\cite{NILAVONGSE2020108765}}); and several papers have shown that macroeconomic shocks can have different impacts depending on the state of the economy.
\textcolor{blue}{\cite{RePEc:eee:jmacro:v:69:y:2021:i:c:s0164070421000379}}, \textcolor{blue}{\cite{RePEc:fip:fedker:00016}}, \textcolor{blue}{\cite{RePEc:cup:macdyn:v:20:y:2016:i:05:p:1219-124600}} and \textcolor{blue}{\cite{RePEc:jae:japmet:v:19:y:2004:i:5:p:551-565}} all take slightly different approaches, however all papers find that the impact of (negative) macroeconomic shocks is a sharper decline when uncertainty is high. Similar processes can be observed when carrying out the analysis with respect to the economy's position in the business cycle. \textcolor{blue}{\cite{RePEc:knz:dpteco:1402}} and \textcolor{blue}{\cite{RePEc:een:camaaa:2020-72}} show that uncertainty shocks have larger impact during economic contractions, and \textcolor{blue}{\cite{RePEc:gro:rugsom:98c36}} finds that monetary transmission is amplified during recessionary periods in the US and Germany. Another field of research on asymmetric dynamics is finding asymmetry in monetary policy-making itself. \textcolor{blue}{\cite{Lin+2021+425+447}} finds that macroeconomic aggregates react asymmetrically to positive and negative monetary policy shocks. \textcolor{blue}{\cite{doi:10.1080/1331677X.2018.1481445}} links uncertainty and asymmetries in monetary policy by assuming a difference in the policy reaction function in over the business cycle. \\

With this paper, I aim to contribute to the pre-existing literature, by linking the effect of uncertainty to the efficacy of monetary transmission. The results could have key implications for policy making, more specifically, it could highlight the importance of forward guidance measures, especially during times, when economic uncertainty is at its peak. \\

The rest of the paper will be outlined as follows: Section 2 gives a brief overview of the data used, Section 3 discusses the empirical strategy, in Section 4 I discuss the results and Section 5 concludes.

\section{Data}

For the purposes of the analysis I will be estimating VAR models with three variables, the federal funds rate, consumer price index, and industrial production index. All data are monthly from ranging from January 1985 to December 2019 and all are retrieved from the federal reserve database. \\

\begin{center}
\begin{figure}[h!]
	\includegraphics[width=0.9\linewidth]{plot1.png}
		\caption{Time series plots of the Federal Funds Rate, Consumer Price Index and Industrial Production Index for the USA.}
\end{figure}
\end{center}

As most macroeconomic variables, these data series are not stationary either, which is not desirable for the stability of the VAR models. For this reason I will be using the first difference of the industrial production index and the CPI. However, I will use the FFR at its level. This way I have better stability in my models and all variables have clear economic interpretation, as the two first-differenced data series can be considered as monthly growth rates of production and inflation respectively.\\

Another key variable I will be using as a "quasi transition variable" is a measure of uncertainty. This however is not a trivial metric, as several data series can be interpreted as a measurement of economic turmoil. The most commonly used are among others the VIX index and the TED spread. 
\begin{center}
	\begin{figure}[h!]
		\includegraphics[width=0.9\linewidth]{plotcompare.png}
		\caption{Time series plots of the VIX index, TED spread and the Economic Policy Uncertainty index, log scales.}
	\end{figure}
\end{center}
The former derives from the market expectations on the price changes of the S\&P500 index, while the latter is the difference between the 3 month London interbank rate and the 3 month US treasury bill. Both are widely used in measuring financial stress or uncertainty, however in recent years in the literature on uncertainty, news based indices have gained a lot of traction. One such is the Economic Policy Uncertainty index created by \textcolor{blue}{\cite{baker2016measuring}}.

\begin{center}
\begin{figure}[h!]
	\includegraphics[width=0.9\linewidth]{epu.png}
		\caption{Time series plot of the BBD Economic Policy Uncertainty Index. The shaded areas indicate time periods with high uncertainty spikes over which the Bayesian VAR models will be estimated for the purposes of this analysis.}
\end{figure}
\end{center}
\noindent The index is primarily news based, on uncertainty related keywords in influential papers in the US. The index also takes into account temporary tax measures reported by the Congressional Budget Office and the dispersion of individual forecasts from the Federal Reserve Bank of Philadelphia's Survey of Professional Forecasters in the consumer price index and government expenditure. The index is constructed by normalizing the sub-components by their respective standard deviation and the normalized series weighted averaged (with the news based component having the highest weight of 0.5). As this index 1) covers a broader spectrum of expectations than the financial market based uncertainty counterparts, and 2) is more specific to economic policy, I believe it should serve as the most adequate measurement of uncertainty for the purposes of this study.\\

In order to estimate the asymmetric effect of uncertainty on monetary transmission, I will be estimating sign restricted VAR models over periods with high uncertainty spikes. These time periods are July 1990 - January 1994,  August 2000 - January 2004, and January 2008 - January 2014. In order to have a sufficient amount of data points, I will also include 12-12 months before and after these periods, as shown with a lighter gray highlight in the graph above.

\section{Empirical strategy}
An empirical challenge for this analysis was coercing the macroeconomic aggregates to behave the way it is written in all macroeconomics textbooks, i.e. to solve the price puzzle. 
There have already been numerous papers on the topic of solving the price puzzle (e.g.: \textcolor{blue}{\cite{RePEc:eee:moneco:v:51:y:2004:i:7:p:1385-1413}}, \textcolor{blue}{\cite{RePEc:hhs:hastef:0414}}, \textcolor{blue}{\cite{RePEc:spr:empeco:v:46:y:2014:i:2:p:701-731}}, \textcolor{blue}{\cite{RePEc:rba:rbardp:rdp2017-02}}, \textcolor{blue}{\cite{RePEc:aea:aejmac:v:8:y:2016:i:4:p:75-102}}, \textcolor{blue}{\cite{RePEc:aea:aecrev:v:94:y:2004:i:4:p:1055-1084}}), however with the available data series the best empirical approach would be to apply sign restrictions in the model specification. For this purpose, I will be estimating Bayesian VAR models closely following \textcolor{blue}{\cite{RePEc:eee:moneco:v:52:y:2005:i:2:p:381-419}} with a non-informative Gaussian inverted-Wirshart prior. The computation is done by implementing the R package created by \textcolor{blue}{\cite{Danne2015}}. More specifically, out of the two methods proposed, I will be using the "penalty" algorithm instead of the "rejection" algorithm. The latter one is constructed to only accept simulations that exactly satisfy the given sign restrictions, while in the former case, the goodness of fit is also taken into account. This method finds the results that are as close to satisfying the sign restrictions by minimizing a function that penalizes the impulse responses that do not meet the sign restrictions imposed.\\

This empirical strategy however is by no means perfect. The use of sign restricted vector autoregressive models can be criticized for artificially creating the desired model structure. For this, at the time of submitting, I have no solution, as the results would not be robust to a non-restricted model specification due to the price puzzle effect. Another disadvantage of using the sign restriction method is that the system of shocks is not completely identified. This however, is not a major issue in this paper, as only the identification of a shock to the Federal Funds Rate is necessary.\\

In order to be able to find asymmetries in monetary transmission, I will first need a basis for comparison. For this purpose, I will first estimate a baseline model over the whole available time period. Next, I will estimate a number of models in a 5 year (12 month) rolling window over time periods highly affected by uncertainty spikes. All models will be of lag order 1, and the reduced form VAR model can be written as:
\begin{equation}
	B(L)Y_{t} = \alpha_{0} + \epsilon_{t},
\end{equation}
where  $B(L)$ is the matrix of coefficients for at lag L, $Y_{t}$ is the vector of endogenous variables, $\alpha_{0}$ is the vector of constants and $\epsilon_{t}$ is the error term. The ordering of the variables will be
\begin{equation}
	Y_{t} = 
	\begin{Bmatrix}
		FFR_{t} \\
		CPI_{t} \\
		INDPRO_{t}
	\end{Bmatrix}
\end{equation}
\noindent which if important for decomposition, as in the algorithm used, the first variable is considered the shock variable (this is given a positive sign restriction) and the sign restriction of the following variables is set accordingly (negative in both cases).\\

In order to compare and analyze results, I find the use of impulse responses to be the best tool. For the baseline model, this is a quite straightforward process, but for the rolling window estimates, some aggregation will be required. For this purpose, I will be taking point-by-point medians from impulse responses of each estimated model.

\section{Results}
To be able to compare results, first we must define the effectiveness of the monetary transmission mechanism. For our purposes, I will consider the mechanism more effective the better it is at anchoring inflation at a smaller trade-off in economic activity. In other words, the higher impact on the CPI inflation and a lower impact on the industrial growth index.\\

Figure 3 below shows the impulse responses from the benchmark model (first row) and from the rolling window estimation (second row). From the graphs we can see that the impact of the interest rate shock to inflation is quite similar in both cases, with the response in the benchmark model being slightly more persistent. Looking at the FFR shock's response to a shock to itself might give an explanation. We can see that the estimated shock in the baseline model is much larger in volume and much more persistent compared to the results from the rolling window estimation. This by itself is an interesting outcome (which I will get to later), however this seems to partially contradict my prior beliefs, as the relative impact of the shock to inflation would be larger in uncertain times. Looking at the impulse responses of the industrial growth index, we can see a more minor but somewhat more persistent response in the baseline model comparing to the rolling window median response. As to the persistence, a similar explanation can be given, however the drop in output is much sharper in the latter case, which confirms my prior assumption that an interest rate shock during higher uncertainty would slow down the economy much more, thus being less effective. As mentioned before, an interesting result is the path of the FFR shock. In the baseline estimate, the shock seems to be rather persistent,

\begin{center}
\begin{figure}[h!]
	\includegraphics[width=0.9\linewidth]{irfplot1.png}
			\caption{Impulse responses from the benchmark model (first row) and median impulse responses from the rolling window estimation (second row).}
\end{figure}
\end{center}
\noindent indicating a near extreme interest rate smoothing by the Federal Reserve in a general setting, however, in the second case the shock decays much faster, which could indicate - similarly to recession periods - a more active policy style with less focus on smoothing interest rates. This could mean that not only could there be asymmetric effects of monetary policy during periods of high uncertainty, but policy making could also be asymmetrical, similarly to the propositions of \textcolor{blue}{\cite{doi:10.1080/1331677X.2018.1481445}}.\\

To further investigate, I will also be taking a look at the decomposition of the median impulse responses of the rolling window, by taking separate median impulse responses for each of the three highlighted time periods. This can be seen in Figure 4 below.   
\begin{center}
\begin{figure}[h!]
	\includegraphics[width=0.9\linewidth]{irfplot2.png}
			\caption{Separate median impulse responses from rolling widow estimation over the three high-uncertainty periods.}
\end{figure}
\end{center}
By taking a look at the separate IRF-s, we get mixed results. While the uncertainty spikes in each period are similar in volume and pattern, the impulse responses can differ substantially.\\ 
\noindent During the period January 2007 - January 2015, the reasons for uncertainty are pretty self explanatory, as most of this time period is characterized by the devastating economic impact and aftermath of the Great Recession. An explanation as to why the uncertainty index was volatile for such a great length could be the Euro Area crisis which followed shortly after the Great Recession. As Europe is one of the most important trading partners of the US, the overseas crisis must have gained substantial news coverage in the US as well, which would by design cause the index to rise considerably. In the impulse response graphs, we see a fast decaying shock and responses similar to the baseline case in inflation and industrial growth. The path of the FFR response could be explained by the active policy making at the beginning of the period - as it was necessary during the Great Recession. \\
\noindent In the case of time period August 1999 - January 2005, an explanation of higher uncertainty could be the fact that the industrial growth peaked in the early 2000s, after which we see a brief decline. The peak in output is usually considered the end of a boom cycle, and a start of a recession, which could cause a surge in uncertainty. During this period we see a relatively steady growth in CPI, however there were major interest rate cuts in the beginning of the century as a reaction to the signs of economic decline - another reason for the uncertainty index to spike. Following these first few years, the FFR was kept steadily at a rate of about 2.5\%. The estimated responses of inflation and industrial growth in this period are nearly negligible, which could possibly indicate the lack of effect of monetary policy during non/mild-recession times. 
\noindent The explanation for high uncertainty in the period June 1989 – January 1995 is similar to the previous case, as we can see the industrial growth reaching a local maximum in the early 90s. The economic contraction was primarily caused by tightening monetary policy, as we can also see rates hike in Figure 1. Another cause for the contraction was the Tax Reform Act of 1986, leading to a bust in growing real estate prices, thus also disincentivizing investments. The economy coming to a downward shift, the monetary policy tightenings and the tax reform bill altogether would cause much news coverage as well, which by construction causes the EPU index to soar. An interesting observation of this period is that the sharpest drop in inflation and industrial growth can be observed in the impulse responses, as inflation drops by close to 5 percentage points and the industrial growth index drops by an outstanding amount of close to 12 percentage points. The reason why the economic impact of an interest rate shock is much more severe compared to other time periods could be by the underlying economic theory “losing its explanatory potential”, as in later times the key drivers behind the business cycle could be shifting to factors not included in economic models. All in all, we can see close to no asymmetry in the effect of monetary transmission on the inflation rate, and a minor asymmetric effect on output.

\section{Conclusion}
In this paper, I aimed to find asymmetries in the monetary transmission mechanism in high-uncertainty times. For this purpose, I estimated several sign restricted VAR models in five-year rolling windows, and a sign restricted model as baseline. In the results of the estimation, I found close to no asymmetry in the effects of an interest rate shock to inflation, however there is a minor difference in how it impacts output, as the drop in industrial growth in noticeably sharper during periods of high uncertainty. While the results are minor, they have a somewhat important policy making aspect. More specifically, it highlights the importance of forward guidance measures alongside interest rate decisions as a counter-measure to higher economic uncertainty in order to mitigate its effect on the contraction of economic growth.

\printbibliography[title = References]

\end{document}

### 2023-10-15 US/depo/old_version/uncertainty.rmd
Initial commit

---
title: "Untitled"
author: "Horváth Péter"
date: '2022 05 26 '
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, results='hide',fig.keep='all')
```

```{r}
#loading packages
#packages
suppressPackageStartupMessages({
  library(readxl)
  library(tidyverse)
  library(tseries)
  library(forecast)
  library(TSA)
  library(fpp2)
  library(lmtest)
  library(vars)
  library(mFilter)
  library(ggplot2)
  library(tsDyn)
  library(fredr)
  library(VARsignR)
  library(lubridate)
  library(cowplot)
  library(gridExtra)
  library(patchwork)
  library(data.table)
  library(tsibble)
  library(mvnfast)
})

```
# Setup és plotok

A lenti kódsorokkal importálom az adatokat a FRED adatbázisból:

```{r}
#set api key
fredr_set_key("cda47ae66b38ed7988c0a9c2ec80c94f")

#download data
params <- list(
  series_id = c("USEPUINDXD", "DFF", "USACPIALLMINMEI", "INDPRO"),
  frequency = "m",
  observation_start = as.Date("1950-01-01")
)

import  <- pmap_dfr(
  .l = params,
  .f = ~ fredr(series_id = .x, frequency = .y)
) %>%
  dplyr::select(date, series_id, value) %>%
  spread(key = series_id, value = value) %>%
  drop_na() %>% as_tsibble() %>% rename(epu = USEPUINDXD,
                                        ffr = DFF,
                                        cpi = USACPIALLMINMEI,
                                        indpro = INDPRO
                                        )

#select data range
data <- import %>%
  dplyr::select(date, epu, ffr, cpi, indpro) %>%
  drop_na() %>% filter(date <= as.Date("2019-12-01"))
```

Ábra az FFR, CPI és industrial procuction adatsorokról:

```{r}
#plotting the data series
data %>%
  gather(key = "variable", value = "value", ffr, cpi, indpro) %>%
  mutate(variable = case_when(variable == "ffr" ~ "Federal Funds Rate",
                              variable == "cpi" ~ "Consumer Prices",
                              variable == "indpro" ~ "Industrial Production Index"),
         variable = factor(variable, levels = c("Federal Funds Rate", "Consumer Prices", "Industrial Production Index"))) %>%
  ggplot(aes(x = date, y = value)) +
  geom_line() +
  facet_wrap(~variable, scales = "free", nrow = 3) + theme_minimal() +
  labs(x = "",
       y = "") 
```

Még egy kis adat set-up:

```{r}
#generate first differences / growth rates  
data <- data %>%
  mutate(cpi = cpi - lag(cpi),
         indpro = indpro - lag(indpro)) %>% drop_na()

#creating threshold values
data <- data %>%
  mutate(epu50 = median(epu),
         epu75 = quantile(epu, probs = 0.75),
         epu25 = quantile(epu, probs = 0.25))
```

Uncertainty index plot a medián és kvartilis értékekkel:

```{r}
#plot epu index with threshold values
data %>% dplyr::select(date, epu, epu50, epu25, epu75) %>%
  ggplot(aes(x = date, y = epu)) +
  geom_line() +
  geom_hline(aes(yintercept = epu50), color = "darkred") +
  geom_hline(aes(yintercept = epu25), color = "darkgreen") +
  geom_hline(aes(yintercept = epu75), color = "darkgreen") +
  theme_minimal() +
  theme(plot.title = element_text(size = 11, hjust=0.5),
        axis.title.y = element_text(size=11)) +
  labs(x = "",
       y = "",
       title = "BBD economic policy uncertainty index")
```

Függvény amivel az irf értékeket számolom. Ez majd később fog kelleni, hogy a bayesi var szimulációiból legyen grafikonra tehető mediánom, illetve megadott konfidencia-szint melletti intervallum.

```{r}
#function to retrieve impulse responses
get_irf <- function(dim, prob, irflength, irfdata){
  prob <- prob
  x <- irfdata[ , ,dim] %>% t() 
  t <- matrix(nrow = nrow(x), ncol = 1) 
  med <- matrix(nrow = nrow(x), ncol = 1)
  up <- matrix(nrow = nrow(x), ncol = 1)
  dn <- matrix(nrow = nrow(x), ncol = 1)
  for(i in(1:irflength)){
    t[i] <- i-1
    med[i] <- median(x[i,])
    up[i] <- quantile(x[i,], probs = (1-prob))
    dn[i] <- quantile(x[i,], probs = prob)
  }
  irf <- bind_cols(t, med, up, dn) %>%
    rename(t = ...1,
           median = ...2,
           upper = ...3,
           lower = ...4)
  irf
}
```

Paraméterek a becsléshez: két késleltetés volt amivel még találtam az előjelmegkötésnek megfelelő impulzus válaszokat, kell egy szám a  "set.seed" parancsnak. 0.84-es konfidenciaintervallum. az irf adatpontjainak (20 helyett) 21-re állítása pedig csak technikai szépítgetés miatt kell az ábrákhoz.

```{r}
#some parameters for estimation and irfs
nlag <- 2
seed <- 2022
prob <- 0.16
irflength <- 21

```

# Két regime-es modell

Ábra az adatsorokról a regime indicator alapján színezve. Ez alapján szerintem kellően eltér egy sima recession indicator-tól, hogy az eredményeknek legyen hozzáadott értéke. A két regime-es változatban a bizonytalansági index mediánja felett értem a felső, alatta pedig az alsó regime-et.

```{r}
#plot variables coloured by regime indicator
data %>%
  dplyr::select(date, ffr, cpi, indpro, epu, epu50) %>%
  mutate(indicator = ifelse(epu > epu50, 1, 2),
         indicator = case_when(indicator == 1 ~ "High uncertainty regime",
                               indicator == 2 ~ "Low uncertainty regime")) %>%
  gather(key = "variable", value = "value", ffr, cpi, indpro) %>%
  mutate(variable = case_when(variable == "ffr" ~ "Federal Funds Rate",
                              variable == "cpi" ~ "Consumer Prices",
                              variable == "indpro" ~ "Industrial Production Index"),
         variable = factor(variable, levels = c("Federal Funds Rate", "Consumer Prices", "Industrial Production Index"))) %>%
  ggplot(aes(x = date, y = value, color = indicator, group = 1)) +
  geom_line(size = 1) +
  facet_wrap(~variable, scales = "free", nrow = 3) + theme_minimal() +
  labs(x = "",
       y = "") + 
  theme(legend.title = element_blank()) 
```

Két regime-es modell. Először a felső regime sokkjait azonosítom, amely az első három adatsorom, ezért az első három változóra teszek megkötést. Alatta pedig az alsó regime-re megismétlem, itt a 4-6. változóta teszem a megkötést. Az elsőben az első, a másodikban pedig a 4. változó van sokkolva (mindkét esetben az FFR a megfelelő regime indicator-ral szorozva).

Az előjelmegkötés az sokk utáni 1.-6. periódusig tartja meg, amit a "KMIN" és "KMAX" argumentumok adnak meg. A függvény 200 posterior eloszlást szimulál és minden eloszlásra 200 impulzus választ (draws és subdraws), ideális esetben pedig 1000 szimulált impulzus választ tart meg (a rejection algoritmus jellemzően kevesebbet talál).

```{r}
# fit 2 regime model
data_2regime <- data %>%
  dplyr::select(date, ffr, cpi, indpro, epu, epu50) %>%
  mutate(indicator = ifelse(epu > epu50, 1, 0),
         indicator_inv = ifelse(indicator == 1, 0, 1),
         ffr_up = ffr*indicator,
         cpi_up = cpi*indicator,
         indpro_up = indpro*indicator,
         ffr_dn = ffr*indicator_inv,
         cpi_dn = cpi*indicator_inv,
         indpro_dn = indpro*indicator_inv) %>% as_tibble() %>%
  dplyr::select(ffr_up, cpi_up, indpro_up, ffr_dn, cpi_dn, indpro_dn) %>% ts()

set.seed(seed)
constr_up <- c(+1, -2, -3)
model_up <- uhlig.reject(Y=data_2regime, nlags=nlag, draws=200, subdraws=200, nkeep=1000, KMIN=1,
                         KMAX=6, constrained=constr_up, constant=TRUE, steps=21)
irfs1 <- model_up$IRFS

set.seed(seed)
constr_dn <- c(+4, -5, -6)
model_dn <- uhlig.reject(Y=data_2regime, nlags=nlag, draws=200, subdraws=200, nkeep=1000, KMIN=1,
                       KMAX=6, constrained=constr_dn, constant=TRUE, steps=21)
irfs2 <- model_dn$IRFS
```

Kirajzolom az impulzus válasz függvényeket:

```{r}
ffr_up <- get_irf(1, prob, irflength, irfs1) %>% mutate(label = "FFR",
                                                        regime = "High uncertainty")
cpi_up <- get_irf(2, prob, irflength, irfs1) %>% mutate(label = "CPI",
                                                        regime = "High uncertainty")
indpro_up <- get_irf(3, prob, irflength, irfs1) %>% mutate(label = "INDPRO",
                                                           regime = "High uncertainty")

ffr_dn <- get_irf(4, prob, irflength, irfs2) %>% mutate(label = "FFR",
                                                        regime = "Low uncertainty")
cpi_dn <- get_irf(5, prob, irflength, irfs2) %>% mutate(label = "CPI",
                                                        regime = "Low uncertainty")
indpro_dn <- get_irf(6, prob, irflength, irfs2) %>% mutate(label = "INDPRO",
                                                           regime = "Low uncertainty")

bind_rows(ffr_up, cpi_up, indpro_up,
          ffr_dn, cpi_dn, indpro_dn) %>%
  mutate(label = factor(label, levels = c("FFR", "CPI", "INDPRO"))) %>%
  ggplot(aes(x = t, y = median, ymin = lower, ymax = upper)) +
  geom_line() +
  geom_hline(yintercept = 0, color = "red") +
  geom_ribbon(fill="grey", alpha=.2, color="grey50", linetype="dashed") + 
  facet_wrap(regime~label, scales = "free")+
  theme_minimal() +
  theme(plot.title = element_text(size = 11, hjust=0.5),
        axis.title.y = element_text(size=11)) +
  labs(x = "",
       y ="")
```

High uncertaintyben valamivel magasabb értékről indul a kamatsokk, picivel jobban fogja vissza az inflációt, és érzhetően erősebb az output visszaesése is.

# Három regime-es modell

Itt gyakorlatilag az előzőket ismétlem meg csak a regime meghatározás lesz más: High uncertainty a felső kvartilis felett, Low uncertainty az alsó kvartilis alatt, közte pedig Medium.

```{r}
#plot variables coloured by regime indicator
data %>%
  dplyr::select(date, ffr, cpi, indpro, epu, epu25, epu75) %>%
  mutate(indicator_up = ifelse(epu > epu75, 1,0),
         indicator_dn = ifelse(epu < epu25, 1, 0),
         indicator_mid = ifelse(indicator_up == 0 & indicator_dn == 0, 1, 0),
         indicator = case_when(indicator_up == 1 ~ "High uncertainty regime",
                               indicator_mid == 1 ~ "Medium uncertainty regime",
                               indicator_dn == 1 ~ "Low uncertainty regime"),
         indicator = factor(indicator, levels = c("High uncertainty regime",
                                                  "Medium uncertainty regime",
                                                  "Low uncertainty regime"))) %>%
  gather(key = "variable", value = "value", ffr, cpi, indpro) %>%
  mutate(variable = case_when(variable == "ffr" ~ "Federal Funds Rate",
                              variable == "cpi" ~ "Consumer Prices",
                              variable == "indpro" ~ "Industrial Production Index"),
         variable = factor(variable, levels = c("Federal Funds Rate", "Consumer Prices", "Industrial Production Index"))) %>%
  ggplot(aes(x = date, y = value, color = indicator, group = 1)) +
  geom_line(size = 1) +
  facet_wrap(~variable, scales = "free", nrow = 3) + theme_minimal() +
  labs(x = "",
       y = "") + 
  theme(legend.title = element_blank())

#fit 3 regime model
data_3regime <- data %>%
  dplyr::select(date, ffr, cpi, indpro, epu, epu25, epu75) %>%
  mutate(indicator_up = ifelse(epu > epu75, 1,0),
         indicator_dn = ifelse(epu < epu25, 1, 0),
         indicator_mid = ifelse(indicator_up == 0 & indicator_dn == 0, 1, 0),
         ffr_up = ffr*indicator_up,
         cpi_up = cpi*indicator_up,
         indpro_up = indpro*indicator_up,
         ffr_mid = ffr*indicator_mid,
         cpi_mid = cpi*indicator_mid,
         indpro_mid = indpro*indicator_mid,
         ffr_dn = ffr*indicator_dn,
         cpi_dn = cpi*indicator_dn,
         indpro_dn = indpro*indicator_dn) %>% 
  as_tibble() %>%
  dplyr::select(ffr_up, cpi_up, indpro_up, ffr_mid, cpi_mid, indpro_mid, ffr_dn, cpi_dn, indpro_dn) %>% ts()

set.seed(seed)
constr_up <- c(+1, -2, -3)
model_up <- uhlig.reject(Y=data_3regime, nlags=nlag, draws=200, subdraws=200, nkeep=1000, KMIN=1,
                         KMAX=6, constrained=constr_up, constant=TRUE, steps=21)
irfs1 <- model_up$IRFS

set.seed(seed)
constr_mid <- c(+4, -5, -6)
model_mid <- uhlig.reject(Y=data_3regime, nlags=nlag, draws=200, subdraws=200, nkeep=1000, KMIN=1,
                         KMAX=6, constrained=constr_mid, constant=TRUE, steps=21)
irfs2 <- model_mid$IRFS

set.seed(seed)
constr_dn <- c(+7, -8, -9)
model_dn <- uhlig.reject(Y=data_3regime, nlags=nlag, draws=200, subdraws=200, nkeep=1000, KMIN=1,
                         KMAX=6, constrained=constr_dn, constant=TRUE, steps=21)
irfs3 <- model_dn$IRFS

prob <- 0.16
irflength <- 21

ffr_up <- get_irf(1, prob, irflength, irfs1) %>% mutate(label = "FFR",
                                                        regime = "High uncertainty")
cpi_up <- get_irf(2, prob, irflength, irfs1) %>% mutate(label = "CPI",
                                                        regime = "High uncertainty")
indpro_up <- get_irf(3, prob, irflength, irfs1) %>% mutate(label = "INDPRO",
                                                           regime = "High uncertainty")

ffr_mid <- get_irf(4, prob, irflength, irfs2) %>% mutate(label = "FFR",
                                                        regime = "Medium uncertainty")
cpi_mid <- get_irf(5, prob, irflength, irfs2) %>% mutate(label = "CPI",
                                                        regime = "Medium uncertainty")
indpro_mid <- get_irf(6, prob, irflength, irfs2) %>% mutate(label = "INDPRO",
                                                           regime = "Medium uncertainty")

ffr_dn <- get_irf(7, prob, irflength, irfs3) %>% mutate(label = "FFR",
                                                        regime = "Low uncertainty")
cpi_dn <- get_irf(8, prob, irflength, irfs3) %>% mutate(label = "CPI",
                                                        regime = "Low uncertainty")
indpro_dn <- get_irf(9, prob, irflength, irfs3) %>% mutate(label = "INDPRO",
                                                           regime = "Low uncertainty")

bind_rows(ffr_up, cpi_up, indpro_up,
          ffr_mid, cpi_mid, indpro_mid,
          ffr_dn, cpi_dn, indpro_dn) %>%
  mutate(label = factor(label, levels = c("FFR", "CPI", "INDPRO")),
         regime = factor(regime, levels = c("High uncertainty", "Medium uncertainty", "Low uncertainty"))) %>%
  ggplot(aes(x = t, y = median, ymin = lower, ymax = upper)) +
  geom_line() +
  geom_hline(yintercept = 0, color = "red") +
  geom_ribbon(fill="grey", alpha=.2, color="grey50", linetype="dashed") + 
  facet_wrap(regime~label, scales = "free")+
  theme_minimal() +
  theme(plot.title = element_text(size = 11, hjust=0.5),
        axis.title.y = element_text(size=11)) +
  labs(x = "",
       y ="")
```

Nagyjából hasonló eredmény, high-ból a low felé haladva egyre kisebb a kamatsokk. A legalacsonyabb-ban még kicsit kevésbé perzisztens is. Az inflációt nagyjából hasonló mértékben fogja vissza a kamatemelés, a gazdasági visszaesés pedig a magas regime-ben érezhetően magasabb mint alatta.

Pár apróság, ami még megemlítendő:

 - Mindegyik modellből csak a saját regime-jét sokkolom, de az a másik regime-ekre is kihat. Ez szerintem nem probléma, ha pl. high regime-ben vagyok, ott a többi idősor értéke a konstrukció alapján 0-val van megszorozva - tehát elméletileg ott nem lenne impact. Teljesen még nem győztem meg magam ezzel kapcsolatban, de szerintem ez nem problémás.

 - Van még más algoritmus is a 'VarSignR' csomagban az Uhlig féle rejection és penalty-n kívül. Van még az Arias-Rubio-Wagoner féle rejection algoritmus (ők kritizálták egyébként az Uhlig féle penalty method hátrányait) - ezzel még értemes lehet kísérletezni - elvileg hatékonyabb, mint az Uhlig féle method. Illetve van még Fry-Pagan median targeting algoritmusa is - ezt még nem ismerem annyira.

 - Az MCMC szimulációk viszonylag gyorsak, így akár növelhetek a szimulált posterior-ok / impulzus válaszok számán, illetve a megtartott impulzus válaszok on is (hátha van ahol komolyabb szimulációval talál mégtöbb IRF-et ami megfelel az előjel megkötéseknek).

 - Lehet majd robustness check-et csinálni más indexekkel (VIX / TED spread / Financial stress indicator, stb), az akár itt, akár - ha értelmes eredmény jön ki - a local projection-ös módszerhez (amikor eljutok odáig) jó lehet.

### 2023-10-15 US/depo/uncertainty_transmission.Rmd
Initial commit

---
title: "The Asymmetric Effect of Uncertainty on Monetary Transmission"
author: "Péter Horváth"
output: 
  pdf_document:
    includes:
      in_header: "preamble.tex"
fontsize: 12pt
bibliography: "ref.bib"
link-citations: true
geometry: "a4paper,outer=25mm,inner=35mm,top=25mm,bottom=25mm"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      results = FALSE)
```

## Abstract

This paper discusses the asymmetric effect of uncertainty on the monetary transmission mechanism. The notion is that high uncertainty dampens the transmission of the interest rate channel, thus implying the importance of using forward guidance alongside interest rate decisions. To study this nonlinear relationship, I use a threshold VAR approach to estimate sign restricted VAR models. The results will show that at the more extreme levels of uncertainty there is considerable asymmetry in the transmission mechanism. Moreover in such instances monetary policy is less effectively able to anchor inflation and is only able to do so with a larger drop in economic activity.

\bigskip

\noindent \textbf{Keywords:} Uncertainty, Monetary transmission, Asymmetry, Sign restriction, Vector-Autoregression

\newpage

## Introduction

Over the recent years, uncertainty and asymmetries are two subjects that gained considerable traction in macroeconomics research. Both fields of research already have a rich literature, with there being a consensus on uncertainty shocks having an impact resembling that of a negative demand shock on economic activity (@CARRIERESWALLOW2013316, @COLOMBO201339, @CALDARA2016185, @CHENG2018305, @BONCIANI2020102236, @NILAVONGSE2020108765); and several papers have shown that macroeconomic shocks can have different impacts depending on the state of the economy.
@RePEc:eee:jmacro:v:69:y:2021:i:c:s0164070421000379, @RePEc:fip:fedker:00016, @RePEc:cup:macdyn:v:20:y:2016:i:05:p:1219-124600 and @RePEc:jae:japmet:v:19:y:2004:i:5:p:551-565 all take slightly different approaches, however all papers find that the impact of (negative) macroeconomic shocks is a sharper decline when uncertainty is high. Similar processes can be observed when carrying out the analysis with respect to the economy's position in the business cycle. @RePEc:knz:dpteco:1402 and @RePEc:een:camaaa:2020-72 show that uncertainty shocks have larger impact during economic contractions, and @RePEc:gro:rugsom:98c36 finds that monetary transmission is amplified during recessionary periods in the US and Germany. Another field of research on asymmetric dynamics is finding asymmetry in monetary policy-making itself. @Lin+2021+425+447 finds that macroeconomic aggregates react asymmetrically to positive and negative monetary policy shocks. @doi:10.1080/1331677X.2018.1481445 links uncertainty and asymmetries in monetary policy by assuming a difference in the policy reaction function in over the business cycle.

With this paper, I aim to contribute to the pre-existing literature, by linking the effect of uncertainty to the efficacy of monetary transmission. The results could have key implications for policy making, more specifically, it could highlight the importance of forward guidance measures, especially during times, when economic uncertainty is at its peak.

The rest of the paper will be outlined as follows: Section 2 gives a brief overview of the data used, Section 3 discusses the empirical strategy, in Section 4 I discuss the results and Section 5 concludes.

## Data

For the purposes of the analysis I will be estimating VAR models with three variables, the federal funds rate, consumer price index, and industrial production index. All data are monthly from ranging from January 1985 to December 2019 and all are retrieved from the federal reserve database.

```{r}
#loading packages
#packages
suppressPackageStartupMessages({
  library(readxl)
  library(tidyverse)
  library(tseries)
  library(forecast)
  library(TSA)
  library(fpp2)
  library(lmtest)
  library(vars)
  library(mFilter)
  library(ggplot2)
  library(tsDyn)
  library(fredr)
  library(VARsignR)
  library(lubridate)
  library(cowplot)
  library(gridExtra)
  library(patchwork)
  library(data.table)
  library(tsibble)
  library(mvnfast)
})

#set api key
fredr_set_key("cda47ae66b38ed7988c0a9c2ec80c94f")

#download data
params <- list(
  series_id = c("USEPUINDXD", "DFF", "USACPIALLMINMEI", "INDPRO"),
  frequency = "m",
  observation_start = as.Date("1950-01-01")
)

import  <- pmap_dfr(
  .l = params,
  .f = ~ fredr(series_id = .x, frequency = .y)
) %>%
  dplyr::select(date, series_id, value) %>%
  spread(key = series_id, value = value) %>%
  drop_na() %>% as_tsibble() %>% rename(epu = USEPUINDXD,
                                        ffr = DFF,
                                        cpi = USACPIALLMINMEI,
                                        indpro = INDPRO
                                        )

#select data range
data <- import %>%
  dplyr::select(date, epu, ffr, cpi, indpro) %>%
  drop_na() %>% filter(date <= as.Date("2019-12-01"))

```

```{r, fig.cap = "Time series plots of the Federal Funds Rate, Consumer Price Index and Industrial Production Index for the USA."}
#plotting the data series
data %>%
  gather(key = "variable", value = "value", ffr, cpi, indpro) %>%
  mutate(variable = case_when(variable == "ffr" ~ "Federal Funds Rate",
                              variable == "cpi" ~ "Consumer Prices",
                              variable == "indpro" ~ "Industrial Production Index"),
         variable = factor(variable, levels = c("Federal Funds Rate", "Consumer Prices", "Industrial Production Index"))) %>%
  ggplot(aes(x = date, y = value)) +
  geom_line() +
  facet_wrap(~variable, scales = "free", nrow = 3) + theme_minimal() +
  labs(x = "",
       y = "") 
```

As most macroeconomic variables, these data series are not stationary either, which is not desirable for the stability of the VAR models. For this reason I will be using the first difference of the industrial production index and the CPI. However, I will use the FFR at its level. This way I have better stability in my models and all variables have clear economic interpretation, as the two first-differenced data series can be considered as monthly growth rates of production and inflation respectively.

Another key variable I will be using as a "quasi transition variable" is a measure of uncertainty. This however is not a trivial metric, as several data series can be interpreted as a measurement of economic turmoil. The most commonly used are among others the VIX index and the TED spread.

```{r, fig.cap = "Time series plots of the VIX index, TED spread and the Economic Policy Uncertainty index, log scales."}
epu <- fredr(series_id = "USEPUINDXD",
             frequency = "m") %>%
  dplyr::select(date, value) %>%
  mutate(id = "EPU index",
         id = as.factor(id))

ted <- fredr(series_id = "TEDRATE",
             frequency = "m") %>%
  dplyr::select(date, value)%>%
  mutate(id = "TED spread",
         id = as.factor(id))

vix <- fredr(series_id = "VIXCLS",
             frequency = "m") %>%
  dplyr::select(date, value)%>%
  mutate(id = "VIX index",
         id = as.factor(id))

vis <- epu %>%
  bind_rows(ted) %>%
  bind_rows(vix) %>%
  ggplot(aes(x = date, y = value, group = id, color = id)) +
  geom_line() + 
  geom_smooth(method = "gam", se = FALSE, linetype = "dashed") +
  theme_classic() +
  theme(legend.title = element_blank()) +
  scale_y_log10() +
  labs(x = "",
       y = "")
vis
```

The former derives from the market expectations on the price changes of the S&P500 index, while the latter is the difference between the 3 month London interbank rate and the 3 month US treasury bill. Both are widely used in measuring financial stress or uncertainty, however in recent years in the literature on uncertainty, news based indices have gained a lot of traction. One such is the Economic Policy Uncertainty index created by @baker2016measuring .

```{r}
#generate first differences / growth rates  
data <- data %>%
  mutate(cpi = cpi - lag(cpi),
         indpro = indpro - lag(indpro)) %>% drop_na()

#creating threshold values
data <- data %>%
  mutate(epu50 = median(epu),
         epu75 = quantile(epu, probs = 0.75),
         epu25 = quantile(epu, probs = 0.25))
```

```{r, fig.cap = "Time series plot of the BBD Economic Policy Uncertainty Index. The red line indicates the median value, the green lines indicate the upper and lower quartiles."}

#plot epu index with threshold values
data %>% dplyr::select(date, epu, epu50, epu25, epu75) %>%
  ggplot(aes(x = date, y = epu)) +
  geom_line() +
  geom_hline(aes(yintercept = epu50), color = "darkred") +
  geom_hline(aes(yintercept = epu25), color = "darkgreen") +
  geom_hline(aes(yintercept = epu75), color = "darkgreen") +
  theme_minimal() +
  theme(plot.title = element_text(size = 11, hjust=0.5),
        axis.title.y = element_text(size=11)) +
  labs(x = "",
       y = "",
       title = "BBD economic policy uncertainty index")
```

\noindent The index is primarily news based, on uncertainty related keywords in influential papers in the US. The index also takes into account temporary tax measures reported by the Congressional Budget Office and the dispersion of individual forecasts from the Federal Reserve Bank of Philadelphia's Survey of Professional Forecasters in the consumer price index and government expenditure. The index is constructed by normalizing the sub-components by their respective standard deviation and the normalized series weighted averaged (with the news based component having the highest weight of 0.5). As this index 1) covers a broader spectrum of expectations than the financial market based uncertainty counterparts, and 2) is more specific to economic policy, I believe it should serve as the most adequate measurement of uncertainty for the purposes of this study.

## Empirical Strategy

An empirical challenge for this analysis was coercing the macroeconomic aggregates to behave the way it is written in all macroeconomics textbooks, i.e. to solve the price puzzle. 
There have already been numerous papers on the topic of solving the price puzzle (e.g.: @RePEc:eee:moneco:v:51:y:2004:i:7:p:1385-1413, @RePEc:hhs:hastef:0414, @RePEc:spr:empeco:v:46:y:2014:i:2:p:701-731, @RePEc:rba:rbardp:rdp2017-02, @RePEc:aea:aejmac:v:8:y:2016:i:4:p:75-102, @RePEc:aea:aecrev:v:94:y:2004:i:4:p:1055-1084), however with the available data series the best empirical approach would be to apply sign restrictions in the model specification. One of the most frequently used methods is the @RePEc:eee:moneco:v:52:y:2005:i:2:p:381-419 penalty algorithm. This method however has some "hidden features" as pointed out by @arias2018inference. Most importantly in this paper, using the penalty function approach would artificially narrow confidence intervals for the estimates, thus giving us a false sense of robustness in the results. For this reason, I will be implementing the full bayesian rejection algorithm of @rubio2010structural. The computation is done by using the R package created by @Danne2015.

I will introduce nonlinearity by using the BBD uncertainty index to create an indicator function. Taking the interactions of the indicator with each variable, then estimating the sign restricted models will give us a set of impulse responses for each regime which can be compared. First, I will experiment with a two regime model where the high uncertainty regime is considered when the uncertainty index is above its median. Next I will also estimate a three regime version, where the threshold values will be the upper and lower quartile values of the index. Thus, a generalized mathematical representation of the model would be as follows:
 \begin{equation}
 	Y_{t} = \sum\limits_{k=1}^{K}\Theta_{k}I(X_{t-1})Y_{t-1} + \epsilon_{t},
 \end{equation}
where $Y_{t}$ is the vector of endogenous variables, $\Theta_{k}$ is the matrix of coefficients in regime k, $X_{t}$ is the regime indicator function, and $\epsilon_{t}$ is the error term.
The ordering of the variables will be
\begin{equation}
	Y_{t} = 
	\begin{Bmatrix}
		FFR_{t} \\
		CPI_{t} \\
		INDPRO_{t}
	\end{Bmatrix}
\end{equation}
\noindent which is only important for syntactical purposes, as the R package is only able to carry out partial identifications,and in a vector of a given set of sign restrictions, the first one is considered as the shock. As for the identification scheme, I will use a full set of sign restrictions in each regime, and repeat the simulation of the shock for each regime.

With monthly data, ideally one would try to utilize as many lagged values in a VAR model as possible. However, due to the complexity of the task, the inclusion of numerous lags will have to be sacrificed. Through experimentation, I found that for a lag length of $n = 3$ yields an adequate number of accepted IRF draws (at least 100) when a sufficiently large large simulation is run - meaning 20000 MCMC replications and 200 draws per replication. A minor caveat to mention here is that the results are somewhat sensitive to the lag length of choice. With regards to higher lags this is not as troubling, as due to a lack of accepted draws, those results are less credible, however the initial impact on inflation and economic activity can vary noticeably if we reduce the number of included lags to one or two.

In order to compare and analyze results, I find the use of impulse responses to be the best tool. As the sign restriction algorithm gives us a set of impulse responses that satisfy the restrictions, I will be taking the median of the impulse responses and an 84% confidence interval in each case. I will normalize the shocks to a one percentage point interest rate rise, so that the IRF-s are more easily compared.

```{r}
#function to retrieve impulse responses
get_irf <- function(dim, prob, irflength, irfdata){
  prob <- prob
  x <- irfdata[ , ,dim] %>% t() 
  t <- matrix(nrow = nrow(x), ncol = 1) 
  med <- matrix(nrow = nrow(x), ncol = 1)
  up <- matrix(nrow = nrow(x), ncol = 1)
  dn <- matrix(nrow = nrow(x), ncol = 1)
  for(i in(1:irflength)){
    t[i] <- i-1
    med[i] <- median(x[i,])
    up[i] <- quantile(x[i,], probs = (1-prob))
    dn[i] <- quantile(x[i,], probs = prob)
  }
  irf <- bind_cols(t, med, up, dn) %>%
    rename(t = ...1,
           median = ...2,
           upper = ...3,
           lower = ...4)
  irf
}
```

## Results

First let us take a look at the data series colored by the regime indicator. In the plot below, we can see that while the indicator created from the uncertainty index (especially the two regime version) is somewhat similar to a recession indicator, however I do not believe the two are to be considered interchangeable.

```{r, fig.cap = "Time series plots of the endogenous variables colored by the regime indicators.", fig.height=7}

#plot variables coloured by regime indicator - 2
p2 <- data %>%
  dplyr::select(date, ffr, cpi, indpro, epu, epu50) %>%
  mutate(indicator = ifelse(epu > epu50, 1, 2),
         indicator = case_when(indicator == 1 ~ "High uncertainty regime",
                               indicator == 2 ~ "Low uncertainty regime")) %>%
  gather(key = "variable", value = "value", ffr, cpi, indpro) %>%
  mutate(variable = case_when(variable == "ffr" ~ "Federal Funds Rate",
                              variable == "cpi" ~ "Consumer Prices",
                              variable == "indpro" ~ "Industrial Production Index"),
         variable = factor(variable, levels = c("Federal Funds Rate", "Consumer Prices", "Industrial Production Index"))) %>%
  ggplot(aes(x = date, y = value, color = indicator, group = 1)) +
  geom_line(size = .8) +
  facet_wrap(~variable, scales = "free", nrow = 3) + theme_minimal() +
  labs(x = "",
       y = "") + 
  theme(legend.title = element_blank())

#plot variables coloured by regime indicator - 3
p3 <- data %>%
  dplyr::select(date, ffr, cpi, indpro, epu, epu25, epu75) %>%
  mutate(indicator_up = ifelse(epu > epu75, 1,0),
         indicator_dn = ifelse(epu < epu25, 1, 0),
         indicator_mid = ifelse(indicator_up == 0 & indicator_dn == 0, 1, 0),
         indicator = case_when(indicator_up == 1 ~ "High uncertainty regime",
                               indicator_mid == 1 ~ "Medium uncertainty regime",
                               indicator_dn == 1 ~ "Low uncertainty regime"),
         indicator = factor(indicator, levels = c("High uncertainty regime",
                                                  "Medium uncertainty regime",
                                                  "Low uncertainty regime"))) %>%
  gather(key = "variable", value = "value", ffr, cpi, indpro) %>%
  mutate(variable = case_when(variable == "ffr" ~ "Federal Funds Rate",
                              variable == "cpi" ~ "Consumer Prices",
                              variable == "indpro" ~ "Industrial Production Index"),
         variable = factor(variable, levels = c("Federal Funds Rate", "Consumer Prices", "Industrial Production Index"))) %>%
  ggplot(aes(x = date, y = value, color = indicator, group = 1)) +
  geom_line(size = .8) +
  facet_wrap(~variable, scales = "free", nrow = 3) + theme_minimal() +
  labs(x = "",
       y = "") + 
  theme(legend.title = element_blank())

grid.arrange(p2, p3)

```

Next, taking a look at the impulse responses of the two regime model, we can see mostly a lack of evidence for nonlinearity. The simulated path of the FFR shock and the response of inflation is rather similar in both regimes. The major difference in the two regimes is in the response of economic activity, as in the low uncertainty case the drop in output is much less persistent compared to the high regime. This could be explained by low uncertainty periods generally being characterized by stable GDP growth as seen in the previous graphs.

```{r, fig.cap = "Impulse responses from two regime model.",             fig.height = 4.5}
#some parameters for estimation and irfs
nlag <- 3
seed <- 2022
prob <- 0.16
irflength <- 21
draw <- 20000
subdraw <- 200

# fit 2 regime model
data_2regime <- data %>%
  dplyr::select(date, ffr, cpi, indpro, epu, epu50) %>%
  mutate(indicator = ifelse(epu > epu50, 1, 0),
         indicator_inv = ifelse(indicator == 1, 0, 1),
         ffr_up = ffr*indicator,
         cpi_up = cpi*indicator,
         indpro_up = indpro*indicator,
         ffr_dn = ffr*indicator_inv,
         cpi_dn = cpi*indicator_inv,
         indpro_dn = indpro*indicator_inv) %>% as_tibble() %>%
  dplyr::select(ffr_up, cpi_up, indpro_up, ffr_dn, cpi_dn, indpro_dn) %>% ts()

set.seed(seed)
constr_up <- c(+1, -2, -3)
model_up <- rwz.reject(Y=data_2regime, nlags=nlag, draws=draw, subdraws=subdraw, nkeep=1000, KMIN=1,
                         KMAX=6, constrained=constr_up, constant=TRUE, steps=21)
irfs1 <- model_up$IRFS

set.seed(seed)
constr_dn <- c(+4, -5, -6)
model_dn <- rwz.reject(Y=data_2regime, nlags=nlag, draws=draw, subdraws=subdraw, nkeep=1000, KMIN=1,
                       KMAX=6, constrained=constr_dn, constant=TRUE, steps=21)
irfs2 <- model_dn$IRFS

ffr_up <- get_irf(1, prob, irflength, irfs1) %>% mutate(label = "FFR",
                                                        regime = "High uncertainty")
cpi_up <- get_irf(2, prob, irflength, irfs1) %>% mutate(label = "CPI",
                                                        regime = "High uncertainty")
indpro_up <- get_irf(3, prob, irflength, irfs1) %>% mutate(label = "INDPRO",
                                                           regime = "High uncertainty")

ffr_dn <- get_irf(4, prob, irflength, irfs2) %>% mutate(label = "FFR",
                                                        regime = "Low uncertainty")
cpi_dn <- get_irf(5, prob, irflength, irfs2) %>% mutate(label = "CPI",
                                                        regime = "Low uncertainty")
indpro_dn <- get_irf(6, prob, irflength, irfs2) %>% mutate(label = "INDPRO",
                                                           regime = "Low uncertainty")

bind_rows(ffr_up, cpi_up, indpro_up,
          ffr_dn, cpi_dn, indpro_dn) %>%
  mutate(label = factor(label, levels = c("FFR", "CPI", "INDPRO")),
         regime = factor(regime, levels = c("High uncertainty", "Low uncertainty"))) %>% 
  left_join(bind_rows(ffr_up, cpi_up, indpro_up,
                      ffr_dn, cpi_dn, indpro_dn) %>%
              mutate(label = factor(label, levels = c("FFR", "CPI", "INDPRO")),
                     regime = factor(regime, levels = c("High uncertainty", "Low uncertainty"))) %>% 
              filter(t == 0) %>% 
              mutate(mult = 1/median) %>% 
              mutate(mult = ifelse(mult < 0, NA, mult)) %>% 
              group_by(regime) %>% 
              mutate(mult = min(mult, na.rm = TRUE)) %>% 
              ungroup() %>% 
              mutate(median = median * mult,
                     upper = upper * mult,
                     lower = lower * mult) %>% 
              dplyr::select(label, regime, mult), 
            by = c("regime", 'label')) %>% 
  mutate(median = median * mult,
         upper = upper * mult, 
         lower = lower * mult) %>% 
  ggplot(aes(x = t, y = median, ymin = lower, ymax = upper)) +
  geom_line() +
  geom_hline(yintercept = 0, color = "red") +
  geom_ribbon(fill="grey", alpha=.2, color="grey50", linetype="dashed") + 
  facet_wrap(regime~label, scales = "free")+
  theme_minimal() +
  theme(plot.title = element_text(size = 11, hjust=0.5),
        axis.title.y = element_text(size=11)) +
  labs(x = "",
       y ="")
```

With the results of the two regime model in mind, I believe there is room for further developing the model to investigate the relationship between uncertainty and monetary policy. A problem of the two regime model could be that the transmission is very similar when uncertainty is just above or just below the median - i.e. not being able to grasp the two extreme sides. Thus, a three regime variant where the threshold points are at the upper and lower quartiles could yield us a clearer picture. The impulse responses of the three regime model can be seen below:

```{r, fig.cap = "Impulse responses from three regime model.",          fig.height = 6}
#fit 3 regime model
data_3regime <- data %>%
  dplyr::select(date, ffr, cpi, indpro, epu, epu25, epu75) %>%
  mutate(indicator_up = ifelse(epu > epu75, 1,0),
         indicator_dn = ifelse(epu < epu25, 1, 0),
         indicator_mid = ifelse(indicator_up == 0 & indicator_dn == 0, 1, 0),
         ffr_up = ffr*indicator_up,
         cpi_up = cpi*indicator_up,
         indpro_up = indpro*indicator_up,
         ffr_mid = ffr*indicator_mid,
         cpi_mid = cpi*indicator_mid,
         indpro_mid = indpro*indicator_mid,
         ffr_dn = ffr*indicator_dn,
         cpi_dn = cpi*indicator_dn,
         indpro_dn = indpro*indicator_dn) %>% 
  as_tibble() %>%
  dplyr::select(ffr_up, cpi_up, indpro_up, ffr_mid, cpi_mid, indpro_mid, ffr_dn, cpi_dn, indpro_dn) %>% ts()

set.seed(seed)
constr_up <- c(+1, -2, -3)
model_up <- rwz.reject(Y=data_3regime, nlags=nlag, draws=draw, subdraws=subdraw, nkeep=1000, KMIN=1,
                         KMAX=6, constrained=constr_up, constant=TRUE, steps=21)
irfs1 <- model_up$IRFS

set.seed(seed)
constr_mid <- c(+4, -5, -6)
model_mid <- rwz.reject(Y=data_3regime, nlags=nlag, draws=draw, subdraws=subdraw, nkeep=1000, KMIN=1,
                         KMAX=6, constrained=constr_mid, constant=TRUE, steps=21)
irfs2 <- model_mid$IRFS

set.seed(seed)
constr_dn <- c(+7, -8, -9)
model_dn <- rwz.reject(Y=data_3regime, nlags=nlag, draws=draw, subdraws=subdraw, nkeep=1000, KMIN=1,
                         KMAX=6, constrained=constr_dn, constant=TRUE, steps=21)
irfs3 <- model_dn$IRFS

prob <- 0.16
irflength <- 21

ffr_up <- get_irf(1, prob, irflength, irfs1) %>% mutate(label = "FFR",
                                                        regime = "High uncertainty")
cpi_up <- get_irf(2, prob, irflength, irfs1) %>% mutate(label = "CPI",
                                                        regime = "High uncertainty")
indpro_up <- get_irf(3, prob, irflength, irfs1) %>% mutate(label = "INDPRO",
                                                           regime = "High uncertainty")

ffr_mid <- get_irf(4, prob, irflength, irfs2) %>% mutate(label = "FFR",
                                                        regime = "Medium uncertainty")
cpi_mid <- get_irf(5, prob, irflength, irfs2) %>% mutate(label = "CPI",
                                                        regime = "Medium uncertainty")
indpro_mid <- get_irf(6, prob, irflength, irfs2) %>% mutate(label = "INDPRO",
                                                           regime = "Medium uncertainty")

ffr_dn <- get_irf(7, prob, irflength, irfs3) %>% mutate(label = "FFR",
                                                        regime = "Low uncertainty")
cpi_dn <- get_irf(8, prob, irflength, irfs3) %>% mutate(label = "CPI",
                                                        regime = "Low uncertainty")
indpro_dn <- get_irf(9, prob, irflength, irfs3) %>% mutate(label = "INDPRO",
                                                           regime = "Low uncertainty")

bind_rows(ffr_up, cpi_up, indpro_up,
          ffr_mid, cpi_mid, indpro_mid,
          ffr_dn, cpi_dn, indpro_dn) %>%
  mutate(label = factor(label, levels = c("FFR", "CPI", "INDPRO")),
         regime = factor(regime, levels = c("High uncertainty", "Medium uncertainty", "Low uncertainty"))) %>% 
  left_join(bind_rows(ffr_up, cpi_up, indpro_up,
                      ffr_mid, cpi_mid, indpro_mid,
                      ffr_dn, cpi_dn, indpro_dn) %>%
              mutate(label = factor(label, levels = c("FFR", "CPI", "INDPRO")),
                     regime = factor(regime, levels = c("High uncertainty", "Medium uncertainty", "Low uncertainty"))) %>% 
              filter(t == 0) %>% 
              mutate(mult = 1/median) %>% 
              mutate(mult = ifelse(mult < 0, NA, mult)) %>% 
              group_by(regime) %>% 
              mutate(mult = min(mult, na.rm = TRUE)) %>% 
              ungroup() %>% 
              mutate(median = median * mult,
                     upper = upper * mult,
                     lower = lower * mult) %>% 
              dplyr::select(label, regime, mult), 
            by = c("regime", 'label')) %>% 
  mutate(median = median * mult,
         upper = upper * mult, 
         lower = lower * mult) %>% 
  ggplot(aes(x = t, y = median, ymin = lower, ymax = upper)) +
  geom_line() +
  geom_hline(yintercept = 0, color = "red") +
  geom_ribbon(fill="grey", alpha=.2, color="grey50", linetype="dashed") + 
  facet_wrap(regime~label, scales = "free")+
  theme_minimal() +
  theme(plot.title = element_text(size = 11, hjust=0.5),
        axis.title.y = element_text(size=11)) +
  labs(x = "",
       y ="")

```

As a contrast to the two regime version, here we can see some sharper differences, hinting at the existence of the nonlinear relationship. In terms of the persistence of the FFR shock we see close to no differences, however differences in the responses of inflation and economic activity between the regimes is quite curious. The first thing we can notice is the persistence of the shock especially on output is considerably stronger when uncertainty is high compared to the other two regimes. Rising the interest rate is also considerably less effective at anchoring inflation as in the high uncertainty regime the impact nearly halves compared to the other two regimes. The drop in output is similar in all cases, however in the high uncertainty regime it is mildly amplified - hinting at a moderately larger trade-off in output drop versus inflation anchoring capability. These results suggest that interest rate decisions alone might not be solely sufficient at effectively reducing inflation, thus implying the importance of utilizing additional measures such as forward guidance in order to moderate economic uncertainty. An important caveat here is that the BBD Economic Policy uncertainty index by construction would likely be higher around the time of important policy actions such as interst rate rises. This could imply that the true impact of an interest rate decision is closer to what we see from the impulse responses of the high uncertainty regime as opposed to the other two set of impulse responses.

## Conclusion

This paper attempts at quantifying the impact of uncertainty on the monetary transmission mechanism. To study this asymmetric relationship I fit a two and three regime VAR model with sign restrictions using an indicator function created from the BBD Economic Policy Uncertainty Index. While in the two regime case there seems to be close to no asymmetry in the interest rate channel, in the more extreme sides of the distribution we can observe that high levels of uncertainty above its upper quartile seems to dampen the inflation anchoring capability of interest rate shocks at a mildly larger output drop trade-off. This would imply two conclusions: 1) as the index measuring uncertainty tracks news coverage - taking higher values around important policy decisions - it is likely the true impact of the interest rate channel is quantitatively more similar to the impact observed in high uncertainty times; and 2) interest rate decisions alone are less effective at anchoring inflation - pointing to the neccessity of combining it with alternative measures like forward guidance. Quantifying the impact of such measures is however beyond the scopes of this paper.

## References
<div id="ref"></div>

### 2023-10-15 US/depo/uncertainty_transmission_clean.Rmd
Initial commit

---
title: "The Impact of Uncertainty on Monetary Transmission"
author: "Péter Horváth"
output: 
  pdf_document:
    includes:
      in_header: "preamble.tex"
fontsize: 12pt
bibliography: "ref.bib"
csl: "custom-citations.csl"
link-citations: true
linkcolor: "blue"
geometry: "a4paper,outer=25mm,inner=35mm,top=25mm,bottom=25mm"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      results = FALSE)
```

## Abstract

This paper discusses the nonlinear relationship between uncertainty on the monetary transmission mechanism. The notion is that high uncertainty dampens the transmission of the interest rate channel, thus implying the importance of using forward guidance alongside interest rate decisions. To study this nonlinear relationship, I use a threshold VAR approach to estimate sign restricted VAR models. The results show that at more extreme levels of uncertainty there are considerable differences in the transmission mechanism. Moreover, in such instances monetary policy is able to anchor inflation less effectively and is only able to do so with a larger drop in economic activity.

\bigskip

\noindent \textbf{Keywords:} Uncertainty, Monetary transmission, Non-linearity, Sign restriction, Vector-Autoregression

\newpage

## Introduction

Over the recent years, uncertainty and asymmetries are two subjects that gained considerable traction in macroeconomics research. Both fields of research already have a rich literature, with there being a consensus on uncertainty shocks having an impact resembling that of a negative demand shock on economic activity (@CARRIERESWALLOW2013316, @COLOMBO201339, @CALDARA2016185, @CHENG2018305, @BONCIANI2020102236, @NILAVONGSE2020108765); and several papers have shown that macroeconomic shocks can have different impacts depending on the state of the economy.
@RePEc:eee:jmacro:v:69:y:2021:i:c:s0164070421000379, @RePEc:fip:fedker:00016, @RePEc:cup:macdyn:v:20:y:2016:i:05:p:1219-124600 and @RePEc:jae:japmet:v:19:y:2004:i:5:p:551-565 all take slightly different approaches, however all papers find that macroeconomic shocks lead to a sharper decline when uncertainty is high. Similar processes can be observed when carrying out the analysis with respect to the economy's position in the business cycle. @RePEc:knz:dpteco:1402 and @RePEc:een:camaaa:2020-72 show that uncertainty shocks have larger impact during economic contractions, and @RePEc:gro:rugsom:98c36 finds that monetary transmission is amplified during recessionary periods in the US and Germany. Another field of research on asymmetric dynamics is finding asymmetry in monetary policy-making itself. @Lin+2021+425+447 finds that macroeconomic aggregates react asymmetrically to positive and negative monetary policy shocks. @doi:10.1080/1331677X.2018.1481445 links uncertainty and asymmetries in monetary policy by assuming a difference in the policy reaction function over the business cycle.

With this paper, I aim to contribute to the pre-existing literature, by linking the effect of uncertainty to the efficacy of monetary transmission. The results could have key implications for policy making, more specifically, it could highlight the importance of forward guidance measures, especially during times, when economic uncertainty is at its peak.

The rest of the paper will be outlined as follows: Section 2 gives a brief overview of the data used, Section 3 discusses the empirical strategy, in Section 4 I discuss the results, in Section 5 I check the results' robustness to alternate specifications and Section 6 concludes.

## Data

For the purposes of the analysis I will be estimating VAR models with three variables, the federal funds rate, consumer price index, and industrial production index. All data are monthly from ranging from January 1985 to December 2019 and all are retrieved from the federal reserve database.

```{r}
#loading packages
#packages
library(fredr)
library(tidyverse)
library(gridExtra)
library(fredr)
library(VARsignR)
library(rlang)
library(lubridate)

#set api key
fredr_set_key("cda47ae66b38ed7988c0a9c2ec80c94f")

#download data
params <- list(
  series_id = c("USEPUINDXD", "DFF", "USACPIALLMINMEI", "INDPRO"),
  frequency = "m",
  observation_start = as.Date("1950-01-01")
)

import  <- pmap_dfr(
  .l = params,
  .f = ~ fredr(series_id = .x, frequency = .y)
) %>%
  select(date, series_id, value) %>%
  spread(key = series_id, value = value) %>%
  drop_na() %>% rename(epu = USEPUINDXD,
                       ffr = DFF,
                       cpi = USACPIALLMINMEI,
                       indpro = INDPRO
  )

#select data range
data <- import %>%
  select(date, epu, ffr, cpi, indpro) %>%
  drop_na() %>% filter(date <= as.Date("2019-12-01"))
```

```{r, fig.cap = "Time series plots of the Federal Funds Rate, Consumer Price Index and Industrial Production Index for the USA."}
#plotting the data series
data %>%
  gather(key = "variable", value = "value", ffr, cpi, indpro) %>%
  mutate(variable = case_when(variable == "ffr" ~ "Federal Funds Rate",
                              variable == "cpi" ~ "Consumer Prices",
                              variable == "indpro" ~ "Industrial Production Index"),
         variable = factor(variable, levels = c("Federal Funds Rate", "Consumer Prices", "Industrial Production Index"))) %>%
  ggplot(aes(x = date, y = value)) +
  geom_line() +
  facet_wrap(~variable, scales = "free", nrow = 3) + theme_minimal() +
  labs(x = "",
       y = "") 
```

As most macroeconomic variables, these data series are not stationary either, which is not desirable for the stability of the VAR models. For this reason I will be using the first difference of the industrial production index and the CPI. However, I will use the FFR at its level. This way I have better stability in my models and all variables have clear economic interpretation, as the two first-differenced data series can be considered as monthly growth rates of production and inflation respectively.

Another key variable I will be using as a "quasi transition variable" is a measure of uncertainty. This however is not a trivial metric, as several data series can be interpreted as a measurement of economic turmoil. Commonly used indices to measure uncertainty are are the VIX index and the TED spread.

```{r, fig.cap = "Time series plots of the VIX index, TED spread and the Economic Policy Uncertainty index, log scales."}
unc_measures <- bind_rows(
  fredr(series_id = "USEPUINDXD",
        frequency = "m") %>%
    dplyr::select(date, value) %>%
    mutate(id = "EPU index",
           id = as.factor(id)),
  fredr(series_id = "TEDRATE",
        frequency = "m") %>%
    dplyr::select(date, value)%>%
    mutate(id = "TED spread",
           id = as.factor(id)),
  fredr(series_id = "VIXCLS",
        frequency = "m") %>%
    dplyr::select(date, value)%>%
    mutate(id = "VIX index",
           id = as.factor(id))
) %>% filter(date <= as.Date("2019-12-01"))

#Plot uncertainty measures
unc_measures %>% 
  ggplot(aes(x = date, y = value, group = id, color = factor(id))) +
  geom_line() + 
  geom_smooth(method = "gam", se = FALSE, linetype = "dashed") +
  scale_y_log10() +
  scale_color_viridis_d(begin = 0.05, end = 0.4, name = "Index", option = "turbo") +
  theme_minimal() +
  labs(x = "",
       y = "")
```

The former derives from the market expectations on the price changes of the S&P500 index, while the latter is the difference between the 3 month London interbank rate and the 3 month US treasury bill. Both are widely used in measuring financial stress or uncertainty, however in recent years in the literature on uncertainty, news based indices have gained a lot of traction. One such is the Economic Policy Uncertainty index created by @baker2016measuring .

```{r}
#generate first differences / growth rates  
data <- data %>%
  mutate(cpi = cpi - lag(cpi),
         indpro = indpro - lag(indpro)) %>% drop_na()

#creating threshold values
data <- data %>%
  mutate(epu50 = median(epu),
         epu75 = quantile(epu, probs = 0.75),
         epu25 = quantile(epu, probs = 0.25))
```

```{r, fig.cap = "Time series plot of the Economic Policy Uncertainty Index. The red line indicates the median value, the blue lines indicate the upper and lower quartiles."}

#plot epu index with threshold values
data %>% 
  select(date, starts_with("epu")) %>% 
  gather(key = "variable", value = "value", starts_with("epu")) %>% 
  mutate(variable = case_when(variable == "epu" ~ "EPU",
                              variable == "epu50" ~ "Median",
                              variable == "epu25" ~ "Lower quartile",
                              variable == "epu75" ~ "Upper quartile")) %>% 
  ggplot(aes(x = date, y = value, group = variable, size = variable, color = variable)) +
  geom_line() +
  scale_size_manual(breaks = c("EPU", "Median", "Lower quartile", "Upper quartile"),
                    values = c(.5, 1, 1, 1)) +
  scale_color_manual(breaks = c("EPU", "Median", "Lower quartile", "Upper quartile"),
                     values = c("black", "#df2c14", "#0457ac", "#0457ac")) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(x = "",
       y = "")
```

\noindent The index is primarily news based, on uncertainty related keywords in influential newspapers in the US. The index also takes into account temporary tax measures reported by the Congressional Budget Office and the dispersion of individual forecasts from the Federal Reserve Bank of Philadelphia's Survey of Professional Forecasters in the consumer price index and government expenditure. The index is constructed by normalizing the sub-components by their respective standard deviation and then these values are averaged out. Out of the sub-components, the news based part has the highest weight of 0.5 in the index.

As this index 1) covers a broader spectrum of expectations than the financial market based uncertainty counterparts, and 2) is more specific to economic policy, I believe it should serve as the most adequate measurement of uncertainty for the purposes of this study.

```{r}
unc_measures <- unc_measures %>% 
  mutate(id = case_when(id == "EPU index" ~ "epu",
                        id == "TED spread" ~ "ted",
                        id == "VIX index" ~ "vix")) %>% 
  pivot_wider(names_from = id, values_from = value) %>% 
  drop_na()

cortab_create <- function(data, digits){
  cortab <- data %>% select(-date) %>% drop_na() %>% cor()
  cortab <- round(cortab, digits = digits)
  cortab <- cortab*lower.tri(cortab)
  diag(cortab) <- 1
  for(i in 1:nrow(cortab)) {
    for(j in 1:ncol(cortab)){
      if(cortab[i,j] == 0) {
        cortab[i,j] <- " "
      }
    }
  }
  cortab
  rownames(cortab) <- c("EPU", "TED", "VIX")
  colnames(cortab) <- c("EPU", "TED", "VIX")
  cortab
}
```

```{r}

cortab <- cortab_create(unc_measures, 3) %>% 
  stargazer::stargazer(style  = "aer", 
                       title = "Correlation table of uncertainty measures",
                       type = "latex") 
```

```{r, results='asis'}
cortab[4:length(cortab)] %>% cat()
```

We can see from the correlation table, the EPU index has low to moderate correlation with the other indices. This raises the question whether the results would be robust to using one of the other indices in place of the EPU, which is addressed in the robustness check.

## Empirical Strategy

An empirical challenge for this analysis was coercing the macroeconomic aggregates to behave the way it is written in all macroeconomics textbooks, i.e. to solve the price puzzle. 
There have already been numerous papers on the topic of solving the price puzzle (e.g.: @RePEc:eee:moneco:v:51:y:2004:i:7:p:1385-1413, @RePEc:hhs:hastef:0414, @RePEc:spr:empeco:v:46:y:2014:i:2:p:701-731, @RePEc:rba:rbardp:rdp2017-02, @RePEc:aea:aejmac:v:8:y:2016:i:4:p:75-102, @RePEc:aea:aecrev:v:94:y:2004:i:4:p:1055-1084), however with the available data series the best empirical approach would be to apply sign restrictions in the model specification. One of the most frequently used methods is the @RePEc:eee:moneco:v:52:y:2005:i:2:p:381-419 penalty algorithm. This method however has some "hidden features" as pointed out by @arias2018inference. Most importantly in this paper, using the penalty function approach would artificially narrow confidence intervals for the estimates, thus giving us a false sense of robustness in the results. For this reason, I will be implementing the full bayesian rejection algorithm of @rubio2010structural. The computation is done by using the R package created by @Danne2015.

I will introduce non-linearity by using the EPU index to create an indicator function. Taking the interactions of the indicator with each variable, then estimating the sign restricted models will give us a set of impulse responses for each regime which can be compared. First, I will experiment with a two-regime model where the high uncertainty regime is considered when the uncertainty index is above its median. Next I will also estimate a three-regime version, where the threshold values will be the upper and lower quartile values of the index. Thus, a generalized mathematical representation of the model would be as follows:
 \begin{equation}
 	Y_{t} = \sum\limits_{m=1}^{M}\sum\limits_{k=1}^{K}\Theta_{m}I(X_{t})Y_{t-k} + \epsilon_{t},
 \end{equation}
where $Y_{t}$ is the vector of endogenous variables, $\Theta_{m}$ is the matrix of coefficients in regime k, $X_{t}$ is the regime indicator function, and $\epsilon_{t}$ is the error term.
The ordering of the variables will be
\begin{equation}
	Y_{t} = 
	\begin{Bmatrix}
		FFR_{t} \\
		CPI_{t} \\
		INDPRO_{t}
	\end{Bmatrix}
\end{equation}
\noindent which is only important for syntactical purposes, as the R package is only able to carry out partial identifications,and in a vector of a given set of sign restrictions, the first one is considered as the shock. As for the identification scheme, I will use a full set of sign restrictions in each regime, and repeat the simulation of the shock for each regime.

With monthly data, ideally one would try to utilize as many lagged values in a VAR model as possible. However, due to the complexity of the task, the inclusion of numerous lags will have to be sacrificed. Through experimentation, I found that a lag length of $n = 3$ yields an adequate number of accepted IRF draws when a sufficiently large large simulation is run - meaning 20000 MCMC replications and 200 draws per replication. A minor caveat to mention here is that the results are somewhat sensitive to the lag length of choice. This issue is further discussed in the robustness check.

In order to compare and analyze results, I find the use of impulse responses to be the best tool. As the sign restriction algorithm gives us a set of impulse responses that satisfy the restrictions, I will be taking the median of the impulse responses and an 84% confidence interval in each case. I will normalize the shocks to a one percentage point interest rate rise, so that the IRF-s are more easily compared.

```{r}
#function to retrieve impulse responses
get_irf <- function(dim, prob, irflength, irfdata){
  prob <- prob
  x <- irfdata[ , ,dim] %>% t() 
  t <- matrix(nrow = nrow(x), ncol = 1) 
  med <- matrix(nrow = nrow(x), ncol = 1)
  up <- matrix(nrow = nrow(x), ncol = 1)
  dn <- matrix(nrow = nrow(x), ncol = 1)
  for(i in(1:irflength)){
    t[i] <- i-1
    med[i] <- median(x[i,])
    up[i] <- quantile(x[i,], probs = (1-prob))
    dn[i] <- quantile(x[i,], probs = prob)
  }
  irf <- bind_cols(t, med, up, dn) %>%
    rename(t = ...1,
           median = ...2,
           upper = ...3,
           lower = ...4)
  irf
}
```

## Results

First let us take a look at the data series colored by the regime indicator. A question that could be raised is whether a regime indicator created from such an index is different enough from a recession indicator. However as we can see from Figures 4 and 5, neither the two or the three-regime variants can be considered interchangeable with a recession indicator.

```{r, fig.cap = "Time series plots of the endogenous variables colored by the regime indicator (2 regimes)."}

#variables by regime indicator
#####
#plot variables coloured by regime indicator - 2
data %>%
  select(date, ffr, cpi, indpro, epu, epu50) %>%
  mutate(indicator = ifelse(epu > epu50, 1, 2),
         indicator = case_when(indicator == 1 ~ "High uncertainty regime",
                               indicator == 2 ~ "Low uncertainty regime")) %>%
  gather(key = "variable", value = "value", ffr, cpi, indpro) %>%
  mutate(variable = case_when(variable == "ffr" ~ "Federal Funds Rate",
                              variable == "cpi" ~ "Consumer Prices",
                              variable == "indpro" ~ "Industrial Production Index"),
         variable = factor(variable, levels = c("Federal Funds Rate", "Consumer Prices", "Industrial Production Index"))) %>%
  ggplot(aes(x = date, y = value, color = indicator, group = 1)) +
  geom_line(size = .8) +
  facet_wrap(~variable, scales = "free", nrow = 3) + theme_minimal() +
  scale_color_manual(breaks = c("High uncertainty regime", "Low uncertainty regime"),
                     values = c("#E9002D", "#00B000")) +
  labs(x = "",
       y = "",
       caption = "Shaded areas indicate NBER recession preiods") + 
  theme(legend.title = element_blank(),
        legend.position = "bottom") +
  geom_rect(data = fredr(series_id = "USREC",
                         frequency = "m") %>% 
              select(date, recession = value) %>% 
              mutate(diff = recession - lag(recession)) %>% 
              filter(!is.na(diff)) %>% 
              mutate(recession_start = ifelse(diff == 1, as.character(date), NA),
                     recession_end = ifelse(diff == -1, as.character(date), NA)) %>% 
              filter(!is.na(recession_start) | !is.na(recession_end)) %>% 
              mutate(recession_end = ifelse(!is.na(recession_start), lead(recession_end), NA)) %>%
              filter(!is.na(recession_start)) %>% 
              mutate(across(.cols = c(recession_start, recession_end), .fns = ~as.Date(.x))) %>% 
              select(recession_start, recession_end) %>% 
              filter(recession_start >= min(data$date),
                     recession_start <= max(data$date)),
            inherit.aes = F,
            aes(xmin = recession_start, xmax = recession_end, ymin = -Inf, ymax = Inf), 
            fill = "grey50", alpha = 0.5)

```

```{r, fig.cap = "Time series plots of the endogenous variables colored by the regime indicator (3 regimes)."}
#plot variables coloured by regime indicator - 3
data %>%
  select(date, ffr, cpi, indpro, epu, epu25, epu75) %>%
  mutate(indicator_up = ifelse(epu > epu75, 1,0),
         indicator_dn = ifelse(epu < epu25, 1, 0),
         indicator_mid = ifelse(indicator_up == 0 & indicator_dn == 0, 1, 0),
         indicator = case_when(indicator_up == 1 ~ "High uncertainty regime",
                               indicator_mid == 1 ~ "Medium uncertainty regime",
                               indicator_dn == 1 ~ "Low uncertainty regime"),
         indicator = factor(indicator, levels = c("High uncertainty regime",
                                                  "Medium uncertainty regime",
                                                  "Low uncertainty regime"))) %>%
  gather(key = "variable", value = "value", ffr, cpi, indpro) %>%
  mutate(variable = case_when(variable == "ffr" ~ "Federal Funds Rate",
                              variable == "cpi" ~ "Consumer Prices",
                              variable == "indpro" ~ "Industrial Production Index"),
         variable = factor(variable, levels = c("Federal Funds Rate", "Consumer Prices", "Industrial Production Index"))) %>%
  ggplot(aes(x = date, y = value, color = indicator, group = 1)) +
  geom_line(size = .8) +
  facet_wrap(~variable, scales = "free", nrow = 3) + theme_minimal() +
  scale_color_manual(breaks = c("High uncertainty regime", "Medium uncertainty regime", "Low uncertainty regime"),
                     values = c("#E9002D", "#9EC9E2" , "#00B000")) +
  labs(x = "",
       y = "",
       caption = "Shaded areas indicate NBER recession preiods") + 
  theme(legend.title = element_blank(),
        legend.position = "bottom") +
  geom_rect(data = fredr(series_id = "USREC",
                         frequency = "m") %>% 
              select(date, recession = value) %>% 
              mutate(diff = recession - lag(recession)) %>% 
              filter(!is.na(diff)) %>% 
              mutate(recession_start = ifelse(diff == 1, as.character(date), NA),
                     recession_end = ifelse(diff == -1, as.character(date), NA)) %>% 
              filter(!is.na(recession_start) | !is.na(recession_end)) %>% 
              mutate(recession_end = ifelse(!is.na(recession_start), lead(recession_end), NA)) %>%
              filter(!is.na(recession_start)) %>% 
              mutate(across(.cols = c(recession_start, recession_end), .fns = ~as.Date(.x))) %>% 
              select(recession_start, recession_end) %>% 
              filter(recession_start >= min(data$date),
                     recession_start <= max(data$date)),
            inherit.aes = F,
            aes(xmin = recession_start, xmax = recession_end, ymin = -Inf, ymax = Inf), 
            fill = "grey50", alpha = 0.5)
```

Next, taking a look at the impulse responses of the two-regime model in Figure 5, we see a lack of strong evidence for non-linearity. The simulated path of the FFR shock and the response of inflation is rather similar in both regimes. The only noticeable difference is in the response in economic activity, as in the low uncertainty case the drop in output is much less persistent compared to the high-regime. This could be explained by low uncertainty periods generally being characterized by stable GDP growth as seen in the previous graphs.

```{r, fig.cap = "Impulse responses from two-regime model.",             fig.height = 4.5}
#some parameters for estimation and irfs
nlag <- 3
seed <- 2022
prob <- 0.16
irflength <- 21
draw <- 20000
subdraw <- 200

# fit 2 regime model
data_2regime <- data %>%
  select(date, ffr, cpi, indpro, epu, epu50) %>%
  mutate(indicator = ifelse(epu > epu50, 1, 0),
         indicator_inv = ifelse(indicator == 1, 0, 1),
         ffr_up = ffr*indicator,
         cpi_up = cpi*indicator,
         indpro_up = indpro*indicator,
         ffr_dn = ffr*indicator_inv,
         cpi_dn = cpi*indicator_inv,
         indpro_dn = indpro*indicator_inv) %>% as_tibble() %>%
  select(ffr_up, cpi_up, indpro_up, ffr_dn, cpi_dn, indpro_dn) %>% ts()

est_irfs_2regime <- function(data_2regime, nlag, draw, subdraw, irflength, prob, seed){
set.seed(seed)
constr_up <- c(+1, -2, -3)
model_up <- rwz.reject(Y=data_2regime, nlags=nlag, draws=draw, subdraws=subdraw, nkeep=1000, KMIN=1,
                       KMAX=6, constrained=constr_up, constant=TRUE, steps=21)
irfs1 <- model_up$IRFS

set.seed(seed)
constr_dn <- c(+4, -5, -6)
model_dn <- rwz.reject(Y=data_2regime, nlags=nlag, draws=draw, subdraws=subdraw, nkeep=1000, KMIN=1,
                       KMAX=6, constrained=constr_dn, constant=TRUE, steps=21)
irfs2 <- model_dn$IRFS

ffr_up <- get_irf(1, prob, irflength, irfs1) %>% mutate(label = "FFR",
                                                        regime = "High uncertainty")
cpi_up <- get_irf(2, prob, irflength, irfs1) %>% mutate(label = "CPI",
                                                        regime = "High uncertainty")
indpro_up <- get_irf(3, prob, irflength, irfs1) %>% mutate(label = "INDPRO",
                                                           regime = "High uncertainty")

ffr_dn <- get_irf(4, prob, irflength, irfs2) %>% mutate(label = "FFR",
                                                        regime = "Low uncertainty")
cpi_dn <- get_irf(5, prob, irflength, irfs2) %>% mutate(label = "CPI",
                                                        regime = "Low uncertainty")
indpro_dn <- get_irf(6, prob, irflength, irfs2) %>% mutate(label = "INDPRO",
                                                           regime = "Low uncertainty")

bind_rows(ffr_up, cpi_up, indpro_up,
          ffr_dn, cpi_dn, indpro_dn) %>%
  mutate(label = factor(label, levels = c("FFR", "CPI", "INDPRO")),
         regime = factor(regime, levels = c("High uncertainty", "Low uncertainty"))) %>% 
  left_join(bind_rows(ffr_up, cpi_up, indpro_up,
                      ffr_dn, cpi_dn, indpro_dn) %>%
              mutate(label = factor(label, levels = c("FFR", "CPI", "INDPRO")),
                     regime = factor(regime, levels = c("High uncertainty", "Low uncertainty"))) %>% 
              filter(t == 0, label == "FFR") %>% 
              mutate(mult = 1/median) %>% 
              select(regime, mult), 
            by = "regime") %>% 
  mutate(median = median * mult,
         upper = upper * mult, 
         lower = lower * mult)
}

irfs_2regime <- est_irfs_2regime(data_2regime, nlag, draw, subdraw, irflength, prob, seed)

irfs_2regime %>% 
  ggplot(aes(x = t, y = median, ymin = lower, ymax = upper)) +
  geom_line() +
  geom_hline(yintercept = 0, color = "red") +
  geom_ribbon(fill="grey", alpha=.2, color="grey50", linetype="dashed") + 
  facet_wrap(regime~label, scales = "free")+
  theme_minimal() +
  theme(plot.title = element_text(size = 11, hjust=0.5),
        axis.title.y = element_text(size=11)) +
  labs(x = "",
       y ="")
```

With the results of the two-regime model in mind, I believe there is room for further developing the model to investigate the relationship between uncertainty and monetary policy. A problem of the two-regime model could be that the transmission is very similar when uncertainty is just above or just below the median - i.e. not being able to grasp the two extreme sides. Thus, a three-regime variant where the threshold points are at the upper and lower quartiles could yield us a clearer picture. The impulse responses of the three-regime model can be seen below:

```{r, fig.cap = "Impulse responses from three-regime model.",          fig.height = 6}
#fit 3 regime model
data_3regime <- data %>%
  select(date, ffr, cpi, indpro, epu, epu25, epu75) %>%
  mutate(indicator_up = ifelse(epu > epu75, 1,0),
         indicator_dn = ifelse(epu < epu25, 1, 0),
         indicator_mid = ifelse(indicator_up == 0 & indicator_dn == 0, 1, 0),
         ffr_up = ffr*indicator_up,
         cpi_up = cpi*indicator_up,
         indpro_up = indpro*indicator_up,
         ffr_mid = ffr*indicator_mid,
         cpi_mid = cpi*indicator_mid,
         indpro_mid = indpro*indicator_mid,
         ffr_dn = ffr*indicator_dn,
         cpi_dn = cpi*indicator_dn,
         indpro_dn = indpro*indicator_dn) %>% 
  as_tibble() %>%
  select(ffr_up, cpi_up, indpro_up, ffr_mid, cpi_mid, indpro_mid, ffr_dn, cpi_dn, indpro_dn) %>% ts()

est_irfs_3regime <- function(data_3regime, nlag, draw, subdraw, irflength, prob, seed){
set.seed(seed)
constr_up <- c(+1, -2, -3)
model_up <- rwz.reject(Y=data_3regime, nlags=nlag, draws=draw, subdraws=subdraw, nkeep=1000, KMIN=1,
                       KMAX=6, constrained=constr_up, constant=TRUE, steps=21)
irfs1 <- model_up$IRFS

set.seed(seed)
constr_mid <- c(+4, -5, -6)
model_mid <- rwz.reject(Y=data_3regime, nlags=nlag, draws=draw, subdraws=subdraw, nkeep=1000, KMIN=1,
                        KMAX=6, constrained=constr_mid, constant=TRUE, steps=21)
irfs2 <- model_mid$IRFS

set.seed(seed)
constr_dn <- c(+7, -8, -9)
model_dn <- rwz.reject(Y=data_3regime, nlags=nlag, draws=draw, subdraws=subdraw, nkeep=1000, KMIN=1,
                       KMAX=6, constrained=constr_dn, constant=TRUE, steps=21)
irfs3 <- model_dn$IRFS

prob <- 0.16
irflength <- 21

ffr_up <- get_irf(1, prob, irflength, irfs1) %>% mutate(label = "FFR",
                                                        regime = "High uncertainty")
cpi_up <- get_irf(2, prob, irflength, irfs1) %>% mutate(label = "CPI",
                                                        regime = "High uncertainty")
indpro_up <- get_irf(3, prob, irflength, irfs1) %>% mutate(label = "INDPRO",
                                                           regime = "High uncertainty")

ffr_mid <- get_irf(4, prob, irflength, irfs2) %>% mutate(label = "FFR",
                                                         regime = "Medium uncertainty")
cpi_mid <- get_irf(5, prob, irflength, irfs2) %>% mutate(label = "CPI",
                                                         regime = "Medium uncertainty")
indpro_mid <- get_irf(6, prob, irflength, irfs2) %>% mutate(label = "INDPRO",
                                                            regime = "Medium uncertainty")

ffr_dn <- get_irf(7, prob, irflength, irfs3) %>% mutate(label = "FFR",
                                                        regime = "Low uncertainty")
cpi_dn <- get_irf(8, prob, irflength, irfs3) %>% mutate(label = "CPI",
                                                        regime = "Low uncertainty")
indpro_dn <- get_irf(9, prob, irflength, irfs3) %>% mutate(label = "INDPRO",
                                                           regime = "Low uncertainty")

bind_rows(ffr_up, cpi_up, indpro_up,
          ffr_mid, cpi_mid, indpro_mid,
          ffr_dn, cpi_dn, indpro_dn) %>%
  mutate(label = factor(label, levels = c("FFR", "CPI", "INDPRO")),
         regime = factor(regime, levels = c("High uncertainty", "Medium uncertainty", "Low uncertainty"))) %>% 
  left_join(bind_rows(ffr_up, cpi_up, indpro_up,
                      ffr_mid, cpi_mid, indpro_mid,
                      ffr_dn, cpi_dn, indpro_dn) %>%
              mutate(label = factor(label, levels = c("FFR", "CPI", "INDPRO")),
                     regime = factor(regime, levels = c("High uncertainty", "Medium uncertainty", "Low uncertainty"))) %>% 
              filter(t == 0, label == "FFR") %>% 
              mutate(mult = 1/median) %>% 
              select(regime, mult), 
            by = "regime") %>% 
  mutate(median = median * mult,
         upper = upper * mult, 
         lower = lower * mult)

}

irfs_3regime <- est_irfs_3regime(data_3regime, nlag, draw, subdraw, irflength, prob, seed)

irfs_3regime %>% 
  ggplot(aes(x = t, y = median, ymin = lower, ymax = upper)) +
  geom_line() +
  geom_hline(yintercept = 0, color = "red") +
  geom_ribbon(fill="grey", alpha=.2, color="grey50", linetype="dashed") + 
  facet_wrap(regime~label, scales = "free")+
  theme_minimal() +
  theme(plot.title = element_text(size = 11, hjust=0.5),
        axis.title.y = element_text(size=11)) +
  labs(x = "",
       y ="")

```

As a contrast to the two-regime version, here we can see some sharper differences, hinting at the existence of the nonlinear relationship. In terms of the persistence of the FFR shock we see close to no differences, however differences in the responses of inflation and economic activity between the regimes is intriguing. The first thing we can notice is that the drop in inflation and especially output is considerably more persistent when uncertainty is high compared to the other two-regimes. Raising the interest rate is also considerably less effective at anchoring inflation as in the high uncertainty regime the impact nearly halves compared to the low uncertainty regime. The drop in output is similar in all cases, however in the high uncertainty regime it is mildly amplified - hinting at a moderately larger trade-off in output drop versus inflation anchoring capability. These results suggest that interest rate decisions alone might not be sufficient at effectively reducing inflation, thus implying the importance of utilizing additional measures such as forward guidance in order to moderate economic uncertainty. An important caveat here is that the EPU index by construction would likely be higher around the time of important policy actions such as interest rate rises. This could imply that the true impact of an interest rate decision is closer to what we see from the impulse responses of the high uncertainty regime as opposed to the other two set of impulse responses. This implication could be further reinforced by the fact that only a small proportion of observations from the high uncertainty regime fall into time periods characterized by the zero lower bound. Naturally, this raises a question about the credibility of the other two sets of impulse responses - especially the ones from the low uncertainty regime, as the numerous ZLB data points could introduce some bias to these results. In this respect, there is room for further exploring this relationship with a more complex model framework.

## Robustness Check

In order to further solidify the validity of my results, I test the robustness of my results with alternative specifications. When it comes to a model such as the one I outlined in previous sections, there are three aspects of the specification which can serve as basis for a robustness check: 1) lag structure, 2) the choice of index as the measure of uncertainty, and 3) sign restrictions. Out of these three, I address the former two, as due to the price puzzle the results do not hold without the sign restrictions. As the evidence of non-linearity is found in the three-regime model, I am testing its validity to alternate specifications.

As stated in the Empirical Strategy section, the results are slightly sensitive to the choice of lags included in the models. The median impulse responses with respect to the choice of lags included in the model can be seen in Figure 8.
```{r, fig.cap= "Robustness check with respect to lag structure" ,             fig.height = 6}
#Robcheck for nlags
nlags_list <- NULL
for(i in 1:3){
  nlags_list[[i]] <- est_irfs_3regime(data_3regime, nlag = i, draw, subdraw, irflength, prob, seed) %>% 
    mutate(lag = paste("Lags = ", i, sep = ""))
}
bind_rows(nlags_list) %>% 
  ggplot(aes(x = t, y = median, color = lag, group = lag)) +
  geom_line(size = .5) +
  geom_hline(yintercept = 0, color = "red") +
  facet_wrap(regime~label, scales = "free")+
  scale_color_viridis_d(begin = 0.05, end = 0.8, name = "Index", option = "magma")+
  theme_minimal() +
  theme(plot.title = element_text(size = 11, hjust=0.5),
        axis.title.y = element_text(size=11)) +
  labs(x = "",
       y ="") +
  theme(legend.title = element_blank())

```
Naturally, with fewer lags, the impulse responses are smoother, thus it is not an issue in terms of robustness. With both one and two lags, we can have a similar conclusion on the drop in output and the persistence of output drop increasing with higher levels of uncertainty. However, the responses of inflation are more sensitive to the number of lags included. With just a single lag, the impact of an interest rate raise does not seem to change within the regimes, and with two lagged values, the inflation anchoring capability seems to increase. This can put the validity of the results into question, however I do not find it as troubling, as with more lags included, the model is better at capturing the underlying data generating processes. With regards to higher lags, I did not further explore the results' robustness, as due to a lack of accepted draws, those results would be less credible.

The other aspect with respect to which robustness can be questioned is the choice of index from which the regime indicator function is created. As the TED spread is only available from 1986, and the VIX index from 1990, for the purposes of this robustness check the data range starts from January 1990 as opposed to the datasets of the previous simulations starting in 1985. Figure 9 shows the impulse responses with the regime indicators created from alternate uncertainty measures.

```{r, fig.cap= "Robustness check with respect to uncertainty measurement" ,             fig.height = 6}
#Robcheck for index
data <- data %>% select(date, ffr, cpi, indpro) %>% 
  inner_join(unc_measures, by = "date")

data_list <- NULL
indices <- c("epu", "ted", "vix")
for(i in 1:length(indices)){
  data_list[[i]] <- 
    data %>% 
    select(ffr, cpi, indpro, !!sym(indices[i])) %>% 
    mutate(indicator_up = ifelse(!!sym(indices[i]) > quantile(!!sym(indices[i]), 0.75), 1, 0),
           indicator_dn = ifelse(!!sym(indices[i]) < quantile(!!sym(indices[i]), 0.25), 1, 0),
           indicator_mid = ifelse(indicator_up == 0 & indicator_dn == 0, 1, 0),
           ffr_up = ffr*indicator_up,
           cpi_up = cpi*indicator_up,
           indpro_up = indpro*indicator_up,
           ffr_mid = ffr*indicator_mid,
           cpi_mid = cpi*indicator_mid,
           indpro_mid = indpro*indicator_mid,
           ffr_dn = ffr*indicator_dn,
           cpi_dn = cpi*indicator_dn,
           indpro_dn = indpro*indicator_dn) %>% 
    as_tibble() %>%
    select(ffr_up, cpi_up, indpro_up, ffr_mid, cpi_mid, indpro_mid, ffr_dn, cpi_dn, indpro_dn) %>% ts()
}
index_list <- NULL
for(i in 1:length(indices)){
  index_list[[i]] <- est_irfs_3regime(data_list[[i]], nlag, draw, subdraw, irflength, prob, seed) %>% 
    mutate(index = str_to_upper(indices[i]))
}
bind_rows(index_list) %>% 
  ggplot(aes(x = t, y = median, color = index, group = index)) +
  geom_line(size = .5) +
  geom_hline(yintercept = 0, color = "red") +
  facet_wrap(regime~label, scales = "free")+
  scale_color_viridis_d(begin = 0.05, end = 0.8, name = "Index", option = "viridis")+
  theme_minimal() +
  theme(plot.title = element_text(size = 11, hjust=0.5),
        axis.title.y = element_text(size=11)) +
  labs(x = "",
       y ="") +
  theme(legend.title = element_blank())

```

With these sets of IRF-s we can see more differences, which was expected, as they are not strongly correlated. However, by comparing each set of IRF-s with each other, most of the results that were previously derived from the model which used the EPU as a basis for its regime indicator seem to be robust. Here with each index we can observe that the inflation anchoring power drops in higher uncertainty, with the magnitude of the initial impact dropping substantially in the high uncertainty regime compared to its low counterpart. However it is important to mention that the persistence of the inflation drop is also slightly elevated in the high uncertainty regime. With respect to the responses of output, we can observe some similarities, however some conflicting outcomes as well. While with the EPU index, the output impact of an interest rate raise seems to grow larger, it is not the case with the VIX index or the TED spread. However with all indices, the persistence of the output drop does increase with higher uncertainty.

## Conclusion

This paper attempts at quantifying the impact of uncertainty on the monetary transmission mechanism. To study this nonlinear relationship I fit a two and three-regime VAR model with sign restrictions using an indicator function created from the EPU index. While in the two-regime case there seems to be close to no evidence of non-linearity in the interest rate channel, in the more extreme sides of the uncertainty distribution we can observe that high levels of uncertainty above its upper quartile seems to dampen the inflation anchoring capability of interest rate shocks at a mildly larger output drop trade-off. This would imply two conclusions: 1) as the index measuring uncertainty tracks news coverage - taking higher values around important policy decisions - it is likely the true impact of the interest rate channel is quantitatively more similar to the impact observed in high uncertainty times; and 2) interest rate decisions alone are less effective at anchoring inflation - pointing to the necessity of combining it with alternative measures like forward guidance. Quantifying the impact of such measures is however beyond the scopes of this paper.

## References
<div id="ref"></div>

### 2023-10-15 US/draft.Rmd
Initial commit

---
title: "The Impact of Uncertainty on Monetary Transmission"
subtitle: "Evidence from the US"
author: "Péter Horváth" 
#date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  pdf_document:
    #toc: true
    #toc_depth: 3
    includes:
      in_header: "preamble.tex"
      before_body: "before_body.tex"
    number_sections: true
abstract: "This paper investigates the nonlinear interaction of the monetary transmission mechanism and economic uncertainty in the United States. I employ a Threshold VAR model with two states dependent on the level of uncertainty, and identify regime specific shocks using sign-restrictions. Results suggest that \\textcolor{red}{FILL IN THE BLANK}. Using a number of alternative specifications as robustness checks \\textcolor{red}{FILL IN THIS BLANK TOO}. Implications for policymakers are \\textcolor{red}{BLANK NUMBER THREE TO BE FILLED}."
fontsize: 12pt
bibliography: "ref.bib"
csl: "custom-citations.csl"
link-citations: true
linkcolor: "blue"
geometry: "a4paper,outer=25mm,inner=25mm,top=25mm,bottom=25mm"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      results = FALSE)

library(tidyverse)
library(readr)
library(lubridate)
library(stringr)
library(rebus)

lw = .5

source('codes/data2.R')

quarterly_format <- function(input_date){
require(lubridate)

# Your input date string
input_date <- input_date

# Convert the input date to a date object
date_obj <- ymd(input_date)

# Extract the year and quarter
year <- year(date_obj)
quarter <- quarter(date_obj)

# Create the desired output format "YYYY-QN"
output_format <- paste0(year, "-Q", quarter)

output_format
}

estout_clean <- read_rds('data/estout_clean.Rds')

irfs <- estout_clean[['irfs']]
```

# Introduction

This paper serves as a natural followup to @horvathTBAuncertain. The paper establishes how the interest rate channel of the monetary transmission intertwined with economic uncertainty influences economies from across the globe. The results indicate that in times of high uncertainty, interest rate shocks lead to a larger contraction of macroeconomic demand, primarily reflecting in a larger and longer lasting slump in economic activity.

While the main implications are clear, there is still room for further investigation. As a result of using a wide variety of countries for the estimation, some factors - such as a more nuanced investigation into the source of economic uncertainty; checking to what extent the Zero Lower Bound influences the results; whether or not excess financial stress caused by uncertainty is reflected in macroeconomic aggregates; or whether uncertainty is endogenous to policy shocks - was not possible to incorporate. This paper intends to fill this knowledge gap, by changing from a birds-eye view to a more microscopic perspective focusing on implications solely from the United States.

I adopt the same Threshold Vector-Autoregression (TVAR) model as well as sign-restriction approach to identify regime specific shocks. Following the previous paper, the core data behind the estimations are macroeconomic aggregates on a quarterly frequency, consisting of interest rates, inflation and GDP.

However, given the wider range of data available for the US, there is room for a number of alternative specifications, such as using the @wu2016measuring shadow rate to measure implied interest rates at the ZLB, or incorporate the St. Louis Fed Financial Stress Index as a control variable in the model.

As opposed to using the global versions of uncertainty indices, I switch to the US specific ones. For the main measure of uncertainty, I use the @baker2016measuring Economic Policy Uncertainty (EPU) index. As the EPU for the US has a number of supplementary Categorical indices (e.g.: indices specific to Monetary Policy, Financial Regulation, or Sovereign debt and currency crises) available, I also explore how different sources of uncertainty influence the transmission mechanism.

The potential endogeneity issue between uncertainty and policy shocks is addressed using simple Granger-causality tests, as well as explicitly incorporating the uncertainty index into the TVAR model as a control variable.

\textcolor{red}{WRITE SOMETHING ABOUT THE RESULTS HERE 1-2 PARAGRAPHS.}

The rest of the paper is outlined as follows: Section 2 gives a brief overview of the related literature, Section 3 describes the data used in this paper, Section 4 discusses the methodology, Section 5 showcases the results and their implications and Section 6 concludes.

# Related literature

Following the seminal works such as @baker2016measuring or @caldara2022measuring, a vast body of literature has emerged in the field of empirical macroeconomics focusing on economic uncertainty and its implications for economies. The aforementioned papers establish the key implications on the economy. Uncertainty shocks cause a demand-side contraction, the primary channel being uncertainty disincentivizing investments. Moreover (specifically contractive) uncertainty shocks have a more pronounced impact in times of economic turmoil and financial stress.

Following such works, @caldara2016macroeconomic show that financial and uncertainty shocks are pivotal drivers of business cycles, exemplified during events like the Great Recession. Likewise, @bonciani2020international show that global financial uncertainty shocks bear down on economies, particularly in open economies during downturns.

The global importance of uncertainty shocks is further studied in works such as @colombo2013economic, @nilavongse2020economic or @biljanovska2021fear, showing that uncertainty shocks of large economies - most predominantly shocks in the US - have a considerable spillover effect on economies such as the UK, Euro Area economies, as well as countries of the Western Hemisphere.

@cheng2018important show that geopolitical uncertainty is a significantly influential factor in driving business cycle fluctuations of emerging economies. Works such as @carriere2013impact, @ahir2019caution, @ahir2022world show that these economies suffer greater downturns as the impact of uncertainty shocks is magnified by credit constraints, albeit there is some dispute in these findings, as @das2018international argue the opposite.

As demonstrated by a number of papers such as @ebeke2018trade, @wang2021trade, @chen2021tariff, and @william2022economic, shocks to uncertainty related to trade policy are shown to exert similar adverse effects on investment by reducing risk appetite; additionally increasing firm markups as well as deterring exports.

On the methodological side of the literature, the use of nonlinear models - such as threshold or smooth-transition techniques - for modeling time series have been steadily rising. TVAR models - as a method of introducing state dependency into multivariate time series models - have seen a wide range of use cases in works such as @baum2011impact studying the state dependency of fiscal multipliers along the business cycle; @galvao2006structural using such models to predict recession timings; or @alessandri2019financial showing the state dependency of uncertainty shocks on financial conditions.

Such techniques are used for research on monetary policy as well. @avdjiev2014credit show that policy shocks have a more pronounced impact during low-growth periods while monetary policy having a aggressive responses during booms; @schmidt2020risk find evidence for state-dependency on asset prices, with macro-risk and policy risk playing distinct roles for the transmission; while @li2010financial and later @fry2016effects highlight monetary policy efficacy in times of financial stress, emphasizing short-term trade-offs between output and inflation.

The state-dependent nature of uncertainty shocks is further established using nonlinear time series techniques. @schuler2014asymmetric and @colombo2020asymmetric both find that uncertainty shocks have a more significant effect during recessions, also showing the vital role of monetary policy measures - especially balance sheet-based tools - in mitigating such shocks. @nalban2021asymmetric also shows that the impact of uncertainty varies with the economic state, amplifying during financial distress but rebounding faster due to monetary policy reactions.

Works such as @aastveit2017economic and @aquino2022does investigate the state-dependency of monetary shocks conditional on economic uncertainty. While the methods differ slightly, their results suggest that economic uncertainty hinders the transmission completely, as the responses of both output and inflation are decreased in high uncertainty times compared to the responses in times of low uncertainty.

@aastveit2017economic uses an interacted SVAR approach, where the uncertainty-based regime indicator is interacted with only the interest rate in the model, all else is kept in a linear form. The regimes are selected based on the very extreme sides of the distribution however, which is criticized by @aquino2022does. The latter introduces the threshold assumption into the interest rate equation of the SVAR model - implying a state dependent Taylor Rule - and additionally classify the high and low regimes using a grid-search algorithm[^1].

[^1]: Results from this algorithm in the paper suggest the regime switches are at the 57th, 58th and 65th percentile of the historical distribution depending on the index. While it is true, that the correct specification of the threshold value is important for goodness-of-fit, the more ad-hoc definition of high uncertainty at the median and the 70th percentile is kept as in @horvathTBAuncertain for the sake of simplicity and comparability.

While theoretical arguments are compelling, taking into account how the equations describing the evolution of inflation and economic activity change across regimes should be paramount in measuring how the transmission changes with the uncertainty regimes. Additionally, arguments can be made for i) high uncertainty inflating the variance of outcomes in periods affected and thus enlarging the impact of all shocks; ii) uncertainty influencing the future expectations of agents; iii) uncertainty can introduce excess financial stress and thus influencing the transmission through this channel; or iv) uncertainty could prove to be endogenous to policy shocks.

Works that might support the results of this paper could be @gbohoui2021uncertainty where an increase in investment multipliers is shown in times of high uncertainty, implying larger and longer lasting effects; or @ying2022role finding that monetary policy uncertainty moderates the transmission of forward guidance shocks on long-term yields, but not the transmission of Federal Funds Rate or asset price shocks, furthermore, the moderation effect is shown to be a consequence of changes in the term-premium and not through the channel of short rates.

# Data

## Macroeconomic aggregates

In order to keep the estimation streamlined, I rely on quarterly macroeconomic aggregates from `r max(min(data2[['uncertainty']][['date']]), min(data2[['macro']][['date']])) %>% quarterly_format()` to `r min(max(data2[['uncertainty']][['date']]), max(data2[['macro']][['date']])) %>% quarterly_format()`[^2] as the core set of data behind the estimations. The three main aggregates are interest rates, quarterly inflation rates - measured as the first difference of the CPI - and quarterly real GDP growth rates - measured as the first difference of the chain volume index calculated from quarterly real GDP.

[^2]: The range of data does change in some specifications, as the Shadow Rate or the St. Louis Fed Financial Stress Index are only available on a shorter period of time.

Figure \ref{fig:interest} below shows the three measures of interest rates I consider in different specifications. In the first set of estimations, policy shocks are measured as innovations of the Federal Funds Rate (FFR). In order to add some further nuance to the estimations, two other series are also considered.

```{=latex}
\begin{figure}[!htbp]
```
```{r interest rates}
full_join(
  data2[["macro"]] %>% 
    select(date, ffr = r),
  data2[["macro_alt"]] %>% 
    select(date, r1y, ffr_shadow),
  by = 'date') %>% 
  gather(variable, value, -date) %>% 
  mutate(variable = case_when(variable == 'ffr' ~ 'Federal Funds Rate',
                              variable == 'ffr_shadow' ~ 'Shadow rate',
                              variable == 'r1y' ~ '1-Year Treasury Yield')) %>% 
  ggplot(aes(x = date, y = value, color = variable)) +
  geom_line(linewidth = lw) +
  scale_color_manual(breaks = c('Federal Funds Rate',
                                'Shadow rate',
                                '1-Year Treasury Yield'),
                     values = c('dodgerblue',
                                'firebrick',
                                'limegreen')) +
  theme_minimal() +
  theme(legend.title = element_blank(),
        legend.position = 'bottom') +
  labs(x = "",
       y = "")

```
```{=latex}
\caption{Interest rates \label{fig:interest}}
\end{figure}
```

In an alternative specifications the FFR is replaced with the 1-Year Treasury Yield. While the two show high co-movement, the latter may carry some additional information such as market expectations or yield curve dynamics, that might have an impact on the results.

A further specification is considered where interest rates are measured by the @wu2016measuring shadow rate. While through the majority of the sample, the FFR and its shadow rate are highly correlated, the latter contains information about the implied rate of interest in the time period between 2008 and 2015 or 2020 and 2021 where the interest rates were close to the Zero Lower Bound. Estimating with the shadow rate is important as sign-restriction algorithms could lead to biased outcomes as a result of the ZLB periods.

As shown by @li2010financial and @fry2016effects, in times of high financial stress, monetary policy can potentially become more effective. As elevated levels of uncertainty are likely associated with higher financial stress, it might be necessary to control for it as well. For this reason, additional to the standard macroeconomic aggregates, the St. Louis Fed Financial Stress Index is considered as a control variable.

## Measuring uncertainty

In the models outlined, my primary measure of uncertainty is the @baker2016measuring EPU index. This measure of uncertainty is constructed from textual data retrieved from articles of influential newspapers in the US. Using the articles, automated text queries are conducted in order to find articles containing terms related to "uncertainty" and either "economic" or "policy". [^3]

[^3]: The data series of uncertainty measures along with more detailed descriptions of each can be retrieved from [policyuncertainty.com](https://www.policyuncertainty.com/).

Why use measures constructed from textual data? A benefit of using the EPU over more standard measures such as the VIX is that it is more tailored to "economic" or "economic policy" related uncertainty, and thus should be more influential on the impact of policy shocks. Additionally, as opposed to being constructed from financial market data, the news articles might be a more suitable source of information to capture expectations of agents in the economy.

In order to define what "high uncertainty" is, I consider two alternatives based on position in the historical distribution of the EPU. In the first case, high uncertainty is defined as values of the EPU above its median, while in the second case, above its 70th percentile. While these are more ad-hoc definitions, for the sake of better conveying the implications of the results, I prefer using such pre-defined levels over an endogenously defined one from a Threshold AR model. The EPU along with the two threshold values can be seen in Figure \ref{fig:epu}.

```{=latex}
\begin{figure}[!htbp]
```
```{r epu}

full_join(
  data2[['uncertainty']],
  data2[['forrob']],
  by = 'date') %>%
  select(date, epu) %>% 
  mutate(med = quantile(epu, prob = .5),
         pc70 = quantile(epu, prob = .7)) %>% 
  ggplot(aes(x = date, y = epu)) + 
  geom_line(linewidth = .7) +
  geom_line(aes(x = date, y = med), color = '#0074D9', linetype = 'dashed', linewidth = .7) + 
  geom_line(aes(x = date, y = pc70), color = '#FF4136', linetype = 'dashed', linewidth = .7) +
  theme_minimal() +
  labs(x = "",
       y = "")

```

```{=latex}
\caption{Economic Policy Uncertainty index \label{fig:epu}}
\bigskip
\footnotesize
\textit{Notes}: The blue and red lines indicate the median and 70th percentile of the historical distribution respectively.
\end{figure}
```

In @baker2016measuring, along with the main EPU index, several EPU Categorical indices are also constructed. These indices are more specific to a certain topic within "economic" or "economic policy" such as monetary policy, fiscal policy, or sovereign debt crises. In order to pinpoint the source of uncertainty that might be most influential for the transmission of monetary policy shocks, I repeat the estimations with the EPU Categorical indices in place of the main EPU index. The EPU categorical indices can be seen in Figure \ref{fig:epucat} of the Appendix.

It is important to mention that there are a number of alternative^[These measures can be seen in Figure \ref{fig:unces} of the Appendix.] measurements of uncertainty in the literature. These indices are for example the Equity Market-related Economic Uncertainty Index of @baker2016measuring; the Geopolitical Risk Index and the Trade Policy Uncertainty Index of @caldara2022measuring and @caldara2020economic; the World Uncertainty Index of @ahir2022world; as well as the VIX Index and the TED spread. In this paper, only the EPU and its Categorical indices are considered as measures of uncertainty.

# Methodology

## Threshold VAR model

As discussed in the Introduction and Literature sections, TVAR models are suitable for modeling non-linearity problems, such as the one discussed in this paper. For the sake of comparability, I adopt the same baseline setup as in @horvathTBAuncertain, however, in alternative specifications, some other factors are considered. Thus the baseline model can bet written as: 
\begin{equation} \label{eq:1}
  Y_{t} = \sum\limits_{k=1}^{4}\Theta_{high}I(X_{t} \ge \mu)Y_{t-k} + \sum\limits_{k=1}^{4}\Theta_{low}I(X_{t} < \mu)Y_{t-k} + \epsilon_{t}
 \end{equation} where $Y_{t}$ is the vector of endogenous variables, $\Theta_{low}$ and $\Theta_{high}$ are the respective coefficient matrices for each regime, $I(.)$ is the regime indicator function, $X_{t}$ is threshold variable, $\mu$ is the threshold value and $\epsilon_{t}$ is the error term.

In this main specification, $Y_{t}$ consists of the FFR, quarterly CPI inflation and quarterly GDP growth and $X_{t}$ is the EPU index. To investigate how different sources of economic uncertainty impact the transmission mechanism, the estimation is repeated by replacing plugging in each EPU categorical index into $X_{t}$ in place of the main EPU index. Two different values of $\mu$ are considered, the median value as well as the 70th percentile of $X_{t}$ in each estimation. This allows to check i) how policy shocks affect the economy conditional on the level of economic uncertainty experienced in the same period, ii) to some extent, how much the level of uncertainty matters for the transmission mechanism, and iii) which source of uncertainty is the most influential in altering the transmission channel.

To address some potential issues, a number of robustness checks are conducted. Firstly, I replace the FFR in $Y_ {t}$ with the 1-Year Treasury Yield and the FFR shadow rate. Secondly, I check if the results hold if I change the construction of the regime indicator from $X_{t}$ to $X_{t-1}$, making it pre-determined, and thereby eliminating the potential of estimating the impact of two contemporaneous shocks. Thirdly, I incorporate the uncertainty index in its levels as well as a financial stress indicator in order to void their effects from the aggregates during the estimation^[Please refer to the [Alternative model specifications] of the Appendix for further details on how Equation \ref{eq:1} above changes].

## Identification

The identification of shocks is done using sign-restrictions, as with a relatively compact set of variables, this is the most straight-forward method. As standard macroeconomic theory dictates, I assume that an increase in interest rates should be a negative demand-type shock, and thus lead to a contraction of output and prices. In other terms, a positive sign-restriction is imposed on the interest rates, and a negative on inflation rates as well as GDP growth. Following @uhlig2005effects, the restriction is imposed on the contemporaneous effects, as well as up to 6 quarters ahead after the shock.

The algorithms proposed in @uhlig2005effects has some shortcomings that should be addressed. As pointed out by @arias2018inference, using the penalty algorithm has some adverse consequences. Firstly, this algorithm can introduce some unintentional sign restrictions on unrestricted variables, thereby leading to biased results. For the main round of estimations, this is no issue, as I work with a full set of sign restrictions. However more notably, the penalty algorithm accepts a larger number of posterior draws, thus narrowing confidence bands. Opposing to this, the rejection algorithm only accepts posterior draws where the impulse responses that exactly fit the imposed sign-restrictions. Additionally, @rubio2010structural proposed a more efficient rejection algorithm, and thus their improved algorithm is implemented in this paper.

In all cases, the models are estimated using Bayesian methods with a non-informative Wishart prior. The estimations yield a set of impulse response draws, which are aggregated into posterior-median impulse responses along with an 84% confidence interval. All computations are done in the R programming language relying on the work of @Danne2015 for the implementation of the sign-restriction algorithms.

\textcolor{red}{I SHOULD WRITE SOMETHING ABOUT ENDOGENEITY CONCERNS AND HOW THEY ARE ADDRESSED.}

# Results

## Replicating the baseline specification

In the baseline specification, the three equation TVAR is estimated on interest rates, CPI inflation and GDP growth using a four-quarter lag structure. Monetary shocks are measured as innovations of the Federal Funds Rate, and the threshold value $\mu$ is calculated as the median of the EPU index. Figure \ref{fig:epumedirfs} shows the regime specific impulse responses of inflation and economic activity to a 1 percentage point rise in interest rates.

```{=latex}
\begin{figure}[!htbp]
```
```{r epumedirfs}

irfs %>% 
  select(-mult, -restr) %>% 
  filter(label != 'R') %>% 
  filter(spec == 'main',
         control == 'none',
         index == 'epu') %>% 
  ggplot(aes(x = t, y = median, color = thresh, fill = thresh,
             ymin = lower, ymax = upper
  )) +
  geom_line(linewidth = lw) +
  geom_hline(yintercept = 0, color = "red") +
  geom_ribbon(alpha=.1, linetype=0) + 
  facet_wrap(regime~label, scales = "free")+
  theme_minimal() +
  theme(plot.title = element_text(size = 11, hjust=0.5),
        axis.title.y = element_text(size=11),
        legend.position = 'none') +
  scale_color_manual(breaks = c('0.5', '0.7'),
                     values = c('#0074D9', '#FF4136')) +
  scale_fill_manual(breaks = c('0.5', '0.7'),
                    values = c('#0074D9', '#FF4136')) +
  labs(x = "",
       y ="")

```

```{=latex}
\caption{Regime specific impulse responses, original specification. \label{fig:epumedirfs}}
\bigskip
\footnotesize
\textit{Notes}: The blue solid line represents the posterior median obtaied with the threshold value $\mu$ set to the median, the red one with it set to the 70th percentile of the EPU. The shaded areas represent the 84\% confidence bands calculated from the posterior impulse response draws.
\end{figure}
```

There are some noteworthy differences from these results compared to previous literature. As discussed in the Literature section of this paper, @aastveit2017economic and @aquino2022does show that the transmission efficacy is eroded by high economic uncertainty, while @horvathTBAuncertain find a considerably more prolonged period of contraction induced by monetary shocks in high uncertainty.

Results from this estimate suggest that the contemporaneous drop in output is slightly larger in low uncertainty, with the following quarters showing a similar level of contraction. Alongside this, disinflation caused by the monetary shock is significantly enhanced in the high uncertainty regime compared to the low uncertainty one.

Comparing the median and the 70th percentile as threshold levels for the regime switches, the responses of inflation in the lower regime and of output in the higher regime are not too dissimilar, however their opposite uncertainty regime counterparts show some key differences. Firstly, the output contraction in the low regime is elevated when switching to the higher uncertainty threshold. This is in line with the findings of @horvathTBAuncertain, as allocating a larger proportion of observations previously labelled as high uncertainty would support the notion of larger contraction of high uncertainty times. Secondly, we can observe that the response of inflation shifts upwards as we increase the threshold level - albeit retaining a similar shape. This lies closer to the findings of @aastveit2017economic and @aquino2022does as the implication from this suggest that more extreme levels of uncertainty would erode the transmission.

\textcolor{red}{EZT AMÚGY MAJD MÁSHOVA RAKNÁM.}^[To further clarify, global uncertainty measures are likely better at capturing one of the mechanisms, through which geoeconomic fragmentation impacts economies, as noted in @ilyna2023geo. While here and and in the cited previous literature, country-specific uncertainty measures carry a different implication, whereby policy uncertainty explicitly reduces the efficacy of the transmission of policy actions.]

## An updated specification

Through conducting a number of robustness tests, controlling for the level of uncertainty, using a more traditional TVAR model where the indicator is constructed from $t-1$ values of the threshold variable, or using 1-Year rates instead of the FFR has no significant impact on the results. However, controlling for the Zero-Lower-Bound, as well as financial stress both change the results significantly.

```{=latex}
\begin{figure}[!htbp]
```
```{r}
irfs %>% 
  select(-mult, -restr) %>% 
  filter(label != 'R') %>% 
  filter(spec == 'shadow',
         control == 'financial_stress',
         index == 'epu',
         thresh == '0.7') %>% 
  ggplot(aes(x = t, y = median)) +
  geom_line(linewidth = lw) +
  geom_hline(yintercept = 0, color = "red") +
  facet_wrap(regime~label, scales = "free")+
  theme_minimal() +
  theme(plot.title = element_text(size = 11, hjust=0.5),
        axis.title.y = element_text(size=11),
        legend.position = 'none') +
  labs(x = "",
       y ="")
```
```{=latex}
\caption{Regime specific impulse responses, updated specification. \label{fig:epumedirfs}}
\bigskip
\footnotesize
\textit{Notes}: Interest rates are measured by the Shadow Rate, the threshold value is set to the 70th percentile of the EPU index.
\end{figure}
```

## Narrative of uncertainty

First, the "narrative" of the uncertainty index could matter. In @baker2016measuring, additional to the EPU index, several categorical indices are constructed^[Please refer to Figure \ref{fig:epucat}.]. Replicating the simulations with each categorical index could provide insight into how such narratives can impact the transmission mechanism. Notably, uncertainty specifically related to monetary policy should be the most insightful. Below in Figure \ref{fig:narrative}

```{=latex}
\begin{figure}[!htbp]
```
```{r narrative, fig.height=10}
irfs %>% 
  select(-mult, -restr) %>% 
  filter(label != 'R',
         spec == 'shadow',
         index != 'epu',
         control == 'financial_stress') %>% 
  group_by(label, regime, spec, control, thresh, index) %>% 
  mutate(median = purrr::accumulate(.x = median, 
                                    .f = sum,
                                    .dir = 'forward'),
         upper = purrr::accumulate(.x = upper, 
                                    .f = sum,
                                    .dir = 'forward'),
         lower = purrr::accumulate(.x = lower, 
                                    .f = sum,
                                    .dir = 'forward')) %>% 
  filter(t == max(t)) %>% 
  select(-t, -upper, -lower) %>% 
  spread(label, median) %>% 
  mutate(sac = Y/PI,
         regime = str_remove_all(regime, ' uncertainty'),
         thresh = case_when(thresh == '0.5' ~ 'Median',
                            thresh == '0.7' ~ '70th Perc.'),
         thresh = factor(thresh, levels = c('Median', '70th Perc.')),
         index = case_when(index == 'epu' ~ 'EPU',
                           index == 'epu_debt' ~ 'EPU: Debt, crises',
                           index == 'epu_finreg' ~ 'EPU: Fin. Regulation',
                           index == 'epu_fis' ~ 'EPU: Fiscal Policy',
                           index == 'epu_mon' ~ 'EPU: Monetary Policy',
                           index == 'epu_reg' ~ 'EPU: Regulation',
                           index == 'epu_trade' ~ 'EPU: Trade Policy')) %>% 
  ggplot(aes(x = regime, y = sac, fill = regime)) + 
  geom_col( color = 'black', width = 0.3, position = position_dodge(width = 0.1)) +
  facet_grid(index~thresh, scales = 'free') +
  theme_minimal()+
  theme(legend.position = 'none') +
  labs(x = '',
       y = '') + 
  scale_fill_manual(breaks = c('High',
                                'Low'),
                     values = c('#8B0000',
                                '#00008B'))

```
```{=latex}
\caption{Sacrifice ratios. \label{fig:narrative}}
\bigskip
\footnotesize
\textit{Notes}: The sacrifice ratios are calculated from the accumulated impact of the policy shocks over the 20 quarter horizon.  
\end{figure}
```

## Endogeneity

# Conclusion

# References

```{=tex}
\bgroup
\renewcommand{\baselinestretch}{1}\normalsize
```
::: {#refs}
:::

::: {#ref}
:::

```{=tex}
\egroup
\newpage
```
# Appendix

## Alternative uncertainty indices

```{=latex}
\begin{figure}[!htbp]
```
```{r epucategorical}

full_join(
  data2[['uncertainty']],
  data2[['forrob']],
  by = 'date') %>% 
  select(- epu) %>% 
  gather(index, value, -date) %>% 
  drop_na() %>%
  group_by(index) %>% 
  mutate(med = quantile(value, prob = .5),
         pc70 = quantile(value, prob = .7)) %>% 
  filter(index %in% c('epu_debt', 
                      'epu_finreg', 
                      'epu_fis', 
                      'epu_mon',
                      'epu_reg',
                      'epu_trade')) %>% 
  mutate(index = case_when(index == 'epu_debt' ~ 'EPU: Debt, crises',
                           index == 'epu_finreg' ~ 'EPU: Fin. Regulation',
                           index == 'epu_fis' ~ 'EPU: Fiscal Policy',
                           index == 'epu_mon' ~ 'EPU: Monetary Policy',
                           index == 'epu_reg' ~ 'EPU: Regulation',
                           index == 'epu_trade' ~ 'EPU: Trade Policy')) %>% 
  ggplot(aes(x = date, y = value)) +
  geom_line(linewidth = lw)+
  geom_line(aes(x = date, y = med), color = '#0074D9', linetype = 'dashed', linewidth = lw) + 
  geom_line(aes(x = date, y = pc70), color = '#FF4136', linetype = 'dashed', linewidth = lw) +
  facet_wrap(~index, scales = 'free') + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
  labs(x = "",
       y = "")

```

```{=latex}
\caption{EPU Categorical uncertainty indices \label{fig:epucat}} 
\bigskip
\footnotesize
\textit{Notes}: The blue and red lines indicate the median and 70th percentile of the historical distribution respectively.
\end{figure}
```

```{=latex}
\begin{figure}[!htbp]
```
```{r otherunc}

full_join(
  data2[['uncertainty']],
  data2[['forrob']],
  by = 'date') %>% 
  select(- epu) %>% 
  gather(index, value, -date) %>% 
  drop_na() %>%
  group_by(index) %>% 
  mutate(med = quantile(value, prob = .5),
         pc70 = quantile(value, prob = .7)) %>%   
  filter(index %in% c('equnc', 
                      'gpr',
                      'ted',
                      'tpu', 
                      'vix',
                      'wui')) %>% 
  mutate(index = case_when(index == 'equnc' ~ 'Eq. Market Unc.',
                           index == 'gpr' ~ 'GPR',
                           index == 'ted' ~ 'TED spread',
                           index == 'tpu' ~ 'TPU',
                           index == 'vix' ~ 'VIX',
                           index == 'wui' ~ 'WUI')) %>% 
  ggplot(aes(x = date, y = value)) +
  geom_line(linewidth = lw)+
  geom_line(aes(x = date, y = med), color = '#0074D9', linetype = 'dashed', linewidth = lw) + 
  geom_line(aes(x = date, y = pc70), color = '#FF4136', linetype = 'dashed', linewidth = lw) +
  facet_wrap(~index, scales = 'free') + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
  labs(x = "",
       y = "")

```

```{=latex}
\caption{Additional uncertainty measures \label{fig:unces}}
\bigskip
\footnotesize
\textit{Notes}: The blue and red lines indicate the median and 70th percentile of the historical distribution respectively.
\end{figure}
```

## Alternative model specifications

To address robustness concerns, some further specifications are considered. As described in the Methodology section, the first round of robustness checks are done by swapping out the FFR for alternative interest rate aggregates. Additionally, a second check is done by constructing the regime indicator from the $t-1$ level of uncertainty as opposed to the $t$ period level.

In the thrird round of robustness checks, additional variables are considered as controls in the TVAR equation. In one specification, I control for the level of uncertainty explicitly in the model, in another specification the level of financial stress, and finally both. Respectively, the TVAR equation changed for each of these specifications in order are:

\begin{equation} \label{eq:2}
  Y_{t} = \sum\limits_{k=1}^{4}\Theta_{high}I(X_{t} \ge \mu)Y_{t-k} + \sum\limits_{k=1}^{4}\Theta_{low}I(X_{t} < \mu)Y_{t-k} + \sum\limits_{k=1}^{4}\beta X_{t-k} + \epsilon_{t}
\end{equation}

\begin{equation} \label{eq:3}
  Y_{t} = \sum\limits_{k=1}^{4}\Theta_{high}I(X_{t} \ge \mu)Y_{t-k} + \sum\limits_{k=1}^{4}\Theta_{low}I(X_{t} < \mu)Y_{t-k} + \sum\limits_{k=1}^{4}\gamma Z_{t-k} + \epsilon_{t}
\end{equation}

where $X_{t-k}$ represents the inclusion of four lags of the uncertainty indicator, and $Z_{t-k}$ represent the inclusion of four lags of the St. Louis Fed Financial Stress Index.

## Robustness checks

As noted in the the Methodology section, there could be two arguments against the specification used to obtain the results seen above. One of them being using time $t$ values of the EPU to specify the regimes, the other being that the variables should be voided of the share of variance explained by uncertainty. The former is achievable by changing Equation \ref{eq:1} to use $X_{t-1}$ instead of $X_{t}$ for specifying the regime indicator, while the latter is done as shown in \ref{eq:2}. Figure \ref{fig:contemp} shows that the outcomes obtained from such alternative specifications is largely similar. Two notable differences are i) when uncertainty is controlled for and the $t-1$ value of the index is used for constructing the indicator, the high uncertainty inflation response is slightly dampened; and ii) both inflation and output responses sharpen in high uncertainty when switching from a $t$ to a $t-1$ quarter-based indicator. These however do not cause robustness concerns and indicate that the four-quarter lag of each variable are sufficient controls, as the former of the two differences is relatively minor, while the latter is washed out of the results once the level of uncertainty is also controlled for - yielding results alike to the baseline specification of Equation \ref{eq:1}.

```{=latex}
\begin{figure}[!htbp]
```
```{r contemp}
cowplot::plot_grid(
irfs %>% 
  select(-mult, -restr) %>% 
  filter(label != 'R',
         index == 'epu',
         thresh == '0.5',
         spec == 'lag' & control == 'none' | spec == 'main' & control == 'uncertainty' | spec == 'main' & control == 'none' | spec == 'lag' & control == 'uncertainty') %>% 
  mutate(specs = case_when(spec == 'lag' & control == 'none' ~ 'lag',
                           spec == 'main' & control == 'uncertainty' ~ 'control',
                           spec == 'main' & control == 'none' ~ 'base',
                           spec == 'lag' & control == 'uncertainty' ~ 'both')) %>% 
  ggplot(aes(x = t, y = median, color = specs
  )) +
  geom_line(linewidth = lw) +
  geom_hline(yintercept = 0, color = "red") +
  facet_wrap(regime~label, scales = "free")+
  scale_color_manual(breaks = c('base', 'both', 'control', 'lag'),
                     values = c('#000000', '#0072B2', '#FF0000', '#009E73')) +
  theme_minimal() +
  theme(plot.title = element_text(size = 11, hjust=0.5),
        axis.title.y = element_text(size=11),
        plot.caption = element_text(hjust=0.5, size=rel(1.2)),
        legend.position = 'none') +
  labs(x = "",
       y ="",
       caption ="μ = 0.5"),
irfs %>% 
  select(-mult, -restr) %>% 
  filter(label != 'R',
         index == 'epu',
         thresh == '0.7',
         spec == 'lag' & control == 'none' | spec == 'main' & control == 'uncertainty' | spec == 'main' & control == 'none' | spec == 'lag' & control == 'uncertainty') %>% 
  mutate(specs = case_when(spec == 'lag' & control == 'none' ~ 'lag',
                           spec == 'main' & control == 'uncertainty' ~ 'control',
                           spec == 'main' & control == 'none' ~ 'base',
                           spec == 'lag' & control == 'uncertainty' ~ 'both')) %>% 
  ggplot(aes(x = t, y = median, color = specs
  )) +
  geom_line(linewidth = lw) +
  geom_hline(yintercept = 0, color = "red") +
  facet_wrap(regime~label, scales = "free")+
  scale_color_manual(breaks = c('base', 'both', 'control', 'lag'),
                     values = c('#000000', '#0072B2', '#FF0000', '#009E73')) +
  theme_minimal() +
  theme(plot.title = element_text(size = 11, hjust=0.5),
        axis.title.y = element_text(size=11),
        plot.caption = element_text(hjust=0.5, size=rel(1.2)),
        legend.position = 'none') +
  labs(x = "",
       y ="",
       caption ="μ = 0.7")
)

```
```{=latex}
\caption{Influence of uncertainty \label{fig:contemp}}
\bigskip
\footnotesize
\textit{Notes}: The black line shows the base specification, the green, red and blue lines show results obtained with one-quarter lag uncertainty-based regime indicator, uncertainty added as a control variable, and the combination of these two specifications respectively.
\end{figure}
```

As discussed in the Data section of this paper, there are alternative interest rate measures to calculate policy shocks from. Using the 1-Year Treasury Yield over the Federal Funds Rate has the benefit of - to some extent - carrying information about future expectations. The other alternative is the @wu2016measuring Shadow Rate, the added benefit of which is the additional information about implied rates of interest in times when the former two were close to the Zero-Lower-Bound. The data for the Shadow Rate is available in a shorter time span from `r min(drop_na(select(data2[['macro_alt']], date, ffr_shadow))$date) %>% quarterly_format()` to `r max(drop_na(select(data2[['macro_alt']], date, ffr_shadow))$date) %>% quarterly_format()`, however the exclusion of data from before the 1990s as well as 2022 should not cause concerns.

```{=latex}
\begin{figure}[!htbp]
```
```{r shockmes}
cowplot::plot_grid(
irfs %>% 
  select(-mult, -restr) %>% 
  filter(label != 'R') %>% 
  filter(spec!= 'lag',
         control == 'none',
         thresh == '0.5',
         index == 'epu') %>% 
  mutate(spec = case_when(spec == 'main' ~ 'FFR',
                          spec == 'r1y' ~ 'Treasury',
                          spec == 'shadow' ~ 'Shadow')) %>% 
  ggplot(aes(x = t, y = median, color = spec
  )) +
  geom_line(linewidth = lw) +
  geom_hline(yintercept = 0, color = "red") +
  facet_wrap(regime~label, scales = "free")+
  theme_minimal() +
  theme(plot.title = element_text(size = 11, hjust=0.5),
        axis.title.y = element_text(size=11),
        plot.caption = element_text(hjust=0.5, size=rel(1.2)),
        legend.position = 'bottom',
        legend.title = element_blank()) +
  labs(x = "",
       y ="",
       caption = "μ = 0.5") + 
  scale_color_manual(breaks = c('FFR',
                                'Shadow',
                                'Treasury'),
                     values = c('dodgerblue',
                                'firebrick',
                                'limegreen')),
irfs %>% 
  select(-mult, -restr) %>% 
  filter(label != 'R') %>% 
  filter(spec!= 'lag',
         control == 'none',
         thresh == '0.7',
         index == 'epu') %>% 
  mutate(spec = case_when(spec == 'main' ~ 'FFR',
                          spec == 'r1y' ~ 'Treasury',
                          spec == 'shadow' ~ 'Shadow')) %>%  
  ggplot(aes(x = t, y = median, color = spec
  )) +
  geom_line(linewidth = lw) +
  geom_hline(yintercept = 0, color = "red") +
  facet_wrap(regime~label, scales = "free")+
  theme_minimal() +
  theme(plot.title = element_text(size = 11, hjust=0.5),
        axis.title.y = element_text(size=11),
        plot.caption = element_text(hjust=0.5, size=rel(1.2)),
        legend.position = 'bottom',
        legend.title = element_blank()) +
  labs(x = "",
       y ="",
       caption = "μ = 0.7") + 
  scale_color_manual(breaks = c('FFR',
                                'Shadow',
                                'Treasury'),
                     values = c('dodgerblue',
                                'firebrick',
                                'limegreen'))
)

```
```{=latex}
\caption{Policy impact calculated with different rates \label{fig:shockmes}}
\bigskip
\footnotesize
\end{figure}
```

As seen in Figure \ref{fig:shockmes}, the difference between responses obtained when the policy shock is calculated using the FFR and the 1-Year Treasury Yield show little difference, however accounting for implied rates via the Shadow Rate has considerable impact on the results. With the threshold level set to the median, both regimes responses shift downwards, and with the 70th percentile, this is only seen in high uncertainty responses. Such results do support the finding of shocks in high uncertainty times having a larger contractionary effect due to uncertainty causing an increase in volatility.

Voiding the effects of financial stress is also a crucial determinant in shaping the results. As seen in Figure \ref{fig:finstress}, after including the financial stress indicator as a control variable, the responses shift downwards drastically, especially in the lower regime. Switching the threshold variable from the median to the 70th percentile, the low uncertainty responses turn almost unaffected, however high regime responses show a drastically increased drop.

```{=latex}
\begin{figure}[!htbp]
```
```{r finstress}
cowplot::plot_grid(
irfs %>% 
  select(-mult, -restr) %>% 
  filter(label != 'R') %>% 
  filter(spec == 'shadow',
         thresh == '0.5',
         index == 'epu',
         control %in% c('none', 'financial_stress')) %>% 
  mutate(control = case_when(control == 'none' ~ 'No Controls',
                             control == 'financial_stress' ~ 'Fin. Stress')) %>% 
  ggplot(aes(x = t, y = median, color = control
  )) +
  geom_line(linewidth = lw) +
  geom_hline(yintercept = 0, color = "red") +
  facet_wrap(regime~label, scales = "free")+
  theme_minimal() +
  theme(plot.title = element_text(size = 11, hjust=0.5),
        axis.title.y = element_text(size=11),
        plot.caption = element_text(hjust=0.5, size=rel(1.2)),
        legend.position = 'bottom',
        legend.title = element_blank()) +
  scale_color_manual(breaks = c('No Controls',
                                'Fin. Stress'),
                     values = RColorBrewer::brewer.pal(3, 'Set1')[1:2]) +
  labs(x = "",
       y ="",
       caption = "μ = 0.5"),
irfs %>% 
  select(-mult, -restr) %>% 
  filter(label != 'R') %>% 
  filter(spec == 'shadow',
         thresh == '0.7',
         index == 'epu',
         control %in% c('none', 'financial_stress')) %>% 
  mutate(control = case_when(control == 'none' ~ 'No Controls',
                             control == 'financial_stress' ~ 'Fin. Stress')) %>% 
  ggplot(aes(x = t, y = median, color = control
  )) +
  geom_line(linewidth = lw) +
  geom_hline(yintercept = 0, color = "red") +
  facet_wrap(regime~label, scales = "free")+
  theme_minimal() +
  theme(plot.title = element_text(size = 11, hjust=0.5),
        axis.title.y = element_text(size=11),
        plot.caption = element_text(hjust=0.5, size=rel(1.2)),
        legend.position = 'bottom',
        legend.title = element_blank()) +
  scale_color_manual(breaks = c('No Controls',
                                'Fin. Stress'),
                     values = RColorBrewer::brewer.pal(3, 'Set1')[1:2]) +
  labs(x = "",
       y ="",
       caption = "μ = 0.7")
)

```
```{=latex}
\caption{Financial stress versus no controls \label{fig:finstress}}
\bigskip
\footnotesize
\textit{Notes}: In both the No Controls and the Fin. Stress estimations, the Shadow Rate is used to estimate policy shocks.
\end{figure}
```

### 2023-10-15 US/codes/data.R

@@ -1,0 +1,246 @@
+data <- function(){
+#loading packages
+#packages
+require(fredr)
+require(tidyverse)
+require(rlang)
+require(lubridate)
+require(readxl)
+  
+  x13adj <- function(value, freq, date){
+    require(forecast)
+    require(seasonal)
+    require(tidyverse)
+    tryCatch(
+      value %>% 
+        ts(frequency = freq,
+           start = c(date %>% min %>% year, 
+                     date %>% min %>% month)) %>% 
+        seas() %>% 
+        seasadj(),
+      error = function(error){value}
+    )
+  }
+  
+fredr_set_key("cda47ae66b38ed7988c0a9c2ec80c94f")
+
+
+macro_codes <- c(
+  'DFF', #FFR
+  'USACPIALLMINMEI', #CPI
+  'GDPC1' #GDP
+  )
+
+unc_codes <- c(
+  'USEPUINDXD', #EPU
+  'EPUMONETARY', #EPU monetary
+  'EPUFISCAL', #EPU fiscal
+  'EPUTRADE', #EPU trade
+  'EPUFINREG', #EPU financial regulation
+  'EPUSOVDEBT', #EPU debt
+  'EPUREG' #EPU regulation
+)
+
+macro_codes_alt <- c(
+  'DGS1', #1 year rate
+  'CORESTICKM159SFRBATL', #Sticky CPI (less food and energy)
+  'PCEPI' #PCE
+)
+
+rob_codes <- c(
+  'TEDRATE', #TED spread
+  'WLEMUINDXD', #Equity market uncertainty
+  'VIXCLS', #VIX index
+  'WUIUSA' #WUI
+)
+
+fred_query <- function(ids, freq, start){
+  
+  require(fredr)
+  require(dplyr)
+  require(purrr)
+  
+  #download data
+  params <- list(
+    series_id = ids,
+    frequency = freq,
+    observation_start = as.Date(start)
+  )
+  
+  
+  data  <- pmap_dfr(
+    .l = params,
+    .f = ~ fredr(series_id = .x, frequency = .y) ) %>%
+    dplyr::select(date, series_id, value) %>%
+    spread(key = series_id, value = value) %>%
+    drop_na() 
+  
+  data
+}
+
+macro <- fred_query(
+  ids = macro_codes,
+  freq = 'q',
+  start = '1950-01-01') %>% 
+  rename(r = DFF,
[…]

### 2023-10-15 US/codes/data2.R

@@ -1,0 +1,230 @@
+data <- function(){
+  #loading packages
+  #packages
+  require(fredr)
+  require(tidyverse)
+  require(rlang)
+  require(lubridate)
+  require(readxl)
+  
+  fredr_set_key("cda47ae66b38ed7988c0a9c2ec80c94f")
+  
+  
+  macro_codes <- c(
+    'DFF', #FFR
+    'USACPIALLMINMEI', #CPI
+    'GDPC1' #GDP
+  )
+  
+  unc_codes <- c(
+    'USEPUINDXD', #EPU
+    'EPUMONETARY', #EPU monetary
+    'EPUFISCAL', #EPU fiscal
+    'EPUTRADE', #EPU trade
+    'EPUFINREG', #EPU financial regulation
+    'EPUSOVDEBT', #EPU debt
+    'EPUREG' #EPU regulation
+  )
+  
+  macro_codes_alt <- c(
+    'DGS1', #1 year rate
+    'CORESTICKM159SFRBATL', #Sticky CPI (less food and energy)
+    'PCEPI' #PCE
+  )
+  
+  rob_codes <- c(
+    'TEDRATE', #TED spread
+    'WLEMUINDXD', #Equity market uncertainty
+    'VIXCLS', #VIX index
+    'WUIUSA' #WUI
+  )
+  
+  fred_query <- function(ids, freq, start){
+    
+    require(fredr)
+    require(dplyr)
+    require(purrr)
+    
+    #download data
+    params <- list(
+      series_id = ids,
+      frequency = freq,
+      observation_start = as.Date(start)
+    )
+    
+    
+    data  <- pmap_dfr(
+      .l = params,
+      .f = ~ fredr(series_id = .x, frequency = .y) ) %>%
+      dplyr::select(date, series_id, value) %>%
+      spread(key = series_id, value = value) %>%
+      drop_na() 
+    
+    data
+  }
+  
+  macro <- fred_query(
+    ids = macro_codes,
+    freq = 'q',
+    start = '1950-01-01') %>% 
+    rename(r = DFF,
+           pi = USACPIALLMINMEI,
+           y = GDPC1) %>% 
+    select(date, r, pi, y) %>% 
+    mutate(date = as_date(date))
+  
+  
+  macro <- macro %>% 
+    mutate(
+      year = year(date)) %>% 
+    group_by(year) %>% 
+    mutate(mean_y = mean(y),
[…]

### 2023-10-15 US/codes/dataprep.R

@@ -1,0 +1,159 @@
+prep <- function(){
+library(fredr)
+library(tidyverse)
+library(rlang)
+library(lubridate)
+source('US_only/codes/data.R')
+source('US_only/codes/functions.R')
+
+data[['macro']] <- data[['macro']] %>% 
+  mutate(pi = pi - lag(pi),
+         y = y - lag(y)) %>% 
+  drop_na()
+
+data[['uncertainty']] <- data[['uncertainty']] %>% 
+  full_join(data[['forrob']])
+
+data[['forrob']] <- NULL
+
+prep1 <- function(data, index, threshold, controls = NA){
+#Define specs
+data_prepped <- NULL
+
+#Main 
+data_prepped[['main']] <- 
+  list(
+    data = 
+      data[['macro']] %>% 
+      inner_join(create_indicator2(data[['uncertainty']], index, threshold), by = "date") %>% 
+      select(-indicator),
+    indicator = 
+      create_indicator2(data[['uncertainty']], index, threshold) %>% 
+      inner_join(data[['macro']], by = "date") %>% 
+      select(date, indicator),
+    controls = controls
+  )
+
+#Lag
+data_prepped[['lag']] <- 
+  list(
+    data = 
+      data[['macro']] %>% 
+      inner_join(create_indicator2(data[['uncertainty']], index, threshold), by = "date") %>% 
+      select(-indicator),
+    indicator = 
+      create_indicator2(data[['uncertainty']], index, threshold) %>% 
+      inner_join(data[['macro']], by = "date") %>% 
+      mutate(indicator = lag(indicator)) %>% 
+      select(date, indicator),
+    controls = controls
+  )
+
+#1y interest rate swap
+data_prepped[['r1y']] <- 
+  list(
+    data = 
+      data[['macro']] %>% 
+      select(-r) %>% 
+      left_join(data[['macro_alt']] %>% 
+                  select(date, r = r1y)) %>% 
+      select(date, r, pi, y) %>% 
+      inner_join(create_indicator2(data[['uncertainty']], index, threshold), by = "date") %>% 
+      select(-indicator),
+    indicator = 
+      create_indicator2(data[['uncertainty']], index, threshold) %>% 
+      inner_join(data[['macro']], by = "date") %>% 
+      select(date, indicator),
+    controls = controls
+  )
+
+#shadow rate swap
[…]

### 2023-10-15 US/codes/estout_cleanup.R

@@ -1,0 +1,126 @@
+library(tidyverse)
+library(rlang)
+library(lubridate)
+
+estout <- read_rds('US_only/data/estout.Rds')
+
+catch <- 'start'
+for(m in 1:length(restrs)){
+  restr <- restrs[m]
+  for(n in 1:length(indices)){
+    index <- indices[n]
+    for(k in 1:length(thresholds)){
+      thresh <- thresholds[k]
+      for(j in 1: length(controlstruct)){
+        cont <- controlstruct[j]
+        for(i in 1:length(specs)){
+          spec <- specs[i]
+          valid <- estout[[restr]][[index]][[thresh]][[cont]][[spec]][[2]] %>% is.data.frame()
+          if(valid){
+            
+          } else {
+            estout[[restr]][[index]][[thresh]][[cont]][[spec]] <- NULL
+            catch <- c(catch, paste(restr, index, thresh, cont, spec, sep = '-'))
+          }
+        }
+        len <- (length(estout[[restr]][[index]][[thresh]][[cont]]) > 0)
+        if(len){ } else {
+          estout[[restr]][[index]][[thresh]][[cont]] <- NULL
+        }
+      }
+      len <- (length(estout[[restr]][[index]][[thresh]]) > 0)
+      if(len){ } else {
+        estout[[restr]][[index]][[thresh]] <- NULL
+      }
+    }
+    len <- (length(estout[[restr]][[index]]) > 0)
+    if(len){ } else {
+      estout[[restr]][[index]] <- NULL
+    }
+  }
+  len <- (length(estout[[restr]]) > 0)
+  if(len){ } else {
+    estout[[restr]] <- NULL
+  }
+}
+
+
+catch <- catch[2:length(catch)] %>% 
+  as_tibble() %>% 
+  separate(value,
+           into = c('restr', 'index', 'threshold', 'control', 'spec'),
+           sep = '-') %>% 
+  select(-restr)
+
+
+namesform <- names(estout)
+for(m in 1:length(namesform)){
+  restr <- namesform[m]
+  
+  irf_out3 <- NULL
+  res_out3 <- NULL
+  namesforn <- names(estout[[restr]])
+  for(n in 1:length(namesforn)){
+    index <- namesforn[n]
+    
+    irf_out2 <- NULL
+    res_out2 <- NULL
+    namesfork <- names(estout[[restr]][[index]])
+    for(k in 1:length(namesfork)){
+      thresh <- namesfork[k]
+      
[…]

### 2023-10-15 US/codes/functions.R

@@ -1,0 +1,233 @@
+create_indicator2 <- function(data, variable, prob){
+  
+  require(dplyr)
+  require(rlang)
+  
+  data <- data %>%
+    mutate(q = quantile(!!sym(variable), probs = prob, na.rm = TRUE), 
+           indicator = ifelse(!!sym(variable) > q, 1, 0)) %>% 
+    select(date, indicator)
+  
+  data
+}
+
+create_indicator3 <- function(data, variable, upper, lower){
+  
+  require(dplyr)
+  require(rlang)
+  
+  data <- data %>%
+    mutate(q_up = quantile(!!sym(variable), probs = upper, na.rm = TRUE), 
+           q_dn = quantile(!!sym(variable), probs = lower, na.rm = TRUE),
+           indicator_up = ifelse(!!sym(variable) > q_up, 1, 0),
+           indicator_dn = ifelse(!!sym(variable) < q_dn, 1, 0),
+           indicator_mid = ifelse(indicator_up == 0 & indicator_dn == 0, 1, 0)) %>% 
+    select(date, indicator_up, indicator_mid, indicator_dn)
+  
+  data
+}
+
+sign_tvar_2regime <- function(data, indicator, constr_mat, seed,
+                              nlags, draws, subdraws, nkeep, KMIN, KMAX, constant, steps){
+  
+  require(dplyr)
+  paste('US_only/codes/varsignr/', list.files('US_only/codes/varsignr') , sep = '') %>% lapply(source)
+  
+  
+  indicator_inv <- indicator %>% mutate(indicator = ifelse(indicator == 1, 0, 1)) %>% as.matrix()
+  indicator <- as.matrix(indicator)
+  data_dn <- data[,1:(ncol(data))] %>% as.matrix()
+  data_up <- data_dn
+  
+  
+  for(i in 1:ncol(data_dn)){
+    colnames(data_dn)[i] <- paste(colnames(data_dn)[i], "dn", sep = "_")
+    colnames(data_up)[i] <- paste(colnames(data_up)[i], "up", sep = "_")
+    data_dn[,i] <- data_dn[,i]*indicator_inv
+    data_up[,i] <- data_up[,i]*indicator
+  }
+  
+  data_ts <- bind_cols(data_up, data_dn) %>% ts()
+  
+  set.seed(seed)
+  constr_up <- constr_mat[1,]
[…]

### 2023-10-15 US/codes/main.R

@@ -1,0 +1,167 @@
+library(fredr)
+library(tidyverse)
+library(rlang)
+library(lubridate)
+library(foreach)
+library(doParallel)
+library(doSNOW)
+
+source('US_only/codes/dataprep.R')
+source('US_only/codes/functions.R')
+clean <- function(estdata){
+  estdata$data <- estdata$data %>% 
+    drop_na()
+  estdata$indicator <- estdata$indicator %>% 
+    drop_na()
+  estdata$controls <- if(!is.logical(estdata$controls)){
+    estdata$controls %>% 
+      drop_na()} else{
+        NA}
+  
+
+  
+  mindate <- max(
+    min(estdata$data$date),
+    min(estdata$indicator$date),
+    if(!is.logical(estdata$controls)){
+      min(estdata$controls$date) 
+    } else{
+      NA
+    },
+    na.rm = T
+  )
+  
+  maxdate <- min(
+    max(estdata$data$date),
+    max(estdata$indicator$date),
+    if(!is.logical(estdata$controls)){
+      max(estdata$controls$date) 
+    } else{
+      NA
+    },
+    na.rm = T
+  )
+  
+  estdata$data <- estdata$data %>% 
+    filter(date >= mindate,
+           date <= maxdate) 
+  estdata$indicator <- estdata$indicator %>% 
+    filter(date >= mindate,
+           date <= maxdate) 
+  estdata$controls <- if(!is.logical(estdata$controls)){
+    estdata$controls %>% 
+      filter(date >= mindate,
+             date <= maxdate)
+  } else {
+    NA
+  }
+  
+  estdata
+}
+
+data2
+
+#Sign restrictions - just follow Uhlig paper with 1-6
+restr <- list(full = list(KMIN = 1,
+                          KMAX = 6)#,
+              #cont = list(KMIN = 1,
+              #            KMAX = 1),
+              #shape = list(KMIN = 2,
+              #             KMAX = 6)
+              )
+
+restrs <- names(restr)
+indices <- names(data)
+#Indices - Focus on EPU only
+indices <- indices[str_detect(indices, 'epu')]
+thresholds <- names(data[[1]])
+controlstruct <- names(data[[1]][[1]])
+specs <- names(data[[1]][[1]][[1]])
+
+ncores <- parallel::detectCores(logical = F)-1
+cl <- makeCluster(ncores)
[…]

### 2023-10-15 US/codes/plotting.R

@@ -1,0 +1,60 @@
+estout_clean <- read_rds('US_only/data/estout_clean.Rds')
+
+irfs <- estout_clean[['irfs']]
+
+
+estout_clean$invalids %>% 
+  mutate(index = case_when(index == 'epu_debt' ~ 'EPU: Debt, crises',
+                           index == 'epu_finreg' ~ 'EPU: Fin. Regulation',
+                           index == 'epu_fis' ~ 'EPU: Fiscal Policy',
+                           index == 'epu_mon' ~ 'EPU: Monetary Policy',
+                           index == 'epu_reg' ~ 'EPU: Regulation',
+                           index == 'epu_trade' ~ 'EPU: Trade Policy'),
+         threshold = case_when(threshold == '0.5' ~ 'Median',
+                               threshold == '0.7' ~ '70th Perc.'),
+         control = case_when(control == 'none' ~ 'No controls',
+                             control == 'uncertainty' ~ 'Uncertainty',
+                             control == 'financial_stress' ~ 'Financial Stress',
+                             control == 'both' ~ 'Uncertainty and Financial Stress'),
+         spec = case_when(spec == 'main' ~ 'Interest rate measured by FFR',
+                          spec == 'lag' ~ 'Regime indicator from lagged level of uncertainty',
+                          spec == 'r1y'  ~ 'Interest rate measured by')) %>% 
+  janitor::tabyl(spec)
+
+lw = .75
+
+
+
+
+  
+
+
+irfs %>% 
+  select(-mult, -restr) %>% 
+  filter(label != 'R',
+         index != 'epu') %>% 
+  filter(spec == 'shadow',
+         control == 'none',
+         thresh == '0.5') %>% 
+  mutate(index = case_when(index == 'epu_debt' ~ 'EPU: Debt, crises',
+                           index == 'epu_finreg' ~ 'EPU: Fin. Regulation',
+                           index == 'epu_fis' ~ 'EPU: Fiscal Policy',
+                           index == 'epu_mon' ~ 'EPU: Monetary Policy',
+                           index == 'epu_reg' ~ 'EPU: Regulation',
+                           index == 'epu_trade' ~ 'EPU: Trade Policy')) %>% 
+  ggplot(aes(x = t, y = median, color = regime
+  )) +
+  geom_line(linewidth = lw) +
+  geom_hline(yintercept = 0, color = "red") +
[…]

### 2023-10-15 US/codes_clean.R

@@ -1,0 +1,784 @@
+#loading packages
+#packages
+library(fredr)
+library(tidyverse)
+library(gridExtra)
+library(VARsignR)
+library(rlang)
+library(lubridate)
+
+#Custom functions
+get_data_macro <- function(key){
+  
+  require(fredr)
+  require(dplyr)
+  require(purrr)
+  
+#set api key
+fredr_set_key(key)
+
+#download data
+params <- list(
+  series_id = c("DFF", "USACPIALLMINMEI", "INDPRO"),
+  frequency = "m",
+  observation_start = as.Date("1950-01-01")
+)
+
+
+data_macro  <- pmap_dfr(
+  .l = params,
+  .f = ~ fredr(series_id = .x, frequency = .y)
+) %>%
+  dplyr::select(date, series_id, value) %>%
+  spread(key = series_id, value = value) %>%
+  drop_na() %>% rename(ffr = DFF,
+                       cpi = USACPIALLMINMEI,
+                       indpro = INDPRO
+  ) 
+
+data_macro
+}
+
+get_data_unc <- function(key){
+  
+  require(fredr)
+  require(dplyr)
+  require(purrr)
+  
+  #set api key
+  fredr_set_key(key)
+  
+  #download data
+  params <- list(
+    series_id = c("USEPUINDXD", "TEDRATE", "VIXCLS"),
+    frequency = "m",
+    observation_start = as.Date("1950-01-01")
+  )
+  
+  
+  data_unc  <- pmap_dfr(
+    .l = params,
+    .f = ~ fredr(series_id = .x, frequency = .y)
+  ) %>%
+    select(date, series_id, value) %>%
+    spread(key = series_id, value = value) %>%
+    rename(epu = USEPUINDXD,
+           ted = TEDRATE,
+           vix = VIXCLS
+    ) 
+  
+  data_unc
+}
+
+cortab_create <- function(data, digits){
+  cortab <- data %>% select(-date) %>% drop_na() %>% cor()
+  cortab <- round(cortab, digits = digits)
+  cortab <- cortab*lower.tri(cortab)
+  diag(cortab) <- 1
+  for(i in 1:nrow(cortab)) {
+    for(j in 1:ncol(cortab)){
+      if(cortab[i,j] == 0) {
+        cortab[i,j] <- " "
+      }
+    }
+  }
+  cortab
+  rownames(cortab) <- c("MPU", "TED", "EPU", "VIX")
[…]

### 2023-10-15 US/depo/codes.R

@@ -1,0 +1,500 @@
+#loading packages
+#packages
+library(fredr)
+library(tidyverse)
+library(gridExtra)
+library(fredr)
+library(VARsignR)
+library(rlang)
+library(lubridate)
+
+#set api key
+fredr_set_key("cda47ae66b38ed7988c0a9c2ec80c94f")
+
+#download data
+params <- list(
+  series_id = c("USEPUINDXD", "DFF", "USACPIALLMINMEI", "INDPRO"),
+  frequency = "m",
+  observation_start = as.Date("1950-01-01")
+)
+
+
+import  <- pmap_dfr(
+  .l = params,
+  .f = ~ fredr(series_id = .x, frequency = .y)
+) %>%
+  select(date, series_id, value) %>%
+  spread(key = series_id, value = value) %>%
+  drop_na() %>% rename(epu = USEPUINDXD,
+                       ffr = DFF,
+                       cpi = USACPIALLMINMEI,
+                       indpro = INDPRO
+  ) 
+
+#select data range
+data <- import %>%
+  select(date, epu, ffr, cpi, indpro) %>%
+  drop_na() %>% filter(date <= as.Date("2019-12-01"))
+
+
+
+data %>%
+  gather(key = "variable", value = "value", ffr, cpi, indpro) %>%
+  mutate(variable = case_when(variable == "ffr" ~ "Federal Funds Rate",
+                              variable == "cpi" ~ "Consumer Prices",
+                              variable == "indpro" ~ "Industrial Production Index"),
+         variable = factor(variable, levels = c("Federal Funds Rate", "Consumer Prices", "Industrial Production Index"))) %>%
+  ggplot(aes(x = date, y = value)) +
+  geom_line() +
+  facet_wrap(~variable, scales = "free", nrow = 3) + theme_minimal() +
+  labs(x = "",
+       y = "") 
+
+#generate first differences / growth rates  
+data <- data %>%
+  mutate(cpi = cpi - lag(cpi),
+         indpro = indpro - lag(indpro)) %>% drop_na()
+
+
+#creating threshold values
+data <- data %>%
+  mutate(epu50 = median(epu),
+         epu75 = quantile(epu, probs = 0.75),
+         epu25 = quantile(epu, probs = 0.25))
+
+#plot the threshold values
+data %>% 
+  select(date, starts_with("epu")) %>% 
+  gather(key = "variable", value = "value", starts_with("epu")) %>% 
+  mutate(variable = case_when(variable == "epu" ~ "EPU",
[…]

### 2023-10-15 US/depo/codes_clean_2023_quarterly.R

@@ -1,0 +1,634 @@
+#loading packages
+#packages
+library(fredr)
+library(tidyverse)
+library(gridExtra)
+library(VARsignR)
+library(rlang)
+library(lubridate)
+
+#Custom functions
+get_data_macro <- function(key){
+  
+  require(fredr)
+  require(dplyr)
+  require(purrr)
+  
+  #set api key
+  fredr_set_key(key)
+  
+  #download data
+  params <- list(
+    series_id = c("DFF", "USACPIALLMINMEI", "INDPRO"),
+    frequency = "q",
+    observation_start = as.Date("1950-01-01")
+  )
+  
+  
+  data_macro  <- pmap_dfr(
+    .l = params,
+    .f = ~ fredr(series_id = .x, frequency = .y)
+  ) %>%
+    dplyr::select(date, series_id, value) %>%
+    spread(key = series_id, value = value) %>%
+    drop_na() %>% rename(ffr = DFF,
+                         cpi = USACPIALLMINMEI,
+                         indpro = INDPRO
+    ) 
+  
+  data_macro
+}
+
+get_data_unc <- function(key){
+  
+  require(fredr)
+  require(dplyr)
+  require(purrr)
+  
+  #set api key
+  fredr_set_key(key)
+  
+  #download data
+  params <- list(
+    series_id = c("USEPUINDXD", "TEDRATE", "VIXCLS"),
+    frequency = "q",
+    observation_start = as.Date("1950-01-01")
+  )
+  
+  
+  data_unc  <- pmap_dfr(
+    .l = params,
+    .f = ~ fredr(series_id = .x, frequency = .y)
+  ) %>%
+    select(date, series_id, value) %>%
+    spread(key = series_id, value = value) %>%
+    rename(epu = USEPUINDXD,
+           ted = TEDRATE,
+           vix = VIXCLS
+    ) 
+  
+  data_unc
+}
+
+cortab_create <- function(data, digits){
+  cortab <- data %>% select(-date) %>% drop_na() %>% cor()
+  cortab <- round(cortab, digits = digits)
+  cortab <- cortab*lower.tri(cortab)
+  diag(cortab) <- 1
+  for(i in 1:nrow(cortab)) {
+    for(j in 1:ncol(cortab)){
+      if(cortab[i,j] == 0) {
+        cortab[i,j] <- " "
+      }
+    }
+  }
+  cortab
[…]

### 2023-10-15 US/depo/old_version/depo/usa.R

@@ -1,0 +1,169 @@
+#packages
+library(readxl)
+library(tidyverse)
+library(tseries)
+library(forecast)
+library(TSA)
+library(fpp2)
+library(lmtest)
+library(vars)
+library(mFilter)
+library(ggplot2)
+library(tsDyn)
+library(fredr)
+
+#set api key
+fredr_set_key("cda47ae66b38ed7988c0a9c2ec80c94f")
+
+#retrieve data from fred database
+ffr <- fredr(series_id = "DFF",
+             frequency = "m") %>%
+  dplyr::select(date, value) %>%
+  rename(ffr = value)
+
+cpi <- fredr(series_id = "USACPIALLMINMEI",
+             frequency = "m",
+             units = "chg") %>%
+  dplyr::select(date, value) %>%
+  rename(cpi = value)
+
+indpro <- fredr(series_id = "INDPRO",
+                frequency = "m",
+                units = "chg") %>%
+  dplyr::select(date, value) %>%
+  rename(indpro = value)
+
+r3m <- fredr(series_id = "DTB3",
+             frequency = "m") %>%
+  dplyr::select(date, value) %>%
+  rename(r3m = value)
+
+r10y <- fredr(series_id = "DGS10",
+              frequency = "m") %>%
+  dplyr::select(date, value) %>%
+  rename(r10y = value)
+
+m1 <- fredr(series_id = "M1REAL",
+            frequency = "m",
+            units = "chg") %>%
+  dplyr::select(date, value) %>%
+  rename(m1 = value)
+
+m2 <- fredr(series_id = "M2REAL",
+            frequency = "m",
+            units = "chg") %>%
+  dplyr::select(date, value) %>%
+  rename(m2 = value)
+
+depint <- fredr(series_id = "IR3TCD01USM156N",
+                frequency = "m") %>%
+  dplyr::select(date, value) %>%
+  fill(value, .direction = "downup")%>%
+  rename(depint = value)
+
+oil <- fredr(series_id = "WTISPLC",
+             frequency = "m",
+             units = "chg") %>%
+  dplyr::select(date, value) %>%
+  rename(oil = value)
+
+totres <- fredr(series_id = "TRESEGUSM052N",
+                frequency = "m",
+                units = "chg") %>%
[…]

### 2023-10-15 US/depo/old_version/uncerainty_lpirf.R

@@ -1,0 +1,85 @@
+#packages
+suppressPackageStartupMessages({
+  library(readxl)
+  library(tidyverse)
+  library(tseries)
+  library(forecast)
+  library(TSA)
+  library(fpp2)
+  library(lmtest)
+  library(vars)
+  library(mFilter)
+  library(ggplot2)
+  library(tsDyn)
+  library(fredr)
+  library(VARsignR)
+  library(lubridate)
+  library(cowplot)
+  library(gridExtra)
+  library(patchwork)
+  library(data.table)
+  library(tsibble)
+  library(mvnfast)
+  library(lpirfs)
+})
+
+folder <- "C:/Users/horva/Documents/Egyetem/PhD/Research Material/Uncertainty project/Új mappa/"
+
+
+#set api key
+fredr_set_key("cda47ae66b38ed7988c0a9c2ec80c94f")
+
+#download data
+params <- list(
+  series_id = c("USEPUINDXD", "DFF", "USACPIALLMINMEI", "INDPRO"),
+  frequency = "m",
+  observation_start = as.Date("1950-01-01")
+)
+
+
+import  <- pmap_dfr(
+  .l = params,
+  .f = ~ fredr(series_id = .x, frequency = .y)
+) %>%
+  dplyr::select(date, series_id, value) %>%
+  spread(key = series_id, value = value) %>%
+  drop_na() %>% as_tsibble() %>% rename(epu = USEPUINDXD,
+                                        ffr = DFF,
+                                        cpi = USACPIALLMINMEI,
+                                        indpro = INDPRO
+  ) 
+
+#select data range
+data <- import %>%
+  dplyr::select(date, epu, ffr, cpi, indpro) %>%
+  drop_na() %>% filter(date <= as.Date("2019-12-01")) %>%
+  mutate(cpi = cpi - lag(cpi),
+         indpro = indpro - lag(indpro)) %>% drop_na()
+
+
+
+switching_data <- ifelse(data$epu > median(data$epu), 1, 0)
+
+data <- data %>% as_tibble() %>% dplyr::select(-date, -epu)
+  
+
+results_lin <- lp_lin(endog_data = data,
+                      lags_endog_lin = 1,
+                      trend = 0,
+                      shock_type = 1,
+                      confint = 1.96,
+                      hor = 20)
+                      
+plot(results_lin)                      
+                      
+
+results_nl <- lp_nl(data,
+                    lags_endog_lin = 1, lags_endog_nl = 1,
[…]

### 2023-10-15 US/depo/old_version/usa2.R

@@ -1,0 +1,881 @@
+#packages
+library(readxl)
+library(tidyverse)
+library(tseries)
+library(forecast)
+library(TSA)
+library(fpp2)
+library(lmtest)
+library(vars)
+library(mFilter)
+library(ggplot2)
+library(tsDyn)
+library(fredr)
+library(VARsignR)
+library(lubridate)
+library(cowplot)
+library(gridExtra)
+library(patchwork)
+library(data.table)
+#https://cran.r-project.org/web/packages/VARsignR/vignettes/VARsignR-vignette.html
+
+folder <- "C:/Users/horva/Documents/Egyetem/PhD/Research Material/Uncertainty project/Új mappa/"
+
+#set api key
+fredr_set_key("cda47ae66b38ed7988c0a9c2ec80c94f")
+
+#retrieve data from fred database
+#####
+epu <- fredr(series_id = "USEPUINDXD",
+             frequency = "m") %>%
+  dplyr::select(date, value) %>%
+  rename(epu = value)
+
+ffr <- fredr(series_id = "DFF",
+             frequency = "m") %>%
+  dplyr::select(date, value) %>%
+  rename(ffr = value)
+
+cpi <- fredr(series_id = "USACPIALLMINMEI",
+             frequency = "m",
+             units = "chg") %>%
+  dplyr::select(date, value) %>%
+  rename(cpi = value)
+
+indpro <- fredr(series_id = "INDPRO",
+                frequency = "m",
+                units = "chg") %>%
+  dplyr::select(date, value) %>%
+  rename(indpro = value)
+
+vix <- fredr(series_id = "VIXCLS",
+             frequency = "m") %>%
+  dplyr::select(date, value) %>%
+  rename(vix = value)
+
+ted <- fredr(series_id = "TEDRATE",
+             frequency = "m") %>%
+  dplyr::select(date, value) %>%
+  rename(ted = value)
+#####
+
+
+#setup ts dataframe
+#####
+datalist <- list(epu, ffr, cpi, indpro)
+mindate <- vector()
+maxdate <- vector()
+
+for (i in 1:length(datalist)){
+  mindate[i] <- min(datalist[[i]]$date)
+  maxdate[i] <- max(datalist[[i]]$date)
+}
[…]

### 2023-10-15 US/depo/old_version/usa4.R

@@ -1,0 +1,326 @@
+#packages
+suppressPackageStartupMessages({
+  library(readxl)
+  library(tidyverse)
+  library(tseries)
+  library(forecast)
+  library(TSA)
+  library(fpp2)
+  library(lmtest)
+  library(vars)
+  library(mFilter)
+  library(ggplot2)
+  library(tsDyn)
+  library(fredr)
+  library(VARsignR)
+  library(lubridate)
+  library(cowplot)
+  library(gridExtra)
+  library(patchwork)
+  library(data.table)
+  library(tsibble)
+  library(mvnfast)
+})
+
+folder <- "C:/Users/horva/Documents/Egyetem/PhD/Research Material/Uncertainty project/?j mappa/"
+
+
+#set api key
+fredr_set_key("cda47ae66b38ed7988c0a9c2ec80c94f")
+
+#download data
+params <- list(
+  series_id = c("USEPUINDXD", "DFF", "USACPIALLMINMEI", "INDPRO"),
+  frequency = "m",
+  observation_start = as.Date("1950-01-01")
+)
+
+
+import  <- pmap_dfr(
+  .l = params,
+  .f = ~ fredr(series_id = .x, frequency = .y)
+) %>%
+  dplyr::select(date, series_id, value) %>%
+  spread(key = series_id, value = value) %>%
+  drop_na() %>% rename(epu = USEPUINDXD,
+                                        ffr = DFF,
+                                        cpi = USACPIALLMINMEI,
+                                        indpro = INDPRO
+                                        ) 
+
+#select data range
+data <- import %>%
+  dplyr::select(date, epu, ffr, cpi, indpro) %>%
+  drop_na() %>% filter(date <= as.Date("2019-12-01"))
+
+#plotting the data series
+data %>%
+  gather(key = "variable", value = "value", ffr, cpi, indpro) %>%
+  mutate(variable = case_when(variable == "ffr" ~ "Federal Funds Rate",
+                              variable == "cpi" ~ "Consumer Prices",
+                              variable == "indpro" ~ "Industrial Production Index"),
+         variable = factor(variable, levels = c("Federal Funds Rate", "Consumer Prices", "Industrial Production Index"))) %>%
+  ggplot(aes(x = date, y = value)) +
+  geom_line() +
+  facet_wrap(~variable, scales = "free", nrow = 3) + theme_minimal() +
+  labs(x = "",
+       y = "") 
+
[…]

### 2023-10-16 US/codes/data.R
Fixed code for folders Fixed inflation seasonal adjustment (code)

@@ -17,7 +17,8 @@ require(readxl)
            start = c(date %>% min %>% year, 
                      date %>% min %>% month)) %>% 
         seas() %>% 
-        seasadj(),
+        seasadj() %>% 
+        as.numeric(),
       error = function(error){value}
     )
   }
@@ -87,8 +88,7 @@ macro <- fred_query(
          y = GDPC1) %>% 
   select(date, r, pi, y) %>% 
   mutate(date = as_date(date)) %>% 
-  mutate(value = x13adj(pi, 4, date))
-
+  mutate(pi = x13adj(pi, 4, date))
 
 macro <- macro %>% 
   mutate(
@@ -125,7 +125,7 @@ macro_alt <- fred_query(
          pi_stick = CORESTICKM159SFRBATL,
          pi_pce = PCEPI) %>% 
   mutate(date = as_date(date)) %>% 
-  left_join(read_excel('US_only/data/shadowrate.xlsx') %>% 
+  left_join(read_excel('US/data/shadowrate.xlsx') %>% 
               drop_na() %>% 
               mutate(date = as_date(date),
                      year = year(date),
@@ -154,7 +154,7 @@ forrob <- fred_query(
          equnc = WLEMUINDXD,
          wui = WUIUSA) %>% 
   mutate(date = as_date(date)) %>% 
-  left_join(read_excel("US_only/data/data_gpr_export.xls", 
+  left_join(read_excel("US/data/data_gpr_export.xls", 
                        col_types = c("date", "numeric", "numeric", 
                                      "numeric", "numeric", "numeric", 
                                      "numeric", "numeric", "numeric", 
@@ -204,7 +204,7 @@ forrob <- fred_query(
               ungroup() %>%
               distinct(year, quarter, .keep_all = T) %>% 
               select(date, gpr)) %>% 
-  left_join(    read_excel("US_only/data/tpu_web_latest.xlsx", 
+  left_join(    read_excel("US/data/tpu_web_latest.xlsx", 
                            sheet = "TPU_MONTHLY") %>% 
                   mutate(date = as_date(DATE),
                          year = year(date),

### 2023-10-16 US/codes/dataprep.R

@@ -3,8 +3,8 @@ library(fredr)
 library(tidyverse)
 library(rlang)
 library(lubridate)
-source('US_only/codes/data.R')
-source('US_only/codes/functions.R')
+source('US/codes/data.R')
+source('US/codes/functions.R')
 
 data[['macro']] <- data[['macro']] %>% 
   mutate(pi = pi - lag(pi),

### 2023-10-16 US/codes/estout_cleanup.R

@@ -2,7 +2,7 @@ library(tidyverse)
 library(rlang)
 library(lubridate)
 
-estout <- read_rds('US_only/data/estout.Rds')
+estout <- read_rds('US/data/estout.Rds')
 
 catch <- 'start'
 for(m in 1:length(restrs)){
@@ -123,4 +123,4 @@ estout_clean <- list(irfs = irf_out,
                      innovs = res_out,
                      invalids = catch)
 
-saveRDS(estout_clean, 'US_only/data/estout_clean.Rds')
+saveRDS(estout_clean, 'US/data/estout_clean.Rds')

### 2023-10-16 US/codes/functions.R

@@ -31,7 +31,7 @@ sign_tvar_2regime <- function(data, indicator, constr_mat, seed,
                               nlags, draws, subdraws, nkeep, KMIN, KMAX, constant, steps){
   
   require(dplyr)
-  paste('US_only/codes/varsignr/', list.files('US_only/codes/varsignr') , sep = '') %>% lapply(source)
+  paste('US/codes/varsignr/', list.files('US/codes/varsignr') , sep = '') %>% lapply(source)
   
   
   indicator_inv <- indicator %>% mutate(indicator = ifelse(indicator == 1, 0, 1)) %>% as.matrix()
@@ -67,7 +67,7 @@ sign_tvar_2regime_controls <- function(data, indicator, controls = NA, constr_ma
                               nlags, draws, subdraws, nkeep, KMIN, KMAX, constant, steps){
   
   require(dplyr)
-  paste('US_only/codes/varsignr/', list.files('US_only/codes/varsignr') , sep = '') %>% lapply(source)
+  paste('US/codes/varsignr/', list.files('US/codes/varsignr') , sep = '') %>% lapply(source)
   
   
   indicator_inv <- indicator %>% mutate(indicator = ifelse(indicator == 1, 0, 1)) %>% as.matrix()
@@ -109,7 +109,7 @@ sign_tvar_3regime <- function(data, indicator, constr_mat, seed,
                               nlags, draws, subdraws, nkeep, KMIN, KMAX, constant, steps){
   
   require(dplyr)
-  paste('IMF paper/codes/varsignr/', list.files('IMF paper/codes/varsignr') , sep = '') %>% lapply(source)
+  paste('US/codes/varsignr/', list.files('US/codes/varsignr') , sep = '') %>% lapply(source)
   
   
   indicator_up <- indicator[,1] %>% as.matrix()

### 2023-10-16 US/codes/main.R

@@ -6,8 +6,8 @@ library(foreach)
 library(doParallel)
 library(doSNOW)
 
-source('US_only/codes/dataprep.R')
-source('US_only/codes/functions.R')
+source('US/codes/dataprep.R')
+source('US/codes/functions.R')
 clean <- function(estdata){
   estdata$data <- estdata$data %>% 
     drop_na()
@@ -59,7 +59,6 @@ clean <- function(estdata){
   estdata
 }
 
-data2
 
 #Sign restrictions - just follow Uhlig paper with 1-6
 restr <- list(full = list(KMIN = 1,
@@ -162,6 +161,6 @@ estout[[m]] <- outs3
 names(estout) <- restrs
 stopCluster(cl)
 
-saveRDS(estout, 'US_only/data/estout.Rds')
+saveRDS(estout, 'US/data/estout.Rds')
 
-source('US_only/codes/estout_cleanup.R')
+source('US/codes/estout_cleanup.R')

### 2023-10-17 US/draft.Rmd
Dealt with seasonal adjustment issues

@@
   filter(spec == 'shadow',
-         control == 'financial_stress',
+         control == 'both',
          index == 'epu',
          thresh == '0.7') %>% 
-  ggplot(aes(x = t, y = median)) +
+  ggplot(aes(x = t, y = median, ymin = lower, ymax = upper)) +
   geom_line(linewidth = lw) +
   geom_hline(yintercept = 0, color = "red") +
+  geom_ribbon(alpha = .2)+
   facet_wrap(regime~label, scales = "free")+
@@
          index != 'epu',
-         control == 'financial_stress') %>% 
+         control == 'both') %>% 
   group_by(label, regime, spec, control, thresh, index) %>% 

### 2023-10-17 US/codes/functions.R
Fixed parametrization

@@ -94,7 +94,7 @@ sign_tvar_2regime_controls <- function(data, indicator, controls = NA, constr_ma
   set.seed(seed)
   constr_up <- constr_mat[1,]
   model_up <- rwz.reject(Y=data_ts, nlags=nlags, draws=draws, subdraws=subdraws, nkeep=nkeep, KMIN=KMIN,
-                         KMAX=6, constrained=constr_up, constant=constant, steps=steps)
+                         KMAX=KMAX, constrained=constr_up, constant=constant, steps=steps)
   
   set.seed(seed)
   constr_dn <- constr_mat[2,]+ncol(constr_mat)*constr_mat[2,]/abs(constr_mat[2,])

### 2023-10-17 US/draft.Rmd
Updates: Endogeneity

--- 2023-10-15 US/draft.Rmd
@@
 
+Results from a TVAR model estimated with a full set of controls indicates that 
 \textcolor{red}{WRITE SOMETHING ABOUT THE RESULTS HERE 1-2 PARAGRAPHS.}
@@
 ```
+
 ```{=latex}
@@
 ```
-
 In an alternative specifications the FFR is replaced with the 1-Year Treasury Yield.
@@
 ```
-
 In @baker2016measuring, along with the main EPU index, several EPU Categorical indices are also constructed.
@@
 
-It is important to mention that there are a number of alternative^[These measures can be seen in Figure \ref{fig:unces} of the Appendix.] measurements of uncertainty in the literature.
+It is important to mention that there are a number of alternative[^4] measurements of uncertainty in the literature.
 These indices are for example the Equity Market-related Economic Uncertainty Index of @baker2016measuring; the Geopolitical Risk Index and the Trade Policy Uncertainty Index of @caldara2022measuring and @caldara2020economic; the World Uncertainty Index of @ahir2022world; as well as the VIX Index and the TED spread.
 In this paper, only the EPU and its Categorical indices are considered as measures of uncertainty.
+
+[^4]: These measures can be seen in Figure \ref{fig:unces} of the Appendix.
 
@@
 For the sake of comparability, I adopt the same baseline setup as in @horvathTBAuncertain, however, in alternative specifications, some other factors are considered.
-Thus the baseline model can bet written as:
-\begin{equation} \label{eq:1}
+Thus the baseline model can bet written as: \begin{equation} \label{eq:1}
   Y_{t} = \sum\limits_{k=1}^{4}\Theta_{high}I(X_{t} \ge \mu)Y_{t-k} + \sum\limits_{k=1}^{4}\Theta_{low}I(X_{t} < \mu)Y_{t-k} + \epsilon_{t}
@@
 Secondly, I check if the results hold if I change the construction of the regime indicator from $X_{t}$ to $X_{t-1}$, making it pre-determined, and thereby eliminating the potential of estimating the impact of two contemporaneous shocks.
-Thirdly, I incorporate the uncertainty index in its levels as well as a financial stress indicator in order to void their effects from the aggregates during the estimation^[Please refer to the [Alternative model specifications] of the Appendix for further details on how Equation \ref{eq:1} above changes].
+Thirdly, I incorporate the uncertainty index in its levels as well as a financial stress indicator in order to void their effects from the aggregates during the estimation[^5].
+
+[^5]: Please refer to the [Alternative model specifications] of the Appendix for further details on how Equation \ref{eq:1} above changes
 
@@
   geom_hline(yintercept = 0, color = "red") +
-  geom_ribbon(alpha=.1, linetype=0) + 
+  geom_ribbon(alpha=.2, linetype=0) + 
   facet_wrap(regime~label, scales = "free")+
@@
 ```
-
 There are some noteworthy differences from these results compared to previous literature.
@@
 Alongside this, disinflation caused by the monetary shock is significantly enhanced in the high uncertainty regime compared to the low uncertainty one.
+Comparing the median and the 70th percentile as threshold levels for the regime switches, the responses of inflation in the lower regime and of output in the higher regime are not too dissimilar, however their opposite uncertainty regime counterparts show some key differences.
 
-Comparing the median and the 70th percentile as threshold levels for the regime switches, the responses of inflation in the lower regime and of output in the higher regime are not too dissimilar, however their opposite uncertainty regime counterparts show some key differences.
 Firstly, the output contraction in the low regime is elevated when switching to the higher uncertainty threshold.
@@
 
-\textcolor{red}{EZT AMÚGY MAJD MÁSHOVA RAKNÁM.}^[To further clarify, global uncertainty measures are likely better at capturing one of the mechanisms, through which geoeconomic fragmentation impacts economies, as noted in @ilyna2023geo.
-While here and and in the cited previous literature, country-specific uncertainty measures carry a different implication, whereby policy uncertainty explicitly reduces the efficacy of the transmission of policy actions.]
+Why might these results differ so drastically from previous works?
+Firstly, in @horvathTBAuncertain globally aggregated uncertainty measures are used.
+As noted in @ilyna2023geo, heightened uncertainty is a channels of geoeconomic fragmentation, and uncertainty measured on the global level is likely a better proxy of this, thereby having somewhat different implications.
+Additionally, @aastveit2017economic and @aquino2022does use a wider range of variables, which can help isolate how uncertainty interacts with the transmission mechanism.
+
+## Endogeneity
+
+Studying the relationship between policy shocks and (policy) uncertainty naturally raises the question of whether or not the uncertainty index is endogenous to policy innovations.
+Investigating the potential endogeneity can also guide in selecting the appropriate model specification.
+Testing for endogeneity is done by running Granger causality tests between the EPU index and the posterior median of the residuals obtained from alternative specifications considered.
+The results from a select number of these is shown here in Table \ref{tab:grang}^[The Linear model is added as a point of comparison for the TVAR models.
+It is estimated on the FFR, CPI inflation rate and GDP growth rate without further control variables.
+Full control TVAR represents models estimated with lagged values of the EPU index and the St.
+Louis FED Financial Stress Indicator added as controls.].
+
+```{=latex}
+\begin{center}
+`r xtable::xtable(read_rds('data/granger.rds'), type = "latex", digits = 3, align = c("p{0.1cm}", "p{9cm}", "p{2cm}", "p{2cm}"),label = 'tab:grang', caption = 'Granger causality tests with alternative model specifications',
+display = c('g', 'g', 'g', 'g'))`
+\end{center}
+```
+
+The results of these tests suggests that in most cases we fail to reject the null, and thus policy shocks do not drive uncertainty, neither the other way around.
+Interpreting the p-values as the likelihood of $H0$ to be true, there are a handful of conclusions to draw.
+The likelihood of policy shocks driving uncertainty stay reasonably high irrespective of specification, however, for the reverse causality case, it varies substantially.
+Firstly, the likelihood of uncertainty being a driver of policy shocks considerably increases when switching from the median to the 70th percentile threshold.
+This is likely due to the low number of observations in the regime.
+Secondly, using the Shadow rate^[Why the Shadow rate might be a more desirable measure is further discussed in the Robustness checks of the Appendix.] over the FFR also considerably reduces this likelihood.
+Given these, the superior specification should be a TVAR with a full set of controls estimated with the threshold level set to the median.
+The next subsection showcases the impulse responses estimated from this.
 
@@
 
-Through conducting a number of robustness tests, controlling for the level of uncertainty, using a more traditional TVAR model where the indicator is constructed from $t-1$ values of the threshold variable, or using 1-Year rates instead of the FFR has no significant impact on the results.
-However, controlling for the Zero-Lower-Bound, as well as financial stress both change the results significantly.
+Through conducting a number of robustness^[Discussed in further detail in the Appendix.] and endogeneity tests, we arrive at a model that should grasp best how uncertainty influences the transmission of policy shocks.
+Figure \ref{fig:updatedirfs} below shows the impulse responses obtained with the new specification after controlling for the level of uncertainty and financial stress.
 
@@
   filter(label != 'R') %>% 
-  filter(spec == 'shadow',
-         control == 'financial_stress',
+  filter(spec == 'main',
+         control == 'both',
          index == 'epu',
-         thresh == '0.7') %>% 
-  ggplot(aes(x = t, y = median)) +
+         thresh == '0.5') %>% 
+  ggplot(aes(x = t, y = median, ymin = lower, ymax = upper)) +
   geom_line(linewidth = lw) +
   geom_hline(yintercept = 0, color = "red") +
+  geom_ribbon(alpha = .2)+
   facet_wrap(regime~label, scales = "free")+
@@
 ```
+
 ```{=latex}
-\caption{Regime specific impulse responses, updated specification. \label{fig:epumedirfs}}
+\caption{Regime specific impulse responses, updated specification. \label{fig:updatedirfs}}
 \bigskip
 \footnotesize
-\textit{Notes}: Interest rates are measured by the Shadow Rate, the threshold value is set to the 70th percentile of the EPU index.
+\textit{Notes}: In the updated specification, interest rates are measured by the FFR, the threshold value is set to the median of the EPU index, lagged values of the EPU and the St.
+Louis FED Financial Stress Indicator are added as controls.
 \end{figure}
@@
 
+These results indicate that some of the findings from both @aastveit2017economic, @aquino2022does and @horvathTBAuncertain can be true at the same time.
+In high uncertainty times, the impact of the monetary shock is considerably elevated on impact, and the first few quarters after the shock.
+However, dissipates much more rapidly from the variables, even reverting from disinflation to inflation around 12-16 quarters after the shock.
+The larger impact around the time of the shock supports the idea, that inflates the variance of expectations, and thereby increasing the impact of shocks.
+
+On the other hand, in times of low uncertainty, the impact is comparatively smaller, however the persistence of the impulse responses is greater, leading to the the adjustment of prices and output being slower, staying below 0 over the 20 quarter horizon.
+This indicates lower variance and more stability in the expectations of agents in the economy.
+
 ## Narrative of uncertainty
 
-First, the "narrative" of the uncertainty index could matter.
-In @baker2016measuring, additional to the EPU index, several categorical indices are constructed^[Please refer to Figure \ref{fig:epucat}.].
+In @baker2016measuring, additional to the EPU index, several categorical indices are constructed[^Please refer to Figure \ref{fig:epucat}.].
+To investigate how the underlying cause behind uncertainty interacts with the transmission mechanism, I replicate the estimation with each of these categorical indices.
+For easier comparability, I calculate sacrifice ratios - foregone output as a fraction of deterred inflation - from the 20-quarter accumulated impulse responses.
+
 Replicating the simulations with each categorical index could provide insight into how such narratives can impact the transmission mechanism.
 Notably, uncertainty specifically related to monetary policy should be the most insightful.
-Below in Figure \ref{fig:narrative}
+The sacrifice ratios can be seen in Figure \ref{fig:narrative} below.
 
@@
   filter(label != 'R',
-         spec == 'shadow',
+         spec == 'main',
          index != 'epu',
-         control == 'financial_stress') %>% 
+         control == 'both') %>% 
   group_by(label, regime, spec, control, thresh, index) %>% 
@@
 \footnotesize
-\textit{Notes}: The sacrifice ratios are calculated from the accumulated impact of the policy shocks over the 20 quarter horizon.  
+\textit{Notes}: The Full control TVAR model is used for estimating the data behind the sacrifice ratios.  
 \end{figure}
@@
 
-## Endogeneity
+\textcolor{red}{CALCULATE OVER DIFFERENT HORIZONS INSTEAD OF THE TWO MU-s YOU DUMMY! SHORT TERM GAINS OVER LONG TERM LOSSES}
 
@@
 ```
-
 ```{=latex}
@@
 ```
-
 ## Alternative model specifications
@@
 
+```{=tex}
 \begin{equation} \label{eq:2}
@@
 \end{equation}
-
+```
+```{=tex}
 \begin{equation} \label{eq:3}
@@
 \end{equation}
+```
+```{=tex}
+\begin{equation} \label{eq:4}
+  Y_{t} = \sum\limits_{k=1}^{4}\Theta_{high}I(X_{t} \ge \mu)Y_{t-k} + \sum\limits_{k=1}^{4}\Theta_{low}I(X_{t} < \mu)Y_{t-k} + + \sum\limits_{k=1}^{4}\beta X_{t-k} + \sum\limits_{k=1}^{4}\gamma Z_{t-k} + \epsilon_{t}
+\end{equation}
+```
 
@@
 ```
-
 As discussed in the Data section of this paper, there are alternative interest rate measures to calculate policy shocks from.
@@
 ```
-
 As seen in Figure \ref{fig:shockmes}, the difference between responses obtained when the policy shock is calculated using the FFR and the 1-Year Treasury Yield show little difference, however accounting for implied rates via the Shadow Rate has considerable impact on the results.

### 2023-10-17 US/codes/endog.R

@@ -1,0 +1,155 @@
+library(fredr)
+library(tidyverse)
+library(rlang)
+library(lubridate)
+library(lmtest)
+
+source('US/codes/data.R')
+invisible(paste('US/codes/varsignr/', list.files('US/codes/varsignr') , sep = '') %>% lapply(source))
+
+
+residuals <- read_rds('US/data/estout_clean.Rds')[['innovs']]
+
+uncertainty <- data[['uncertainty']]
+base_lin_data <- data[['macro']] 
+
+base_lin <- rwz.reject(base_lin_data %>% 
+                         select(-date) %>% 
+                         ts(),
+                       constrained = c(+1, -2, -3),
+                       nlags = 4, 
+                       draws = 20000, 
+                       subdraws = 200, 
+                       nkeep = 1000, 
+                       KMIN = 1, 
+                       KMAX = 6, 
+                       constant = TRUE, 
+                       steps = 21)
+
+res_lin <- bind_cols(
+  base_lin_data %>% 
+  select(date) %>% 
+  slice(-(1:4)), 
+  t(apply(base_lin[['SHOCKS']], 
+          2, 
+          quantile, 
+          probs = c(0.5, .16, 1-.16))) %>% 
+    as_tibble() %>% 
+    rename(median = `50%`,
+           lower = `16%`,
+           upper = `84%`)
+)
+
+
+
+
+granger_clean <- function(residuals, uncertainty){
+df <- inner_join(residuals, uncertainty)
+
+shock <-df %>% 
+  select(median) %>% 
+  ts() 
+
+unc <- df %>% 
+  select(epu) %>% 
+  ts()
+
+#Does the shock granger cause uncertainty
+test_shocks <- lmtest::grangertest(shock,
+                    unc,
+                    order = 4)
+test_shocks <- test_shocks['Pr(>F)'] %>% 
+  as_tibble() %>% 
+  drop_na() %>% 
+  rename(p = `Pr(>F)`) %>% 
+  mutate(test = 'H0: Shock does not cause uncertainty')
+
+#Does uncertainty granger cause the shock
+test_unc <- lmtest::grangertest(unc,
+                    shock,
+                    order = 4)
+test_unc<- test_unc['Pr(>F)'] %>% 
+  as_tibble() %>% 
+  drop_na() %>% 
+  rename(p = `Pr(>F)`) %>% 
+  mutate(test = 'H0: Uncertainty does not cause shock')
[…]

### 2023-10-17 US/codes/main.R

@@ -164,3 +164,5 @@ stopCluster(cl)
 saveRDS(estout, 'US/data/estout.Rds')
 
 source('US/codes/estout_cleanup.R')
+suppressMessages(source('US/codes/endog.R'))
+

### 2023-10-18 US/draft.Rmd
Completed draft V1

@@
 I employ a Threshold VAR model with two states dependent on the level of uncertainty, and identify regime specific shocks using sign-restrictions.
-Results suggest that \\textcolor{red}{FILL IN THE BLANK}.
-Using a number of alternative specifications as robustness checks \\textcolor{red}{FILL IN THIS BLANK TOO}.
-Implications for policymakers are \\textcolor{red}{BLANK NUMBER THREE TO BE FILLED}."
+Results suggest that the short run contraction induced by the policy shock is large on the short term, however dissipates more quickly.
+Comparatively in low uncertainty times, the contraction is smaller with policymakers having a smoother control over the path the economy takes.
+The short-run larger impact can be leveraged to reduce inflation in a more cost-effective way, however the uncerlying cause of uncertainty needs to be carefully considered.
+"
 fontsize: 12pt
@@
 
-While the main implications are clear, there is still room for further investigation.
-As a result of using a wide variety of countries for the estimation, some factors - such as a more nuanced investigation into the source of economic uncertainty; checking to what extent the Zero Lower Bound influences the results; whether or not excess financial stress caused by uncertainty is reflected in macroeconomic aggregates; or whether uncertainty is endogenous to policy shocks - was not possible to incorporate.
-This paper intends to fill this knowledge gap, by changing from a birds-eye view to a more microscopic perspective focusing on implications solely from the United States.
+Works such as @aastveit2017economic and @aquino2022does, however, find opposing results, suggesting that uncertainty reduces transmission efficacy.
+The former use an interacted SVAR approach, where the regime indicator is interacted only with interest rates in the equations, and take the very far ends of the uncertainty distribution to define high and low uncertainty regimes.
+The latter criticizes this and argues for the use of a grid-search algorithm[^1] to find the optimal threshold level for the regime switches.
+Additionally their specification introduces the threshold assumption into the interest rate equation of the SVAR model - implying a state dependent Taylor Rule.
 
+[^1]: Results from this algorithm in the paper suggest the regime switches are at the 57th, 58th and 65th percentile of the historical distribution depending on the index.
+While it is true, that the correct specification of the threshold value is important for goodness-of-fit, the more ad-hoc definition of high uncertainty at the median and the 70th percentile is kept as in @horvathTBAuncertain for the sake of simplicity and comparability.
+
+\textit{What might drive the key differences?}
+
+Firstly, the use of global versus country-specific uncertainty measures.
+While the two are not dissimilar concepts, they could carry different implications.
+As noted in @ilyna2023geo, heightened uncertainty is a channels of geoeconomic fragmentation, which is likely better grasped by the global measures, and thus carry different implications from times of elevated country-specific uncertainty.
+
+Secondly, the application of the threshold assumption.
+@aastveit2017economic and @aquino2022does use partial threshold models, whereas in @horvathTBAuncertain we argue that for correctly measuring the transmission mechanism, it is important to take into account how the entire dynamic system describing the economy changes at the regime switches.
+
+And thirdly, the richness of the datasets used for the estimations.
+In the global context, more compact, trivariate models were estimated to allow for a wide range of countries to be incorporated.
+However, relying solely on US data allows for a more rigorous model with additional variables as seen in @aastveit2017economic and @aquino2022does.
+
+This paper intends to fill this knowledge gap, by changing from a birds-eye view to a more microscopic perspective focusing on implications implications the United States.
 I adopt the same Threshold Vector-Autoregression (TVAR) model as well as sign-restriction approach to identify regime specific shocks.
@@
 
-However, given the wider range of data available for the US, there is room for a number of alternative specifications, such as using the @wu2016measuring shadow rate to measure implied interest rates at the ZLB, or incorporate the St.
-Louis Fed Financial Stress Index as a control variable in the model.
+However, relying on US data opens up the possibility to extend the variables (e.g.
+adding financial stress[^2], or checking for the robustness of the results with the @wu2016measuring shadow rate).
+Additionally, switching from global uncertainty aggregates to the US specific Economic Policy Uncertainty (EPU) index of @baker2016measuring allows for a deeper investigation of how the underlying cause of uncertainty impacts the transmission mechanism, as the authors construct category-specific EPU indices (e.g.: indices specific to Monetary Policy, Financial Regulation, or Sovereign debt and currency crises).
 
-As opposed to using the global versions of uncertainty indices, I switch to the US specific ones.
-For the main measure of uncertainty, I use the @baker2016measuring Economic Policy Uncertainty (EPU) index.
-As the EPU for the US has a number of supplementary Categorical indices (e.g.: indices specific to Monetary Policy, Financial Regulation, or Sovereign debt and currency crises) available, I also explore how different sources of uncertainty influence the transmission mechanism.
+[^2]: As shown by @li2010financial and @fry2016effects, in times of high financial stress, monetary policy can potentially become more effective.
+As elevated levels of uncertainty are likely associated with higher financial stress, is necessary, as shown later.
+For this reason, additional to the standard macroeconomic aggregates, the St.
+Louis Fed Financial Stress Index is considered as a control variable.
 
-The potential endogeneity issue between uncertainty and policy shocks is addressed using simple Granger-causality tests, as well as explicitly incorporating the uncertainty index into the TVAR model as a control variable.
+Switching from global uncertainty measures to the US-specific EPU might bring up endogeneity concerns, as it is not too far-fetched to assume policy shocks driving the policy aspect of the EPU.
+This is addressed using simple Granger-causality tests, as well as explicitly incorporating the uncertainty index into the model.
 
-Results from a TVAR model estimated with a full set of controls indicates that 
-\textcolor{red}{WRITE SOMETHING ABOUT THE RESULTS HERE 1-2 PARAGRAPHS.}
+Results from an extended TVAR model estimated with a full set of controls indicates that uncertainty inflates the variance of outcomes, exacerbating the impact of shocks in the short run, however in such times the impact of shocks is far less persistent.
+This can be leveraged by central banks, as the larger shock affects inflation disproportionately more, and as such, a lower short-run sacrifice ratio can be achieved.
+The source of uncertainty has to carefully be considered, however, as factors such as debt or trade policy concerns can greatly increase the cost of deterring inflation both on the short run, as well as over longer horizons.
 
@@
 
-Works such as @aastveit2017economic and @aquino2022does investigate the state-dependency of monetary shocks conditional on economic uncertainty.
-While the methods differ slightly, their results suggest that economic uncertainty hinders the transmission completely, as the responses of both output and inflation are decreased in high uncertainty times compared to the responses in times of low uncertainty.
-
-@aastveit2017economic uses an interacted SVAR approach, where the uncertainty-based regime indicator is interacted with only the interest rate in the model, all else is kept in a linear form.
-The regimes are selected based on the very extreme sides of the distribution however, which is criticized by @aquino2022does.
-The latter introduces the threshold assumption into the interest rate equation of the SVAR model - implying a state dependent Taylor Rule - and additionally classify the high and low regimes using a grid-search algorithm[^1].
-
-[^1]: Results from this algorithm in the paper suggest the regime switches are at the 57th, 58th and 65th percentile of the historical distribution depending on the index.
-While it is true, that the correct specification of the threshold value is important for goodness-of-fit, the more ad-hoc definition of high uncertainty at the median and the 70th percentile is kept as in @horvathTBAuncertain for the sake of simplicity and comparability.
-
-While theoretical arguments are compelling, taking into account how the equations describing the evolution of inflation and economic activity change across regimes should be paramount in measuring how the transmission changes with the uncertainty regimes.
-Additionally, arguments can be made for i) high uncertainty inflating the variance of outcomes in periods affected and thus enlarging the impact of all shocks; ii) uncertainty influencing the future expectations of agents; iii) uncertainty can introduce excess financial stress and thus influencing the transmission through this channel; or iv) uncertainty could prove to be endogenous to policy shocks.
-
-Works that might support the results of this paper could be @gbohoui2021uncertainty where an increase in investment multipliers is shown in times of high uncertainty, implying larger and longer lasting effects; or @ying2022role finding that monetary policy uncertainty moderates the transmission of forward guidance shocks on long-term yields, but not the transmission of Federal Funds Rate or asset price shocks, furthermore, the moderation effect is shown to be a consequence of changes in the term-premium and not through the channel of short rates.
+Some additional papers, results shown here are more in line with are @gbohoui2021uncertainty where an increase in investment multipliers is shown in times of high uncertainty, implying larger and longer lasting effects; or @ying2022role finding that monetary policy uncertainty moderates the transmission of forward guidance shocks on long-term yields, but not the transmission of Federal Funds Rate or asset price shocks, furthermore, the moderation effect is shown to be a consequence of changes in the term-premium and not through the channel of short rates.
 
@@
 
-In order to keep the estimation streamlined, I rely on quarterly macroeconomic aggregates from `r max(min(data2[['uncertainty']][['date']]), min(data2[['macro']][['date']])) %>% quarterly_format()` to `r min(max(data2[['uncertainty']][['date']]), max(data2[['macro']][['date']])) %>% quarterly_format()`[^2] as the core set of data behind the estimations.
+In order to keep the estimation streamlined, I rely on quarterly macroeconomic aggregates from `r max(min(data2[['uncertainty']][['date']]), min(data2[['macro']][['date']])) %>% quarterly_format()` to `r min(max(data2[['uncertainty']][['date']]), max(data2[['macro']][['date']])) %>% quarterly_format()`[^3] as the core set of data behind the estimations.
 The three main aggregates are interest rates, quarterly inflation rates - measured as the first difference of the CPI - and quarterly real GDP growth rates - measured as the first difference of the chain volume index calculated from quarterly real GDP.
 
-[^2]: The range of data does change in some specifications, as the Shadow Rate or the St. Louis Fed Financial Stress Index are only available on a shorter period of time.
+[^3]: The range of data does change in some specifications, as the Shadow Rate or the St. Louis Fed Financial Stress Index are only available on a shorter period of time.
 
-Figure \ref{fig:interest} below shows the three measures of interest rates I consider in different specifications.
-In the first set of estimations, policy shocks are measured as innovations of the Federal Funds Rate (FFR).
-In order to add some further nuance to the estimations, two other series are also considered.
+Figure \ref{fig:interest} below shows three measures of interest rates I consider in different specifications.
+Primarily, however, I rely on the Federal Funds Rate (FFR) as the policy instrument to estimate interest rate innovations.
 
@@
 ```
-In an alternative specifications the FFR is replaced with the 1-Year Treasury Yield.
-While the two show high co-movement, the latter may carry some additional information such as market expectations or yield curve dynamics, that might have an impact on the results.
-
-A further specification is considered where interest rates are measured by the @wu2016measuring shadow rate.
-While through the majority of the sample, the FFR and its shadow rate are highly correlated, the latter contains information about the implied rate of interest in the time period between 2008 and 2015 or 2020 and 2021 where the interest rates were close to the Zero Lower Bound.
-Estimating with the shadow rate is important as sign-restriction algorithms could lead to biased outcomes as a result of the ZLB periods.
-
-As shown by @li2010financial and @fry2016effects, in times of high financial stress, monetary policy can potentially become more effective.
-As elevated levels of uncertainty are likely associated with higher financial stress, it might be necessary to control for it as well.
-For this reason, additional to the standard macroeconomic aggregates, the St.
-Louis Fed Financial Stress Index is considered as a control variable.
+In robustness specifications, the 1-Year Treasury Yield and the @wu2016measuring shadow rate are considered as replacements of the FFR in estimating the model.
+While these three measures show high co-movement, the former - to some extent - may carry some additional information such as market expectations or yield curve dynamics - while the latter shows the implied rate of interest, which significantly differs from headline rates close to the ZLB.
 
@@
 Using the articles, automated text queries are conducted in order to find articles containing terms related to "uncertainty" and either "economic" or "policy".
-[^3]
+[^4]
 
-[^3]: The data series of uncertainty measures along with more detailed descriptions of each can be retrieved from [policyuncertainty.com](https://www.policyuncertainty.com/).
+[^4]: The data series of uncertainty measures along with more detailed descriptions of each can be retrieved from [policyuncertainty.com](https://www.policyuncertainty.com/).
 
@@
 
-It is important to mention that there are a number of alternative[^4] measurements of uncertainty in the literature.
+It is important to mention that there are a number of alternative[^5] measurements of uncertainty in the literature.
 These indices are for example the Equity Market-related Economic Uncertainty Index of @baker2016measuring; the Geopolitical Risk Index and the Trade Policy Uncertainty Index of @caldara2022measuring and @caldara2020economic; the World Uncertainty Index of @ahir2022world; as well as the VIX Index and the TED spread.
@@
 
-[^4]: These measures can be seen in Figure \ref{fig:unces} of the Appendix.
+[^5]: These measures can be seen in Figure \ref{fig:unces} of the Appendix.
 
@@
 Secondly, I check if the results hold if I change the construction of the regime indicator from $X_{t}$ to $X_{t-1}$, making it pre-determined, and thereby eliminating the potential of estimating the impact of two contemporaneous shocks.
-Thirdly, I incorporate the uncertainty index in its levels as well as a financial stress indicator in order to void their effects from the aggregates during the estimation[^5].
-
-[^5]: Please refer to the [Alternative model specifications] of the Appendix for further details on how Equation \ref{eq:1} above changes
+Thirdly, I incorporate the uncertainty index in its levels as well as a financial stress indicator in order to void their effects from the aggregates during the estimation.
 
@@
 All computations are done in the R programming language relying on the work of @Danne2015 for the implementation of the sign-restriction algorithms.
-
-\textcolor{red}{I SHOULD WRITE SOMETHING ABOUT ENDOGENEITY CONCERNS AND HOW THEY ARE ADDRESSED.}
 
@@
 ```
-There are some noteworthy differences from these results compared to previous literature.
-As discussed in the Literature section of this paper, @aastveit2017economic and @aquino2022does show that the transmission efficacy is eroded by high economic uncertainty, while @horvathTBAuncertain find a considerably more prolonged period of contraction induced by monetary shocks in high uncertainty.
+Results from this estimate suggest that the contemporaneous drop in output is slightly larger in low uncertainty, with the following quarters showing a similar level of contraction.
+Alongside this, a prolongued disinflation caused by the monetary shock can be observed in the high uncertainty regime.
+Comparing the two threshold levels, the changes in the impulse responses are close to negligible aside from the change in contemporaneous effects seen in the low regime.
 
-Results from this estimate suggest that the contemporaneous drop in output is slightly larger in low uncertainty, with the following quarters showing a similar level of contraction.
-Alongside this, disinflation caused by the monetary shock is significantly enhanced in the high uncertainty regime compared to the low uncertainty one.
-Comparing the median and the 70th percentile as threshold levels for the regime switches, the responses of inflation in the lower regime and of output in the higher regime are not too dissimilar, however their opposite uncertainty regime counterparts show some key differences.
-
-Firstly, the output contraction in the low regime is elevated when switching to the higher uncertainty threshold.
-This is in line with the findings of @horvathTBAuncertain, as allocating a larger proportion of observations previously labelled as high uncertainty would support the notion of larger contraction of high uncertainty times.
-Secondly, we can observe that the response of inflation shifts upwards as we increase the threshold level - albeit retaining a similar shape.
-This lies closer to the findings of @aastveit2017economic and @aquino2022does as the implication from this suggest that more extreme levels of uncertainty would erode the transmission.
-
-Why might these results differ so drastically from previous works?
-Firstly, in @horvathTBAuncertain globally aggregated uncertainty measures are used.
-As noted in @ilyna2023geo, heightened uncertainty is a channels of geoeconomic fragmentation, and uncertainty measured on the global level is likely a better proxy of this, thereby having somewhat different implications.
-Additionally, @aastveit2017economic and @aquino2022does use a wider range of variables, which can help isolate how uncertainty interacts with the transmission mechanism.
+These results are highly counter-intuitive, as they neither show reduced transmission efficacy, nor a prolonged contraction induced by the interest rate shock in high uncertainty times.
+The following subsections further investigate factors that might be key in shaping the regime specific responses.
 
@@
 Testing for endogeneity is done by running Granger causality tests between the EPU index and the posterior median of the residuals obtained from alternative specifications considered.
-The results from a select number of these is shown here in Table \ref{tab:grang}^[The Linear model is added as a point of comparison for the TVAR models.
+The results from a select number of these is shown here[^6] in Table \ref{tab:grang}.
+
+[^6]: The Linear model is added as a point of comparison for the TVAR models.
 It is estimated on the FFR, CPI inflation rate and GDP growth rate without further control variables.
 Full control TVAR represents models estimated with lagged values of the EPU index and the St.
-Louis FED Financial Stress Indicator added as controls.].
+Louis FED Financial Stress Indicator added as controls.
 
@@
 ```
-
 The results of these tests suggests that in most cases we fail to reject the null, and thus policy shocks do not drive uncertainty, neither the other way around.
@@
 Firstly, the likelihood of uncertainty being a driver of policy shocks considerably increases when switching from the median to the 70th percentile threshold.
-This is likely due to the low number of observations in the regime.
-Secondly, using the Shadow rate^[Why the Shadow rate might be a more desirable measure is further discussed in the Robustness checks of the Appendix.] over the FFR also considerably reduces this likelihood.
+This is likely due to the low number of observations in the regime, and thus an increased likelihood of policy action being taken to mitigate the negative impact of uncertainty.
+Secondly, using the Shadow rate[^7] over the FFR also considerably reduces this likelihood.
 Given these, the superior specification should be a TVAR with a full set of controls estimated with the threshold level set to the median.
@@
 
-## An updated specification
+[^7]: Why the Shadow rate might be a more desirable measure is further discussed in the Robustness checks of the Appendix.
 
-Through conducting a number of robustness^[Discussed in further detail in the Appendix.] and endogeneity tests, we arrive at a model that should grasp best how uncertainty influences the transmission of policy shocks.
+## Extended model
+
+Through conducting a number of robustness[^8] and endogeneity tests, we arrive at a model that should grasp best how uncertainty influences the transmission of policy shocks.
 Figure \ref{fig:updatedirfs} below shows the impulse responses obtained with the new specification after controlling for the level of uncertainty and financial stress.
+
+[^8]: Discussed in further detail in the Appendix.
 
@@
 ```{=latex}
-\caption{Regime specific impulse responses, updated specification. \label{fig:updatedirfs}}
+\caption{Regime specific impulse responses, extended model. \label{fig:updatedirfs}}
 \bigskip
 \footnotesize
-\textit{Notes}: In the updated specification, interest rates are measured by the FFR, the threshold value is set to the median of the EPU index, lagged values of the EPU and the St.
+\textit{Notes}: In the extended model, interest rates are measured by the FFR, the threshold value is set to the median of the EPU index, lagged values of the EPU and the St.
 Louis FED Financial Stress Indicator are added as controls.
@@
 ```
-
-These results indicate that some of the findings from both @aastveit2017economic, @aquino2022does and @horvathTBAuncertain can be true at the same time.
+Results from the extended model imply that to some extent findings from the previously cited literature on this question all have some truths to them.
 In high uncertainty times, the impact of the monetary shock is considerably elevated on impact, and the first few quarters after the shock.
 However, dissipates much more rapidly from the variables, even reverting from disinflation to inflation around 12-16 quarters after the shock.
-The larger impact around the time of the shock supports the idea, that inflates the variance of expectations, and thereby increasing the impact of shocks.
+The larger impact around the time of the shock supports the idea, that uncertainty inflates the variance of expectations, and thereby increasing the impact of shocks.
 
@@
 
-In @baker2016measuring, additional to the EPU index, several categorical indices are constructed[^Please refer to Figure \ref{fig:epucat}.].
+In @baker2016measuring, additional to the EPU index, several categorical indices are constructed[^9].
 To investigate how the underlying cause behind uncertainty interacts with the transmission mechanism, I replicate the estimation with each of these categorical indices.
-For easier comparability, I calculate sacrifice ratios - foregone output as a fraction of deterred inflation - from the 20-quarter accumulated impulse responses.
+To compare how much the narrative of uncertainty matters, I calculate sacrifice ratios - foregone output as a fraction of deterred inflation - from accumulated impulse responses over 4, 12 and 20 quarters for both high and low uncertainty regimes.
+To further aggregate, I calculate the sacrifice ratio of the high uncertainty regime relative to the low uncertainty regime.
+These can be seen below in Figure \ref{fig:narrative}.
 
-Replicating the simulations with each categorical index could provide insight into how such narratives can impact the transmission mechanism.
-Notably, uncertainty specifically related to monetary policy should be the most insightful.
-The sacrifice ratios can be seen in Figure \ref{fig:narrative} below.
+[^9]: Please refer to Figure \ref{fig:epucat}.
 
@@
 ```
-```{r narrative, fig.height=10}
+```{r narrative, fig.width=10}
 irfs %>% 
@@
          spec == 'main',
-         index != 'epu',
-         control == 'both') %>% 
+         #index != 'epu',
+         control == 'both',
+         thresh == '0.5') %>% 
   group_by(label, regime, spec, control, thresh, index) %>% 
@@
                                    .dir = 'forward')) %>% 
-  filter(t == max(t)) %>% 
-  select(-t, -upper, -lower) %>% 
+  filter(t  %in% c(4, 12, 20)) %>%
+  mutate(t = case_when(t == 4 ~ '1-year',
+                       t == 12 ~ '3-year',
+                       t == 20 ~ '5-year')) %>% 
+  select(-upper, -lower) %>% 
   spread(label, median) %>% 
@@
          regime = str_remove_all(regime, ' uncertainty'),
-         thresh = case_when(thresh == '0.5' ~ 'Median',
-                            thresh == '0.7' ~ '70th Perc.'),
-         thresh = factor(thresh, levels = c('Median', '70th Perc.')),
          index = case_when(index == 'epu' ~ 'EPU',
@@
                            index == 'epu_trade' ~ 'EPU: Trade Policy')) %>% 
-  ggplot(aes(x = regime, y = sac, fill = regime)) + 
-  geom_col( color = 'black', width = 0.3, position = position_dodge(width = 0.1)) +
-  facet_grid(index~thresh, scales = 'free') +
+  select(-PI, -Y) %>% 
+  spread(regime, sac) %>% 
+  mutate(relsac = High/Low) %>% 
+  ggplot(aes(x = t, y = relsac, fill = t)) +
+  geom_col(color = 'black') +
+  facet_wrap(~index, nrow = 1) +
   theme_minimal()+
   theme(legend.position = 'none') +
+  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
   labs(x = '',
        y = '') +
-  scale_fill_manual(breaks = c('High',
-                                'Low'),
-                     values = c('#8B0000',
-                                '#00008B'))
-
+  scale_fill_brewer('Set 1', direction = -1)
 ```
@@
 \footnotesize
-\textit{Notes}: The Full control TVAR model is used for estimating the data behind the sacrifice ratios.  
+\textit{Notes}: The ratios are calculated from impulse responses obtained using the extended model.  
 \end{figure}
 ```
+For most measures, the relative sacrifice ratio decreases over time, which is no surprise given the results seen in Figure \ref{fig:updatedirfs}.
+However, many of these relative sacrifice ratios show values below one, meaning the trade-off between reducing inflation and forgoing growth is smaller in high uncertainty.
+Calculated over the longer run, this is not necessarily counterintuitive, as responses in the low uncertainty regime stay more persistently below zero, however on the shorter run (1-year) this implies that policy authorities can leverage the higher variance the impact on prices is reflected disproportionately higher on prices than output - however taking into account the underlying cause matters significantly.
 
-\textcolor{red}{CALCULATE OVER DIFFERENT HORIZONS INSTEAD OF THE TWO MU-s YOU DUMMY! SHORT TERM GAINS OVER LONG TERM LOSSES}
+An exception in Figure \ref{fig:narrative} being uncertainty related to sovereign debt and currency crises.
+Here the high uncertainty sacrifice ratio is larger on the short run and increases over longer horizons. High uncertainty around debt can increase borrowing costs for the government - potentially necessitating austerity measures.
+An increased demand from governments for debt financing would likely crowd out private investments.
+A rate hike in such context would add further pressure on government debt schedules and thus enlarge the crowding out effect by either the issuance of new debt or tax increases, carrying over to future periods.
+
+Uncertainty related to trade policy also has a considerable impact on the cost of deterring inflation, albeit this slightly decreases over time.
+Trade policy uncertainty is directly connected to supply chain disruptions, reflecting in a reduced export demand and higher import costs.
+Rate hikes can further affect this through the exchange rate channel, as an appreciating exchange rate would further contract exports and mitigate the potential excess import fees.
+
+Uncertainty related to financial regulation also increases the cost of reducing inflation through disrupting financial markets in the short run, however the higher cost dissipates over longer horizons.
+Counter-intuitively, monetary policy uncertainty does not cause neither short-run, nor long-run excess cost in deterring inflation.
+However, it is likely an artifact of the high variance of the monetary policy uncertainty categorical index having a high variance over time, making the regime assignment almost random.
 
 # Conclusion
+
+This paper studies the relationship between uncertainty and the monetary transmission mechanism in the US.
+A TVAR model is estimated over quarterly interest rates, inflation, and growth, also controlling for the level of uncertainty and financial stress.
+The EPU and its categorical indices are as threshold variables, and regime specific shocks are identified using sign restrictions.
+In the short run, high uncertainty is found to inflate the variance of expectations, and enlarge the impact of the interest rate shock.
+In low uncertainty times policymakers are able to contract demand more moderately, however more smoothly and over a more prolonged period.
+The narrative behind uncertainty has significant implications.
+While policymakers might be able to leverage the larger short-run impact through lesser sacrifice ratios, uncertainty related to debt, trade, and financial regulations make it more costly to anchor inflation.
 
@@
 ```
-
 where $X_{t-k}$ represents the inclusion of four lags of the uncertainty indicator, and $Z_{t-k}$ represent the inclusion of four lags of the St. Louis Fed Financial Stress Index.

### 2023-10-18 US/codes/main.R

@@ -166,3 +166,4 @@ saveRDS(estout, 'US/data/estout.Rds')
 source('US/codes/estout_cleanup.R')
 suppressMessages(source('US/codes/endog.R'))
 
+

### 2023-10-25 US/draft.Rmd
Tidied granger causality table

@@
 display = c('g', 'g', 'g', 'g'))`
+
+\bgroup
+
+\renewcommand{\baselinestretch}{0.75}\normalsize
+\textit{Notes:} \linebreak
+Test 1: $H_{0}$: Shock does not cause uncertainty \linebreak
+Test 2: $H_{0}$: Uncertainty does not cause shock
+
+\egroup
+
 \end{center}
 ```
+
 The results of these tests suggests that in most cases we fail to reject the null, and thus policy shocks do not drive uncertainty, neither the other way around.
-Interpreting the p-values as the likelihood of $H0$ to be true, there are a handful of conclusions to draw.
+Interpreting the p-values as the likelihood of $H_{0}$ to be true, there are a handful of conclusions to draw.
 The likelihood of policy shocks driving uncertainty stay reasonably high irrespective of specification, however, for the reverse causality case, it varies substantially.
@@
 An exception in Figure \ref{fig:narrative} being uncertainty related to sovereign debt and currency crises.
-Here the high uncertainty sacrifice ratio is larger on the short run and increases over longer horizons. High uncertainty around debt can increase borrowing costs for the government - potentially necessitating austerity measures.
+Here the high uncertainty sacrifice ratio is larger on the short run and increases over longer horizons.
+High uncertainty around debt can increase borrowing costs for the government - potentially necessitating austerity measures.
 An increased demand from governments for debt financing would likely crowd out private investments.

### 2023-10-25 US/codes/endog.R

@@ -62,7 +62,7 @@ test_shocks <- test_shocks['Pr(>F)'] %>%
   as_tibble() %>% 
   drop_na() %>% 
   rename(p = `Pr(>F)`) %>% 
-  mutate(test = 'H0: Shock does not cause uncertainty')
+  mutate(test = 'Test 1')
 
 #Does uncertainty granger cause the shock
 test_unc <- lmtest::grangertest(unc,
@@ -72,7 +72,7 @@ test_unc<- test_unc['Pr(>F)'] %>%
   as_tibble() %>% 
   drop_na() %>% 
   rename(p = `Pr(>F)`) %>% 
-  mutate(test = 'H0: Uncertainty does not cause shock')
+  mutate(test = 'Test 2')
 
 out <- bind_rows(test_unc, test_shocks)
 
@@ -146,8 +146,8 @@ granger_clean(residuals = residuals %>%
   mutate(model = 'Full control TVAR with shadow rate, 70th perc.')
 ) %>% 
   select(Model = model,
-         `H0: Shock does not cause uncertainty`,
-         `H0: Uncertainty does not cause shock`) %>% 
+         `Test 1`,
+         `Test 2`) %>% 
   mutate(across(-Model, ~round(.x, digits = 3)))
 
 

### 2023-10-25 US/draft.Rmd
Comments revision round 1

@@
 I employ a Threshold VAR model with two states dependent on the level of uncertainty, and identify regime specific shocks using sign-restrictions.
-Results suggest that the short run contraction induced by the policy shock is large on the short term, however dissipates more quickly.
+Results suggest that the short run contraction induced by the policy shock is larger when uncertainty is high, however, in such times, its impact dissipates more quickly.
 Comparatively in low uncertainty times, the contraction is smaller with policymakers having a smoother control over the path the economy takes.
@@
 
-This paper serves as a natural followup to @horvathTBAuncertain.
-The paper establishes how the interest rate channel of the monetary transmission intertwined with economic uncertainty influences economies from across the globe.
+The transmission of monetary policy interventions has been a cornerstone of macroeconomic research for academics and policymakers alike, as understanding how monetary policy can affect the real economy has consequential implications for evaluating the right policy strategy.
+Another field gaining increased attention, especially for empirical macroeconomists, is research on how uncertainty drives business cycles.
+The literature agrees that uncertainty is an important driver of macroeconomic demand, primarily influencing it through investment decisions.
+However, the interplay between uncertainty and the monetary transmission mechanism is relatively less investigated.
+This leads to the purpose of this paper: To extend the existing literature in quantifying how uncertainty - especially uncertainty related to economic policy - affects the transmission mechanism.
+
+Primarily, this paper is a natural followup to @horvathTBAuncertain.
+The paper establishes how the interest rate channel of monetary transmission in times of elevated global economic uncertainty arising from geoeconomic fragmentation influences economies.
 The results indicate that in times of high uncertainty, interest rate shocks lead to a larger contraction of macroeconomic demand, primarily reflecting in a larger and longer lasting slump in economic activity.
 
-Works such as @aastveit2017economic and @aquino2022does, however, find opposing results, suggesting that uncertainty reduces transmission efficacy.
+Earlier works in this field, such as @aastveit2017economic and @aquino2022does, however, find opposing results, suggesting that uncertainty reduces transmission efficacy.
 The former use an interacted SVAR approach, where the regime indicator is interacted only with interest rates in the equations, and take the very far ends of the uncertainty distribution to define high and low uncertainty regimes.
@@
 
-This paper intends to fill this knowledge gap, by changing from a birds-eye view to a more microscopic perspective focusing on implications implications the United States.
+This paper intends to fill this knowledge gap, by changing from a birds-eye view to a more microscopic perspective focusing on implications for the United States.
 I adopt the same Threshold Vector-Autoregression (TVAR) model as well as sign-restriction approach to identify regime specific shocks.
-Following the previous paper, the core data behind the estimations are macroeconomic aggregates on a quarterly frequency, consisting of interest rates, inflation and GDP.
+To measure the transmission of monetary shocks, I employ a set of quarterly macroeconomic as my core dataset comprising of interest rates, inflation and GDP, and as such, it is possible to measure the transmission of policy shocks to the real economy.
 
@@
 
-Following such works, @caldara2016macroeconomic show that financial and uncertainty shocks are pivotal drivers of business cycles, exemplified during events like the Great Recession.
+@caldara2016macroeconomic show that financial and uncertainty shocks are pivotal drivers of business cycles, exemplified during events like the Great Recession.
 Likewise, @bonciani2020international show that global financial uncertainty shocks bear down on economies, particularly in open economies during downturns.
@@
 Such techniques are used for research on monetary policy as well.
-@avdjiev2014credit show that policy shocks have a more pronounced impact during low-growth periods while monetary policy having a aggressive responses during booms; @schmidt2020risk find evidence for state-dependency on asset prices, with macro-risk and policy risk playing distinct roles for the transmission; while @li2010financial and later @fry2016effects highlight monetary policy efficacy in times of financial stress, emphasizing short-term trade-offs between output and inflation.
+@avdjiev2014credit show that policy shocks have a larger impact in recession periods, and that monetary policy reacts more aggressively to rising inflation during booms; @schmidt2020risk find evidence for state-dependency on asset prices, with macro-risk and policy risk playing distinct roles for the transmission; while @li2010financial and later @fry2016effects highlight monetary policy efficacy in times of financial stress, emphasizing short-term trade-offs between output and inflation.
 
@@
 
-[^3]: The range of data does change in some specifications, as the Shadow Rate or the St. Louis Fed Financial Stress Index are only available on a shorter period of time.
+[^3]: Data for interest rates, inflation and economic activity is available from much earlier, however the sample is restricted to start from `r max(min(data2[['uncertainty']][['date']]), min(data2[['macro']][['date']])) %>% quarterly_format()` as it is the earliest available data point for the EPU index.
+The range of data does change in some specifications, as the Shadow Rate or the St.
+Louis Fed Financial Stress Index are only available on a shorter period of time.
 
@@
 
-It is important to mention that there are a number of alternative[^5] measurements of uncertainty in the literature.
-These indices are for example the Equity Market-related Economic Uncertainty Index of @baker2016measuring; the Geopolitical Risk Index and the Trade Policy Uncertainty Index of @caldara2022measuring and @caldara2020economic; the World Uncertainty Index of @ahir2022world; as well as the VIX Index and the TED spread.
-In this paper, only the EPU and its Categorical indices are considered as measures of uncertainty.
+It is important to mention that there are a number of additional[^5] measurements of uncertainty in the literature.
+Measures such as the VIX Index and the TED spread using market data, as well as the search-term based Equity Market-related Economic Uncertainty Index of @baker2016measuring can be used to grasp uncertainty in financial and liquidity markets.
+@caldara2022measuring and @caldara2020economic use textual data to construct the Geopolitical Risk Index and the Trade Policy Uncertainty index which measure uncertainty around geopolitical and trade tensions.
+@ahir2022world use text searches in Country Intelligence Unit reports to construct the World Uncertainty Index, as a broader measure of uncertainty.
+As this paper focuses on the implications of uncertainty for monetary policy, I use only the EPU and its Categorical indices as they are the most tailored to grasping uncertainty around economic policy.
 
@@
 Secondly, I check if the results hold if I change the construction of the regime indicator from $X_{t}$ to $X_{t-1}$, making it pre-determined, and thereby eliminating the potential of estimating the impact of two contemporaneous shocks.
-Thirdly, I incorporate the uncertainty index in its levels as well as a financial stress indicator in order to void their effects from the aggregates during the estimation.
+Thirdly, I incorporate the uncertainty index in its levels as well as a financial stress indicator in order to purge their effects from the aggregates during the estimation.
 
@@
 
-The identification of shocks is done using sign-restrictions, as with a relatively compact set of variables, this is the most straight-forward method.
-As standard macroeconomic theory dictates, I assume that an increase in interest rates should be a negative demand-type shock, and thus lead to a contraction of output and prices.
-In other terms, a positive sign-restriction is imposed on the interest rates, and a negative on inflation rates as well as GDP growth.
-Following @uhlig2005effects, the restriction is imposed on the contemporaneous effects, as well as up to 6 quarters ahead after the shock.
+Empirically, an obstacle in estimating impulse responses that show the correct sign associated with monetary shocks is the price puzzle.
+Estimating a trivariate (interest rates - inflation - GDP) SVAR model yields impulse responses showing a rise in inflation as a consequence of an interest rate shock.
+This, however goes against standard macroeconomic theory, which states that increasing policy rates should contract macroeconomic demand and thus lead to a decline both in prices as well as output.
+Numerous authors (among others @hanson2004price, @giordani2004alternative, @demiralp2014still, @bishop2017anticipatory, @cloyne2016macroeconomic, @romer2004new) investigate the price puzzle, suggesting different explanations or ways of dealing with this phenomenon.
 
-The algorithms proposed in @uhlig2005effects has some shortcomings that should be addressed.
-As pointed out by @arias2018inference, using the penalty algorithm has some adverse consequences.
-Firstly, this algorithm can introduce some unintentional sign restrictions on unrestricted variables, thereby leading to biased results.
-For the main round of estimations, this is no issue, as I work with a full set of sign restrictions.
-However more notably, the penalty algorithm accepts a larger number of posterior draws, thus narrowing confidence bands.
-Opposing to this, the rejection algorithm only accepts posterior draws where the impulse responses that exactly fit the imposed sign-restrictions.
-Additionally, @rubio2010structural proposed a more efficient rejection algorithm, and thus their improved algorithm is implemented in this paper.
+In this paper, I follow a different branch of literature.
+Pioneered by @uhlig2005effects, it is possible to incorporate some prior beliefs into the estimation by imposing sign-restrictions on the impulse responses.
+His seminal work introduces a (full-Bayesian) rejection as well as a penalty algorithm for estimating SVAR-s as such.
+The former approach accepts impulse response draws only if they exactly fit the imposed restrictions.
+The latter accepts draws that violate the sign-restrictions using a goodness-of-fit based penalty score.
 
-In all cases, the models are estimated using Bayesian methods with a non-informative Wishart prior.
-The estimations yield a set of impulse response draws, which are aggregated into posterior-median impulse responses along with an 84% confidence interval.
+At first glance, the penalty approach might seem more desirable due to higher flexibility, however @arias2018inference show some adverse properties of this algorithm.
+Firstly, the penalty algorithm can artificially impose restrictions on the unrestricted variable, and secondly, the - by design - higher number of accepted responses narrows confidence bands.
+In a smaller model, such as the trivariate one with a full set of restrictions is no issue, however, if the model is extended with variables on which no restrictions are placed this would be an issue, and the narrower confidence bands would misrepresent the accuracy of the results.
+Additionally, @rubio2010structural improved on the efficiency of the rejection algorithm of @uhlig2005effects, and for these reasons, their iteration of it is used in this paper.
+
+In all estimations, Bayesian models are estimated with 4 lags and a non-informative Wishart prior.
+For the purposes of this paper, partial identification is sufficient, and as such, sign-restrictions are only imposed on monetary shocks.The sign-restriction is imposed to follow common macroeconomic assumptions, meaning a positive sign on interest rates and negative on inflation and output.
+The mirror image of the results can be obtained by reversing the sign restrictions, and I assume the impulse responses are independent of the size of the monetary shock.
+As such, studying sign and size asymmetries is left for future research.
+
+Posterior impulse responses are estimated in a bootstrap simulation with 200 impulse response draws for each 20000 MCMC replication.
+These are computed separately for both the low and the high uncertainty regime in each case.
+The accepted posterior impulse responses are aggregated into posterior-medians along with and 84% confidence interval.
 All computations are done in the R programming language relying on the work of @Danne2015 for the implementation of the sign-restriction algorithms.
@@
 ```
+
 Results from this estimate suggest that the contemporaneous drop in output is slightly larger in low uncertainty, with the following quarters showing a similar level of contraction.
@@
 Comparing the two threshold levels, the changes in the impulse responses are close to negligible aside from the change in contemporaneous effects seen in the low regime.
-
-These results are highly counter-intuitive, as they neither show reduced transmission efficacy, nor a prolonged contraction induced by the interest rate shock in high uncertainty times.
-The following subsections further investigate factors that might be key in shaping the regime specific responses.
+Such results are not supported by either of the previous literature.
+In the following subsections I investigate if other factors could be influential in shaping the impulse responses.
 
@@
 Interpreting the p-values as the likelihood of $H_{0}$ to be true, there are a handful of conclusions to draw.
-The likelihood of policy shocks driving uncertainty stay reasonably high irrespective of specification, however, for the reverse causality case, it varies substantially.
+The likelihood of policy shocks driving uncertainty stay reasonably low irrespective of specification, however, for the reverse causality case, it varies substantially.
 Firstly, the likelihood of uncertainty being a driver of policy shocks considerably increases when switching from the median to the 70th percentile threshold.
@@
 ```
-Results from the extended model imply that to some extent findings from the previously cited literature on this question all have some truths to them.
-In high uncertainty times, the impact of the monetary shock is considerably elevated on impact, and the first few quarters after the shock.
-However, dissipates much more rapidly from the variables, even reverting from disinflation to inflation around 12-16 quarters after the shock.
-The larger impact around the time of the shock supports the idea, that uncertainty inflates the variance of expectations, and thereby increasing the impact of shocks.
 
-On the other hand, in times of low uncertainty, the impact is comparatively smaller, however the persistence of the impulse responses is greater, leading to the the adjustment of prices and output being slower, staying below 0 over the 20 quarter horizon.
-This indicates lower variance and more stability in the expectations of agents in the economy.
+Partially, the results of the extended model are supported from findings of the previous literature.
+In low uncertainty times, we see that the policy shock has a comparatively smaller advantage, however, the adjustment is slower, steadier and more persistent over the 20 quarter horizon, likely meaning that the central bank has a better control over future expectations.
+However, in high uncertainty times, the initial contraction in the first few quarters is larger, and dissipates more quickly.
+Additionally, the estimated impulse responses are much less smooth, and around 12 quarters after the intervention, a rise in inflation can be observed.
+These could be signs that high uncertainty times cause the variance of future expectations to be larger, in turn causing shocks to have larger impacts on one hand, while on the other causing the less persistent responses.
 
@@
 In @baker2016measuring, additional to the EPU index, several categorical indices are constructed[^9].
-To investigate how the underlying cause behind uncertainty interacts with the transmission mechanism, I replicate the estimation with each of these categorical indices.
-To compare how much the narrative of uncertainty matters, I calculate sacrifice ratios - foregone output as a fraction of deterred inflation - from accumulated impulse responses over 4, 12 and 20 quarters for both high and low uncertainty regimes.
-To further aggregate, I calculate the sacrifice ratio of the high uncertainty regime relative to the low uncertainty regime.
-These can be seen below in Figure \ref{fig:narrative}.
+To investigate how the underlying cause behind uncertainty interacts with the transmission mechanism, I replicate the estimation with each of these Categorical indices.
+For this estimation I use the extended model with the threshold set to the median - as seen in the previous subsection - of each Categorical EPU index.
+
+To compare across estimates, I calculate sacrifice ratios as $Sacrifice Ratio = \frac{\Delta Output}{\Delta Inflation}$ for both uncertainty regimes.
+In essence, these capture how much output (growth) was forgone at the cost of deterring a percentage point of inflation.
+These ratios are calculated over a 1-year (4-quarter), 3-year (12-quarter) and 5-year (20-quarter) horizon.
+
+I perform a further round of aggregation, and calculate the relative sacrifice ratio of high uncertainty sacrifice ratios as a fraction of low uncertainty ones.
+These can be interpreted as how much more (less) the cost is of deterring inflation when uncertainty is high compared to low uncertainty times.
+The relative sacrifice ratios can be seen below in Figure \ref{fig:narrative}.
 
@@
 ```
-For most measures, the relative sacrifice ratio decreases over time, which is no surprise given the results seen in Figure \ref{fig:updatedirfs}.
-However, many of these relative sacrifice ratios show values below one, meaning the trade-off between reducing inflation and forgoing growth is smaller in high uncertainty.
-Calculated over the longer run, this is not necessarily counterintuitive, as responses in the low uncertainty regime stay more persistently below zero, however on the shorter run (1-year) this implies that policy authorities can leverage the higher variance the impact on prices is reflected disproportionately higher on prices than output - however taking into account the underlying cause matters significantly.
 
-An exception in Figure \ref{fig:narrative} being uncertainty related to sovereign debt and currency crises.
+For most of these, we can observe that the relative sacrifice ratio is decreasing over time.
+This is no surprise, given the results seen in Figure \ref{fig:updatedirfs}, seeing as the impact of the policy shock contracts output and prices more persistently in the low regime, and dissipates quicker in the high regime.
+Additionally, we can observe that for a number of the Categorical EPU indices, the relative sacrifice ratio starts and stays below one, meaning that deterring inflation comes at a lower cost.
+However, Figure \ref{fig:narrative} also warns us that the underlying cause behind uncertainty needs to be carefully considered.
+
+For example, uncertainty related to sovereign debt and currency crises tells a much different story compared to what can be seen from the other indices.
 Here the high uncertainty sacrifice ratio is larger on the short run and increases over longer horizons.
@@
 As noted in the the Methodology section, there could be two arguments against the specification used to obtain the results seen above.
-One of them being using time $t$ values of the EPU to specify the regimes, the other being that the variables should be voided of the share of variance explained by uncertainty.
+One of them being using time $t$ values of the EPU to specify the regimes, the other being that the variables should be purged of the share of variance explained by uncertainty.
 The former is achievable by changing Equation \ref{eq:1} to use $X_{t-1}$ instead of $X_{t}$ for specifying the regime indicator, while the latter is done as shown in \ref{eq:2}.
@@
 
-Voiding the effects of financial stress is also a crucial determinant in shaping the results.
+Purging the effects of financial stress is also a crucial determinant in shaping the results.
 As seen in Figure \ref{fig:finstress}, after including the financial stress indicator as a control variable, the responses shift downwards drastically, especially in the lower regime.

### 2023-10-27 US/codes/data.R
Create README.md

@@ -11,6 +11,7 @@ require(readxl)
     require(forecast)
     require(seasonal)
     require(tidyverse)
+    require(lubridate)
     tryCatch(
       value %>% 
         ts(frequency = freq,

### 2023-10-27 US/codes/expect.R

@@ -1,0 +1,54 @@
+library(tidyverse)
+library(fredr)
+
+fredr_set_key("cda47ae66b38ed7988c0a9c2ec80c94f")
+
+epu <- fredr('USEPUINDXD') %>% 
+  select(date, epu = value)
+
+expinf_cleveland <- fredr('EXPINF1YR') %>% 
+  select(date, expinf = value)
+
+expinf_mich <- fredr('MICH') %>% 
+  select(date, expinf = value)
+
+
+test1 <- inner_join(epu, expinf_cleveland) %>% 
+  drop_na() %>% 
+  mutate(med = median(epu),
+         regime = ifelse(epu > med, 'High Uncertainty', 'Low Uncertainty')) %>% 
+  select(regime, expinf) %>% 
+  var.test(expinf ~ regime, data = .,
+           alternative = 'less')
+
+
+test2 <- inner_join(epu, expinf_mich) %>% 
+  drop_na() %>% 
+  mutate(med = median(epu),
+         regime = ifelse(epu > med, 'High Uncertainty', 'Low Uncertainty')) %>% 
+  select(regime, expinf) %>% 
+  var.test(expinf ~ regime, data = .,
+           alternative = 'less')
+
+test1
+test2
+
+
+inner_join(epu, expinf_mich) %>% 
+  drop_na() %>% 
+  mutate(med = median(epu),
+         regime = ifelse(epu > med, 'High Uncertainty', 'Low Uncertainty')) %>% 
+  select(regime, expinf) %>% 
+  t.test(expinf ~ regime, data = .,
+           alternative = 'less')
+
+
+inner_join(epu, expinf_cleveland) %>% 
+  drop_na() %>% 
+  mutate(med = median(epu),
+         regime = ifelse(epu > med, 'High Uncertainty', 'Low Uncertainty')) %>% 
+  select(regime, expinf) %>% 
+  t.test(expinf ~ regime, data = .,
+           alternative = 'less')
+
+
